### 2017-03-28 misc/visu/paper.tex
initial commit

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\title{Insights in tuning EASY-Backfilling}

\author{
\IEEEauthorblockN{J\'er\^ome Lelong, Valentin Reis, Denis Trystram}
\IEEEauthorblockA{Grenoble Alpes University \\ firstname.lastname@imag.fr}
}

\author{J\'er\^ome Lelong \and Valentin Reis \and Denis Trystram}

\institute{Grenoble Alpes University, Grenoble, France \\ \email{firstname.lastname@imag.fr}}

\maketitle

\begin{abstract}

  EASY backfilling is a popular scheduling heuristic for allocating jobs in
  large scale High Performance Computing platforms. While its aggressive
  reservation mechanism is fast to execute and prevents job starvation, it
  remains largely suboptimal with respect to any scheduling objective. We
  consider the problem of tuning EASY using job reordering policies. More
  precisely, we propose a methodology based on the use of a simulator to choose
  the best heuristic for a given system, so that a cumulative scheduling
  objective such as the bounded slowdown is minimized on a yet unseen workload.
  This methodology introduces a risk on the maximum values of typical objective
  values, which are controlled in this paper by a queue thresholding mechanism.
  This new approach is evaluated through a comprehensive experimental campaign
  on several production logs. We show in particular that the performance does
  generalize on all systems under study. Indeed, the average bounded slowdown
  can be reduced between 86\% to 52\% of the its value for EASY on various
  workload logs from the Parallel Workload Archive.  This work studies
  empirical performance generalization of reordering policies in a robust
  framework and demonstrates that effort should be put on learning to reorder
  job queues.

\end{abstract}

\section{Introduction}

High Performance Computing platforms are evolving to extreme scale multi-cores.
Such systems provide the capabilities needed to solve large scaie complex
scientific, economic or social problems.  The number of processors will
drastically increase and more processing capabilities will obviously lead to
more data produced~\cite{DOEASCAC}.  Moreover, new computing systems are
expected to run more flexible workloads.  Seldom supported by the existing
managing resource systems, the future schedulers should take advantage of this
flexibility to optimize the performance of the system.  The extreme scale
generates a huge amount of data at run-time. Collecting relevant information is
a prerequisite for determining efficient allocations.

The resources of such platforms are usually the subject of competition by
many users submitting their jobs, and parallel job scheduling is the key
problem to solve. Efficient scheduling of parallel jobs is a challenging task
which promises great improvement in various ways, including improved machine
utilization, energy efficiency, throughput and response time. Not only are
scheduling problems usually computationally hard, but in HPC practice they are
also plagued with uncertainty as many parameters of the problem are unknown
while taking decisions. As a consequence, the actual production platforms
currently rely on very simple heuristics based on queues of submitted jobs,
ordered in various ways.  The most used heuristic is the well-known
EASY-backfilling policy~\cite{easy}. While EASY is simple, fast to execute and
prevents starvation, it does not fare especially well with respect to
cumulative cost metrics\footnote{These metrics are defined in section
\ref{sec:pbset}} such as the average Bounded Slowdown(\textbf{AvgBsld}) or the
average waiting time (\textbf{AvgWait}) of jobs.  Therefore, many HPC
researchers and system administrators intent to tune this heuristic by
reordering the \textit{job starting/reservation} and \textit{backfilling}
queues. Since such reordering of job queues may introduce starvation in the
scheduling, this results in a dilemna between expected and maximal cost. In
order to solve this dilemna, we introduce a thresholding mechanism that can
effectively manage the risk of reaching maximum objective values. This issue is
further complexified by the dependence of the relative scheduling performances
on system characteristics and workload profiles. We propose in this work the
use of simulations to select queue reordering policies.  Finally, we study the
expirical generalization and stability of this methodology and open doors to
further learning-based approaches.

Section~\ref{sec:pbset} describes the context and details the problem
statement. Section~\ref{sec:rw} reviews existing approaches from the
literature. Sections~\ref{sec:mmax} and~\ref{sec:th} will introduce our
approach , illustrating the discussion with results from the KTH-SP2 trace.
Section~\ref{sec:expe} validates the approach using a comprehensive
experimental campaign on 7 logs from the Parallel Workload
Archive\cite{Feitelson20142967}.

\section{Related Works}
\label{sec:rw}

While parallel job scheduling is a well studied theoretical
problem~\cite{leung2004handbook}, the practical ramifications, varying
hypotheses, and inherent uncertainty of the problem in HPC have driven
practicioners and researchers alike to use and study simple heuristics.
The two most used heuristics for HPC systems are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} Backfilling.

While Conservative Backfilling offers many
advantages~\cite{srinivasan2002characterization}, it has a computational
overhead, perhaps explaining why most of the machines from the
top500~\cite{top500} ranking still use at the time of publication a variant of
EASY Backfilling.

There is a large body of work that seeks to improve EASY. Indeed, while the
heuristic is used by various resource and job management software (most notably
SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tuning by systems
administrators.

There are multiple works that explore how to tune EASY by reordering waiting
and/or backfilling queues\cite{Tsafrir_easypp_2005}, sometimes even in a
randomized manner~\cite{1592720}, as well as some
implementations~\cite{Jackson2001}. However successful they may be, these works
do not address the dependency~\cite{variability} of scheduling metrics on the
workload. Indeed, these studies often report \textit{post-hoc} performance.

The dynP~\cite{streit_selftuning_2002} scheduler does propose a systematic
method to tuning these queues, although it requires simulated scheduling runs
at decision time and therefore does not enjoy the natural speed of execution
of EASY.

There is a recent focus on leveraging the high amount of data available in
large scale computing systems in order to improve their behavior. Some works
use collaborative filtering to colocate tasks in clouds by estimating
application interference~\cite{7516031}. Others~\cite{fmodeling} are closer to
the application level and use binary classification to distinguish benign
memory faults from application errors in order to execute recovery algorithms.

A few works~\cite{Tsafrir_easypp_2005,learningruntimes} use this method in HPC,
hoping that better job runtime estimation should improve scheduling
~\cite{chiang_impact_2002}. Some algorithms estimate runtime
distributions models and choose jobs using probabilistic integration
procedures~\cite{Nissimov2008}.

However, these works do not address the duality between the cumulative and
maximal scheduling costs, as previously outlined in Figure~\ref{tab:maxbsldlit} for
the EASY++~\cite{Tsafrir_easypp_2005} and Easy-Eloss~\cite{learningruntimes}
heuristics.

While these works try to estimate uncertain parameters, this paper takes a more
basic approach and directly tries to learn a good scheduling policy from a
simple policy space.

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.5]{\paretobasicpdf}

  \caption{Performance of the 49 Backfilling heuristics generated by using the
    7 possible policies as queue and backfilling order, averaged over 2000
    resampled weeks from the \textit{training} KTH-SP2 trace, in terms of both
    maximum and average waiting time of the jobs. All values are normalized by
  that of \textbf{FCFS} reservations with \textbf{FCFS} backfilling and the
error bars correspond to the standard deviation.}

  \label{fig:paretobasic}
\end{figure*}

\section{Experimental Protocol}
\label{sec:expe}

This section motivates the statistical approach used to measure performance,
describes the simulation method used and explains the train/test protocol used
to evaluate the generalization ability of heuristics. Finally, it describes the
workload logs and concludes on the results of the experimental campaign.

\subsection{Statistical approach}
\label{sub:sa}

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.5]{\variabilitypdf}
  \includegraphics[scale=0.5]{\variabilitynormpdf}

  \caption{AvgWait cost of using the 7 main queue policies with \textbf{FCFS}
    backfilling for 100 generated weeks. First, in absolute value, and then
    normalized on the cost of using the \textbf{FCFS} policy to order the
  reservation queue (in yellow). }

  \label{fig:variability}
\end{figure*}

The experimental approach used in this paper is statistical in nature.
Figure~\ref{fig:variability} shows how the AvgWait costs of the 7 reservation
policies used along with FCFS backfilling evolves troughout the first 30 weeks
of the original KTH-SP2 trace. The
variability~\cite{variability,frachtenberg2005pitfalls} of cost metrics and
their sensitivity to small changes in the workload logs~\cite{flurries} have been
thouroughly studied in the literature. Our approach to measuring performance
without reporting noise from workload flurries~\cite{flurries} is to aggregate
the cost metric on a large number of generated logs. The approach of this
paper follows in part \cite{feitresampling} and design a trace resampler in
order to generate week-long workload logs from an original dataset. The
resampling technique used is simplistic in nature: for each system user, a
random week of job submissions from the original trace is used.  This approach
is combinatorially sufficient to generate a quasi-infinity of logs while
preserving the natural dependence of the workload on the weekly period and the
variability in load. On the downside, the seasonal effect and the dependence
between users is lost.  Moreover, there is no user model or other feedback loop
in the simulations. In all experiments, the performance of every policy is
evaluated by averaging the cost values over 2000 generated weeks.

\subsection{Simulation method and testbed}

While high quality simulators like SimGrid~\cite{casanova:hal-01017319} are
available in practice, this paper focuses on backfilling behavior and does not
use such advanced models. This is motivated by the fact that one needs to use a
high-performance approach to simulation in order to perform the high number of
scheduling runs necessary for this study (The total number of week-long
simulations in this paper reaches X). Therefore, experiments are run with a
specially designed backfilling scheduler that trades faithfulness for speed by
not including locality. Using this simulator, a week of EASY-fcfs backfilling
can be replayed in under 10'', I/O operations (reading and writing a swf file)
included. All simulations are performed on a Dell PowerEdge T630 machine with
2x Intel(R) Xeon(R) CPU E5-2697 v3 @ 2.60GHz/14 cores (28 cores/node), and 260
GB of RAM. We use a minimalistic aprroach to reproducible
research~\cite{stodden2014implementing} and provide a snapshot of the work that
includes a build system that runs the experiments using the
\textit{zymake}~\cite{breck2008zymake} minimalistic workflow system.  The
archive includes our simulator and a nix~\cite{dolstra2004imposing} file that
describes the dependencies.

\subsection{Generalization protocol}
\label{sub:gp}

The goal of this paper to study how the performance of different heuristics
generalize. The protocol is as follows: The initial workload is split
at temporal midpoint in two parts, the \textit{training} and \textit{testing}
logs. Each of these are used to resample weeks. Morally, this corresponds to
the situation where a systems administrator having retained usage logs from a
HPC center must choose a scheduling policy for the next period using
simulation. For each HPC log from the Parallel Workload archive used in the
experiment, this process results in two databases of 2000 weeks each.

\section{Problem Setting}
\label{sec:pbset}

This section describes a generic platform model used in this paper, presents the EASY
heuristic and defines two scheduling cost metrics to be minimized. Finally, it
motivates and introduces the problem statement of this paper.

\subsection{System Description}
\label{sub:sysdesc}

The problem adressed in this paper is the one faced by Resource and Job
Management Systems (RJMS) such as SLURM\cite{SLURMdocSCHED}, PBS\cite{PBSdoc}
and OAR\cite{capit2005batch}.

The crucial part of these software is the scheduling algorithm that determines
where and when the submitted jobs are executed. The process is as follows: Jobs
are submitted by end-users into queues until the scheduler selects one of them
for running. Each job has an provided bound on the execution time and some
resource requirements (number and type of processing units). Then, the RJMS
drives the search for the resources required to execute this job.  Finally, the
tasks of the job are assigned to the chosen nodes.

In the classical case, these software need to execute a set of concurrent
parallel jobs with rigid (known and fixed) resource requirements on a HPC
platform represented by a pool of $m$ identical resources. This is an online
scheduling problem as jobs are submitted over time.

A job $j$ has the following characteristics:

\begin{itemize}
  \item Submission date $r_j$ (sometimes called \textit{release date})

  \item Resource requirement $q_j$ (processor count)

  \item Actual running time  $p_j$ (sometimes called \textit{processing time})

  \item Requested running time $\widetilde{p_j}$ (sometimes called \textit{walltime}), which is an upper bound on $p_j$
\end{itemize}

The resource requirement $q_j$ of a job is known when the job is submitted at
time $r_j$, while the requested running time $\widetilde{p_j}$ is given by the
user as an estimate. Its actual value $p_j$ is only known \textit{a posteriori}
when the job really completes.  Moreover, the user has incentive to
overshoot the actual value, since his job could get ``killed'' if it
surpasses the provided value.

\subsection{EASY Backfilling}
\label{sub:notation}

The selection of the job to run is performed according to a
scheduling policy that establishes the order in which the jobs are executed.
EASY-Backfilling is the most widely used policy due to its easy and robust
implementation and known benefits such as high system
utilization~\cite{Feitelson}. This strategy has no worst case guarantee beyond
the absence of starvation (i.e. every job will be scheduled at some moment).

The EASY heuristic uses a job queue to perform job
\textit{starting/reservation} and job \textit{backfilling}. These queues can be
dissociated and the heuristic can be parametrized via both a reservation policy
and a backfilling policy. This is typically done by ordering both queues in an
identical manner using job attributes. In the following, we denote by
EASY-$P_R$-$P_B$ the scheduling policy that starts jobs and does the
reservation according to policy $P_R$ and backfills according to policy $P_B$.
For the sake of completeness, Algorithm~\ref{alg:EASY} describes the
EASY-$P_R$-$P_B$ heuristic.

\begin{algorithm}[H]
  \caption{EASY-$P_R$-$P_B$ policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE Queue $Q$ of waiting jobs.
    \ENSURE None (calls to $Start()$)
    \\ \textit{Starting jobs in the $P_R$ order}
    \STATE Sort $Q$ according to $P_R$
    \FOR {job $j$}
    \STATE Pop $j$ from Q
    \IF {$j$ can be started given current system use.}
    \STATE $Start(j)$
    \ELSE
    \STATE Reserve $j$ at the earliest
    time possible according to the estimated running times
    of the currently running jobs.
    \\ \textit{Backfill jobs in the $P_B$ order}
    \STATE $L \leftarrow Q$
    \STATE Sort $L$ according to $P_B$
    \FOR {job $j'$ in $L$}
    \IF {$j'$ can be started without delaying the reservation on $j$.}
    \STATE $Start(j')$
    \ENDIF
    \ENDFOR
    \STATE \textbf{break}
    \ENDIF
    \ENDFOR
  \end{algorithmic}
  \label{alg:EASY}
\end{algorithm}

This paper makes use of 7 simple queue reordering policies, some of which are
often studied in the literature:

\begin{itemize}
  \item FCFS: First-Come First-Serve, the widely used default~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  \item LPF: Longest estimated Processing time $\widetilde{p_{j}}$ First.
  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
  \item LQF: Largest resource requirement $q_j$ First.
  \item SQF: Smallest resource requirement $q_j$ First.

  \item EXP: Largest Expansion Factor~\cite{bfchar} First, where the expansion
    factor is, by writing $start_j$ the starting time of a job:

  \begin{equation} \frac{start_j - r_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
\end{itemize}

\subsection{Scheduling Costs}
\label{sub:scheduling_objectives}

A system administrator may use one or multiple cost metric(s). Our study of
scheduling performance relies on two of the more frequently used:

\begin{itemize}
  \item The Waiting Time of a job:
    \begin{equation}
      \textbf{Wait}_j =  start_j-r_j
    \end{equation}
  \item The Average Bounded Slowdown~\cite{frachtenberg2005pitfalls,feitelson1998metrics} , a more refined metric:
    \begin{equation}
      \textbf{Bsld}_j = \max \left(  1,\frac{\textbf{Wait}_j + p_j}{\max \left( \tau,p_j \right) } \right)
    \end{equation}
    where $\tau$ is set as is usual~\cite{feitelson1998metrics} to 10 seconds.
\end{itemize}

Both of these cost metrics are usually considered in their \textit{cumulative}
version, which means that one seeks to minimize the Average Bounded Slowdown
(\textbf{AvgBsld}) or the (\textbf{AvgWait}). In the following, we will also
use the maximal version of these costs, which we denote by \textbf{MaxWait} and
\textbf{MaxBsld}, a.k.a the maximal value of the waiting time and bounded
slowdown on all the jobs from a scheduling run.

\subsection{Problem Description}
\label{sec:minmax}

There are in the author's view two main difficulties when efficiently tuning the
EASY heuristic. Each of these two issues are illustrated here by a small
scheduling experiment.

First, the relative performance of EASY policies is sensitive to
context~\cite{variability,bfchar}. Table~\ref{tab:context} illustrates this
effect by comparing the AvgBsld of two different queue ordering policies on
logs two different workloads. This hints to the fact that there is no "one size
fits all" choice of reservation and backfilling policies.  In such a situation,
tuning of the EASY heuristic must be done locally for each HPC system. This can
be done via simulation, in which case one is interested in how the simulation
results \textit{generalize} to the future.

\input{\filetablec}

\begin{table*}[ht!]
  \centering
  \ra{1.3}

  \caption{\textbf{AvgBsld} and \textbf{MaxBsld} of two "statistical" EASY
  Policies versus EASY-FCFS-FCFS. This is a single-run experiment on the
initial workload logs described in Section\ref{sec:expe}.}

  \label{tab:maxbsldlit}
  \begin{tabular}{lrrrrrr}
    \hline
                                      & KTH-SP2 & CEA-CURIE & CTC-SP2 & SDSC-BLUE & SDSC-SP2 \\
    \hline
    \textbf{AvgBsld} cost :           &         &           &         &           &          \\
    \hline
    EASY-FCFS                         & 92.5    & 202.0     & 49.5    & 36.5      & 87.8     \\
    EASY-Eloss\cite{learningruntimes} & 51.4    & 27.8      & 20.5    & 12.9      & 74.9     \\
    EASY++\cite{Tsafrir_easypp_2005}  & 63.5    & 193.0     & 85.7    & 20.9      & 79.3     \\
    \hline
    \textbf{MaxBsld} cost :           &         &           &         &           &          \\
    \hline
    EASY-FCFS                         & 14805   & 9971      & 8137    & 3395      & 5981     \\
    EASY-Eloss                        & 39929   & 511197    & 52933   & 21750     & 190656   \\
    EASY++                            & 30250   & 286588    & 66691   & 15530     & 13681    \\
    \hline
  \end{tabular}
\end{table*}

Second, starvation may appear when changing the EASY queue policy away
from FCFS. This issue concerns the method used to measure the objective. Most
systems use a variant of the FCFS-FCFS policy, where the FCFS policy is used
both for job starting and reservation and for backfilling jobs. The main
advantage of this choice is that it controls the \textit{starvation risk} by
greedily minimizing the maximum values of the job waiting times.
Indeed, a job might be indefinitely delayed when not starting jobs in the FCFS
order. Table~\ref{tab:maxbsldlit} illustrates how this effect appears in
several related works \cite{Tsafrir_easypp_2005,learningruntimes} that optimize
the average cost by removing the FCFS constraint.

In this paper, we ask the following: \textbf{How to leverage workload data in
order to improve cumulative cost metrics while controlling their maximum
values?}

In order to answer this question, we investigate the use of simulation to tune
EASY Backfilling by reordering its two queues.  The first conclusion is that
reordering the job starting and reservation queue is more beneficial than
simply reordering the backfilling queue. However, this introduces a risk on the
maximum values of the objective, which we control by hybridizing FCFS and the
reordering policy via a thresholding mechanism.  Finally, we show that the
experimental performance of the thresholded heuristics generalizes well.

\section{Maximum, Average cost and Experimental Generalization}
\label{sec:mmax}

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.5]{\paretopdf}

  \caption{Performance of the 49 Backfilling heuristics generated by using the
    7 possible policies as queue and backfilling order, averaged over 300
    resampled weeks, in terms of both maximum and average waiting time of the
    jobs. The transparent triangular points correspond to the \textit{training}
    trace and the hard round points to the \textit{testing} trace. All
    \textit{training} and \textit{testing} values are normalized with respect
    to the \textbf{EASY} reservations with \textbf{EASY} backfilling.}

  \label{fig:paretogen}
\end{figure*}

This section presents a small experimental campaign that uses the KTH-SP2 trace
in order to illustrate the contradictory effect of averaged and maximum cost.
In order to report results accurately, all results presented from this point
use the experimental protocol described in Section~\ref{sec:expe}.  The reader
is invited to skim on subsections~\ref{sub:sa} and~\ref{sub:gp} before going
further as it will improve the understanding of the following paragraphs.
Figure~\ref{fig:paretobasic} shows how the cumulative and maximal costs of all
policy couples (one policy for the reservation order and one for the
backfilling order) evolve when various policies are used for reservation and
backfilling. There are two main observations.

First, It seems possible to improve the AvgWait on this machine as far
as to divide it by two compared to the EASY FCFS-FCFS baseline. However, such
AvgWait improvements seem to entail an increase in MaxWait
cost. Expectedly, the EASY-FCFS-FCFS heuristic has a good MaxWait behavior.

Second, there seem to be regularities in the behavior of the performance: The
main factor seems to be the reservation policy, while the importance of the
backfilling policy varies depending on the reservation policy used. It appears
that some policies such as SQF do not lead to many backfilling decisions, while
others like LQF encourage frequent backfilling. Additionaly, there are some
backfilling policies, such as SPF and ExpFact that seem to consistently
outperform the others.

The traditional approach to multi-criteria optimization is to choose a point
using the pareto frontier. However, one must keep in mind that these results
are obtained \textit{in hindsight} on the \textit{training} trace and do not
necessarily generalize well to the future.

\subsection{Generalization of average and maximum cost}

This section discusses the generalization of cost metrics in terms of average
and maximum values. The point is illustrated as in the previous section by
using the KTH-SP2 trace and the experimental protocol described in section
\ref{sec:expe}.

Figure~\ref{fig:paretogen} shows the evolution of the AvgWait and MaxWait costs
obtained by the 49 heuristics between \textit{training} and \textit{testing}
logs. Training (transparent) and testing (hard) values are connected by
dashed lines. The relative performance of queue policies seems to be somehow
stable. However, there is a displacement of values due to changes in the
characteristics of the second part of the original trace\footnote{An apparent
  difference between training and testing parts of the KTH-SP2 trace is the
decrease in load.}.

Since the ordinal performances seem roughly conserved, one can hope to
generalize with respect to the cumulative cost. However, the variability in
values indicates that one can not use classic pareto curve methods to
generalize from training to testing logs in regions of low maximum cost. The
next section therefore introduces a thresholding mechanism in order to control
the maximum cost.

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.5]{\paretothpdf}

  \caption{Performance of the 49 Backfilling heuristics generated by using the
    7 possible policies as queue and backfilling order, averaged over 300
    resampled weeks from the \textit{training} trace, in terms of both maximum
    and average waiting time of the jobs. The transparent triangular points
    correspond to the policies without threshold and the hard round points to the
    thresholded version. All values are normalized with respect to the
    \textbf{EASY} reservations with \textbf{EASY} backfilling.}

\end{figure*}

\section{Thresholding}

This section introduces control over maximal costs using a simple thresholding
mechanism.

\subsection{Thresholding and risk}
\label{sec:th}

The future costs $\text{Wait}_j$ and $\text{Bsld}_j$ of a waiting job $j$
are lower-bounded at any time $t$ by the "waiting time so far" $t-r_j$. A
simple way to introduce robustness into the heuristic is therefore to force
jobs with unusually high values of $t-r_j$ ahead of the reservation queue. One
way to do this is to introduce a threshold parameter $T$ and push jobs with
$t-r_j>T$ immediately ahead of the reservation queue after the reservation
queue sorting step (line 1 of Algorithm~\ref{alg:EASY}). If more than one job
is in this situation, these jobs are ordered by submission time $r_j$ at the
head of the queue. Figure~\ref{fig:paretoth} illustrates the effect this has on
the 49 possible heuristics on the KTH-SP2 system with $T = 20$ hours. The
MaxWait cost is greatly reduced, while all AvgWait values are
(perhaps expectedly) moved torwards EASY-FCFS-FCFS. Note that while the
variance in terms of MaxWait is greatly reduced, the variance in terms
of AvgWait is not reduced as much. This is the effect this paper
leverages. The next subsection gives a glimpse of the behavior of
generalization in this framework.

\subsection{Thresholding and generalization}
\label{sec:genth}

The final step is to study how the performance thresholded queue policies
generalizes. Figure~\ref{fig:parallelkth} shows how the performance of all
various queue and backfilling policies evolve from \textit{training} to
\textit{testing} logs when the threshold is set to an example value of 20
hours. While values do change from \textit{training} to \textit{testing}
logs, the relative order of policies seems to be roughly conserved. This
leaves hope for generalization.

\begin{figure}[ht!]
  \centering
  \includegraphics[scale=0.5]{\kthparallelpdf}

  \caption{AvgBsld generalization of thresholded policies obtained by using a
    threshold value of 20 hours. Note that each plot has a different vertical
    axis. The reported AvgBsld and MaxBsld values are averaged over 300
    resampled weeks from the training or testing original logs, normalized over
    the cost of EASY-FCFS-FCFS. The "Worst MaxBsld" value is the highest (300
    week averaged) maximum average bounded slowdown increase percentage
    reported for any of the 49 heuristics. The values reported under Learned
    Performance are the (300 week averaged) decrease percentages obtained by the
    best \textbf{Training} heuristinc on the \textbf{Testing} logs.}

  \label{fig:parallelkth}
\end{figure}

The next section will outline the experimental setup and present the results
from a comprehensive experimental campaign that validates this effect.

\section{Experimental Validation}

\subsection{Workload logs}

Table~\ref{tab:logs} outlines the six workload logs from the Parallel
Workloads Archive~\cite{Feitelson20142967} used in the experiments. The logs
are subjected to pre-filtering. The filtering step excludes jobs with
$\widetilde{p_j} < p_j $ and jobs whose "requested\_cores" and "allocated\_cores"
fields exceed the size of the machine.

\begin{table}[h]
  \vspace{-0.5cm}
  \centering
  \ra{1.3}
  \caption{Workload logs used in the simulations.}
  \label{tab:logs}
  \begin{tabular}{@{}lrrrr@{}}
    \hline
    Name      & Year & \# CPUs      & \# Jobs       & Duration\\
    \hline
    KTH-SP2     & 1996 & 100       & 28k        & 11 Months\\
    CTC-SP2     & 1996 & 338       & 77k        & 11 Months\\
    SDSC-SP2    & 2000 & 128       & 59k        & 24 Months\\
    SDSC-BLUE   & 2003 & 1,152     & 243k       & 32 Months\\
    Curie       & 2012 & 80,640    & 312k       & 3  Months\\
    Metacentrum & 2013 & 3,356     & 495k       & 6  Months\\
    \hline
  \end{tabular}
  \vspace{-0.5cm}
\end{table}

\subsection{Results}

Figure~\ref{fig:threshold} (resp: Figure~\ref{fig:thresholdb}) summarizes the
behavior of Wait (resp: Bsld) generalization and risk with
respect to the value of the threshold $T$. There is a fortunate effect in that
the values from the lower parts of the graphs (the \textit{AvgWait} cost) seem
to decrease faster than values from the upper part (the \textit{MaxWait} cost),
which increase linearly with $T$.

By using an aggressive appproach (no threshold), the AvgWait can be reduced
until 80\% to 65\% compared to the EASY-FCFS-FCFS baseline. However, in that
case the values of the MaxWait can jump as high as 250\% that of the baseline.
The same phenomenon happens with the aggressive approach on the AvgBsld, where
we can attain a reduction of the value until 80\% to 50\%, but this entails an
increase of MaxBsld values up to 120\% times that of the baseline.

By using a conservative approach (thresholding at 20 hours), the AvgWait can be
reduced until 90\% to 70\% in expectation, while keeping the MaxWait increase
under 175\% of the baseline in all cases. We observe again the same phenomenon using the
bounded slowdown, where the AvgBsld can be reduced until 83\% to 60\% while
keeping the MaxBsld under 105\% of the baseline in all cases.

Figure~\ref{fig:par} (resp. Figure~\ref{fig:par}) shows how the AvgWait (resp.
AvgBsld) cost of the 49 combination of queue and backfilling policies evolve
from training to testing logs when we use this conservative threshold of 20
hours, with a higher sample size that was not premitted by the previous
experiment. This confirms the previous values and gives visual insight into the
stability of the performance.

Figure~\ref{fig:thresholdb} shows the same behavior for the AvgBsld
values.

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.4]{\kthprogress}
  \includegraphics[scale=0.4]{\sdscbprogress}
  \includegraphics[scale=0.4]{\sdscsprogress}
  \includegraphics[scale=0.4]{\ctcprogress}

  \caption{AvgWait and MaxWait generalization of thresholded policies as
  affected by the queue threshold. The Value reported as "train" is that of the
least costly heuristic among the 49 possible policy parametrizations averaged
on the \textit{training} logs. The Value reported as "test" is the averaged
cost of the same heuristic on the \textit{testing} logs.}

  \label{fig:threshold}
\end{figure*}

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.4]{\kthprogressb}
  \includegraphics[scale=0.4]{\sdscbprogressb}
  \includegraphics[scale=0.4]{\sdscsprogressb}
  \includegraphics[scale=0.4]{\ctcprogressb}

  \caption{AvgBsld and MaxBsld generalization of thresholded policies as
  affected by the queue threshold. The Value reported as "train" is that of the
least costly heuristic among the 49 possible policy parametrizations averaged
on the \textit{training} logs. The Value reported as "test" is the averaged
cost of the same heuristic on the \textit{testing} logs.}

  \label{fig:thresholdb}
\end{figure*}

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.5]{\sdspbparallelpdf}
  \includegraphics[scale=0.5]{\sdspsparallelpdf}
  \includegraphics[scale=0.5]{\ctcparallelpdf}
  \includegraphics[scale=0.5]{\kthparallelpdf}

  \caption{AvgWait generalization of thresholded policies obtained by using a
  threshold value of 20 hours. Note that each plot has a different vertical y axis. The
reported AvgWait and MaxWait values are averaged over 300 resampled weeks from
the training or testing original logs, normalized over the cost of
EASY-FCFS-FCFS. The "Worst MaxWait" value is the highest (300 week averaged)
maximum average bounded slowdown increase percentage reported for any of the 49
heuristics. The values reported under Learned Performance are the (300 week
averaged) decrease percentages obtained by the best \textbf{Training}
heuristinc on the \textbf{Testing} logs.}

  \label{fig:par}
\end{figure*}

\begin{figure*}[ht!]
  \centering
  \includegraphics[scale=0.5]{\sdspbparallelbpdf}
  \includegraphics[scale=0.5]{\sdspsparallelbpdf}
  \includegraphics[scale=0.5]{\ctcparallelbpdf}
  \includegraphics[scale=0.5]{\kthparallelbpdf}

  \caption{AvgBsld generalization of thresholded policies obtained by using a
    threshold value of 20 hours. Note that each plot has a different vertical y
    axis. The reported AvgBsld and MaxBsld values are averaged over 300
    resampled weeks from the training or testing original logs, normalized over
    the cost of EASY-FCFS-FCFS. The "Worst MaxBsld" value is the highest (300
    week averaged) maximum average bounded slowdown increase percentage
    reported for any of the 49 heuristics. The values reported under Learned
    Performance are the (300 week averaged) decrease percentages obtained by the
    best \textbf{Training} heuristinc on the \textbf{Testing} logs.}

  \label{fig:parb}
\end{figure*}

\section{Conclusion}

This work leverages the fact that the performance of scheduling heuristics
depend on the workload profile of the system. More precisely, we investigate
the use of simulation to tune EASY Backfilling by reordering its two queues.
The first conclusion is that reordering the job starting and reservation queue
is more beneficial than simply reordering the backfilling queue. However, this
introduces a risk on the maximum values of the objective, which we control by
hybridizing FCFS and the reordering policy via a thresholding mechanism.
Finally, we show that the experimental performance of the thresholded
heuristics generalizes well. Therefore, this framework allows a system
administrator to tune EASY using a simulator. Moreover, the attitude torwards
risk in maximum values can be specified via the threshold value. With a low
threshold value, the increase in maximal cost is small but the learned policy
does not take many risks. With a high threshold value, it's possible to gain
more, but this comes with an increase in the maximal cost. Two questions
concerning the learning of EASY policies arise from this work.

First, the stability of other EASY heuristic classes remains unknown. The
simple class of composed of 7 reservation policies and 7 backfilling policies
(cardinality 49) can generalize using thresholding. It is natural to ask
whether it could be possible to learn using a larger class of heuristics.  One
could for instance consider the class of mixed policies that choose a job based
on a linear combination of the 7 criteria. A more ambitious endeavor is to ask
whether it is possible to learn a contextual job ranking
model~\cite{joachims2002optimizing} that performs well.

\textit{ a retirer?
Second, the performance improvement achievable by the best reordering of
backfilling and reservation queues is unknown. Answering this hard optimization
task would give a lower bound on the achievable costs.
}

%\section{Annex}

%This section presents for the sake of completeness the equivalent of
%Figure~\ref{pareto} for the remaining workload logs.

\section{Acknowledgements}

Authors are listed in alphabetical order. We warmly thank Frederic Wagner and
Eric Gaussier for discussions as well as Pierre Neyron and Bruno Breznik for
their invaluable help with experiments. We gracefully thank the contributors of
the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2), Travis Earheart
and Nancy Wilkins-Diehr (SDSC Blue), Lars Malinowsky (KTH SP2), Dan Dwyer and
Steve Hotovy (CTC SP2), Joseph Emeras (CEA Curie), and of course Dror
Feitelson.  The Metacentrum workload log was graciously provided by the Czech
National Grid Infrastructure MetaCentrum. This work has been partially
supported by the LabEx PERSYVAL-Lab (ANR-11-LABX-0025-01) funded by the French
program Investissement d'avenir. Experiments presented in this paper were
carried out using the Digitalis platform\footnote{http://digitalis.imag.fr} of
the Grid'5000 testbed. Grid'5000 is supported by a scientific interest group
hosted by Inria and including CNRS, RENATER and several Universities as well as
other organizations\footnote{https://www.grid5000.fr}. Access to the
experimental machine(s) used in this paper was gracefully granted by research
teams from LIG\footnote{http://www.liglab.fr} and
Inria\footnote{http://www.inria.fr}.

\bibliographystyle{o.splncs03}
\bibliography{o.bibliogr}

\end{document}

### 2017-03-28 misc/extract.sh

@@ -1,0 +1,7 @@
+#!/usr/bin/env bash
+
+rm -rf pdfs
+mkdir -p pdfs
+cp o/*/*.txt pdfs
+cp o/*/*.pdf pdfs
+cp o/*/*.png pdfs

### 2017-03-28 misc/extract_maxprocs.py

@@ -1,0 +1,30 @@
+#!/usr/bin/env python
+# encoding: utf-8
+
+'''
+    extract maxprocs
+
+Usage:
+    extract_maxprocs.py <swf_file> [-h]
+
+Options:
+    -o <option>    some option
+    -h --help      show this help message and exit.
+'''
+from docopt import docopt
+#retrieving arguments
+arguments = docopt(__doc__, version='1.0.0rc2')
+
+with open(arguments['<swf_file>']) as input_file:
+    num_processors=None
+    for line in input_file:
+        if(line.lstrip().startswith(';')):
+            if(line.lstrip().startswith('; MaxProcs:')):
+                num_processors = int(line.strip()[11:])
+                break
+            else:
+                continue
+        else:
+            break
+
+print num_processors

### 2017-03-28 misc/init_queue.sh

@@ -1,0 +1,4 @@
+#!/usr/bin/env bash
+
+ts -K
+TS_SLOTS=3 ts

### 2017-03-28 misc/knife.sh

@@ -1,0 +1,5 @@
+#!/bin/sh
+
+lines=$(wc -l $1 |cut -f1 -d' ')
+
+head $1 -n $((lines/3)) > $2

### 2017-03-28 misc/passthrough.sh

@@ -1,0 +1,51 @@
+#!/usr/bin/env bash
+
+csv=$1
+mcts=$2
+h1=$3
+h2=$4
+out=$5
+
+wmcts=$(cat $mcts| cut -d' ' -f1)
+wh1=$(cat $h1| cut -d' ' -f1)
+wh2=$(cat $h2| cut -d' ' -f1)
+wcl=$(wc -l $csv|cut -d' ' -f1)
+
+if [ "$wcl" -lt "3" ]
+then
+  touch $out
+  exit 0
+fi
+
+btest=$( echo "$wmcts < 10" | bc)
+
+
+if [ $btest -eq 1 ]
+then
+  touch $out
+  exit 0
+fi
+
+
+weight=$(echo "$wmcts $wh1 $wh2 $wcl" | awk '{print (($2+$3)/2 - $1 )/$4 }' )
+
+wmax=$(awk -F , 'max < $1 { max = $1 } END { print max }' $csv)
+
+wmin=$(awk -F , 'NR == 1 || m > $1 { m = $1 } END { print m }' $csv)
+echo $wmcts $wh1 $wh2 $wcl
+
+btest=$( echo "$wmax - $wmin < 0.01" | bc)
+
+echo $wmcts $wh1 $wh2 $wcl
+
+if [ "$btest" -eq 1 ]
+then
+  awkstr=$(echo "\$1=\$1 * $weight")
+else
+  awkstr=$(echo "\$1=(\$1-$wmin)/(($wmax-$wmin)*$weight)")
+fi
+
+
+echo $wmcts $wh1 $wh2 $wcl
+
+cat $csv | awk -F , -v OFS=, $awkstr > $out

### 2017-03-28 misc/special_treatment.py

@@ -1,0 +1,40 @@
+#!/usr/bin/python3
+
+import sys
+from shutil import copyfile
+
+if "CEA-Curie-2011-2.1-cln.swf" in sys.argv[1]:
+  copyfile(sys.argv[1],sys.argv[2])
+elif "CTC-SP2-1996-3.1-cln.swf" in sys.argv[1]:
+# keep the last part of the file
+  copyfile(sys.argv[1],sys.argv[2])
+elif "KTH-SP2-1996-2.1-cln.swf" in sys.argv[1]:
+  copyfile(sys.argv[1],sys.argv[2])
+elif "Sandia-Ross-2001-1.1-cln.swf" in sys.argv[1]:
+  # keep the last part of the file
+  # we keep only the part that seems to have 960procs (30 cabinet of 32 procs)
+  #it starts around 28500000
+  import pandas as pd
+
+  df = pd.read_csv(sys.argv[1], sep=" ", skipinitialspace=True, comment=";", header=None, names=['job_id','submit_time','wait_time','run_time','proc_alloc','cpu_time_used','mem_used','proc_req','time_req','mem_req','status','user_id','group_id','exec_id','queue_id','partition_id','previous_job_id','think_time'])
+
+  df = df[df.submit_time >= 28500000]
+  print("Deleted jobs that are bigger than the machine:", len( df[df.proc_req > 960]))
+  df = df[df.proc_req <= 960]
+
+  f = open(sys.argv[2], 'w')
+
+  f.write(';Version: SWF v2\n')
+  f.write(';MaxNodes: 960\n')
+  f.write('; MaxProcs: 960\n')
+  f.write(';Note: This is a cutted version of '+str(sys.argv[1])+' (cutted at 28500000)\n')
+  df.to_csv(f, sep=" ", header=None, index=False)
+  f.close()
+
+elif "SDSC-BLUE-2000-4.2-cln.swf" in sys.argv[1]:
+  copyfile(sys.argv[1],sys.argv[2])
+elif "SDSC-SP2-1998-4.2-cln.swf" in sys.argv[1]:
+  copyfile(sys.argv[1],sys.argv[2])
+else:
+  print(sys.argv[1],sys.argv[2])
+  # copyfile(sys.argv[1],sys.argv[2])

### 2017-03-28 misc/split.sh

@@ -1,0 +1,13 @@
+#!/usr/bin/env bash
+file=$1
+trf=$2
+tef=$3
+
+l=$(wc -l $file |cut -d' ' -f1)
+h=$((3*l/5))
+t=$((2*l/5))
+
+head -n $h $file > $trf
+tail -n $t $file > $tef
+
+echo "successfully splat the files."

### 2017-03-28 misc/split_swf.sh

@@ -1,0 +1,24 @@
+#!/usr/bin/env bash
+type=$1
+in_file=$2
+out_file=$3
+
+l=$(cat ${in_file} | grep -v ";" | wc -l | cut -d' ' -f1)
+h=$((l/2))
+
+echo "length: $h"
+
+touch "${out_file}"
+
+cat ${in_file} | grep ";" > "${out_file}"
+
+if [ "$type" == "train" ]; then
+  echo "getting training"
+  cat ${in_file} | grep -v ";" | head -n $h >> "${out_file}"
+fi
+
+if [ "$type" == "test" ]; then
+  echo "getting testing set"
+  cat ${in_file} | grep -v ";" | tail -n $h >> "${out_file}"
+fi
+

### 2017-03-28 misc/store_annex.sh

@@ -1,0 +1,9 @@
+#!/usr/bin/env bash
+
+mkdir ~/annex/experiments/$1
+cp -r o pdfs ~/annex/experiments/$1
+git log | head -n 1 > ~/annex/experiments/$1/metadata
+cd ~/annex/experiments
+echo "adding"
+git annex add $1
+git commit -m "added experiment: $1" -a

### 2017-03-28 misc/submit.sh

@@ -1,0 +1,4 @@
+#!/usr/bin/env bash
+
+jid=$(ts bash -c "$(cat $1)")
+ts -w $jid

### 2017-03-28 misc/visu/accuracy.R

@@ -1,0 +1,26 @@
+#!/usr/bin/Rscript
+
+library(docopt)
+'usage: accuracy.R <input> ... [--mean=<output1>] [--max=<output2>]
+
+options:
+<input>                   The input data.
+' -> doc
+
+args<- docopt(doc)
+
+maxscore=0
+best=NULL
+for (filename in args$input) {
+  d=read.table(filename)
+  pos=length(which(d$V1>0))/length(d$V1)
+  neg=length(which(d$V1<0))/length(d$V1)
+  score=pos/(neg+pos)
+  if (score>maxscore  ) {
+    best=filename
+    maxscore=score 
+  }
+  print(score)
+}
+print(maxscore)
+print(best)

### 2017-03-28 misc/visu/accuracy_plot.R

@@ -1,0 +1,40 @@
+#!/usr/bin/Rscript
+
+library(docopt)
+library(TTR)
+
+'usage: accuracy.R <input> ... [--out=<out>]
+
+options:
+<input>                   The input data.
+--out=<out>               The output pdf.
+' -> doc
+
+args<- docopt(doc)
+
+theme_bw_tuned<-function()
+{
+	return(theme_bw() +theme(
+		plot.title = element_text(face="bold", size=10),
+		axis.title.x = element_text(face="bold", size=10),
+		axis.title.y = element_text(face="bold", size=10, angle=90),
+		axis.text.x = element_text(size=10),
+		axis.text.y = element_text(size=10),
+		panel.grid.minor = element_blank(),
+		legend.key = element_rect(colour="white"))
+         )
+}
+
+print(args)
+maxscore=0
+best=NULL
+for (filename in args$input) {
+  d=read.table(filename)
+
+  #ema=EMA(d,n=5000)
+  ema=EMA(d,n=5000)
+  pdf(file=paste(filename,args$out,sep="-"),width=10,height=7)
+  #plot(cumsum(d$V1))
+  plot(ema)
+  dev.off()
+}

### 2017-03-28 misc/visu/average_cumobj.R

@@ -1,0 +1,68 @@
+#!/usr/bin/env Rscript
+
+'usage: tool.R <input_files> ...  [-o <output>]
+tool.R -h | --help
+
+options:
+<input_files>                   The input data.
+-o <output>                     Output file.
+-h , --help                     Show this screen.
+' -> doc
+
+library(docopt)
+args<- docopt(doc)
+
+#library('TTR')
+#library('gridExtra')
+library('ggplot2')
+#library('reshape2')
+#library('plyr')
+library('dplyr')
+
+theme_bw_tuned<-function()
+{
+  return(theme_bw() +theme(
+                           plot.title = element_text(face="bold", size=10),
+                           axis.title.x = element_text(face="bold", size=10),
+                           axis.title.y = element_text(face="bold", size=10, angle=90),
+                           axis.text.x = element_text(size=10),
+                           axis.text.y = element_text(size=10),
+                           panel.grid.minor = element_blank(),
+                           legend.key = element_rect(colour="white"))
+  )
+}
+
+swf_read <- function(f)
+{
+  df <- read.table(f,comment.char=';')
+  names(df) <- c('job_id','submit_time','wait_time','run_time','proc_alloc','cpu_time_used','mem_used','proc_req','time_req','mem_req','status','user_id','group_id','exec_id','queue_id','partition_id','previous_job_id','think_time')
+  return(df)
+}
+
+timestamp_to_date <- function(timestamp){
+  return(as.POSIXct(timestamp, origin="1970-01-01 01:00.00", tz="Europe/Paris"))
+}
+
+d=swf_read(args$input_files[1])
+mis = min(d$submit_time)
+mas = max(d$submit_time)
+interpolation_points = seq(from = 0, to = mas-mis, by = 60000)
+
+interpolateds=data.frame(time=double(),value=double(),id=character())
+
+for (filename in args$input_files) {
+  d=swf_read(filename)
+
+  d=d[order(d$submit_time),]
+  d$csum=cumsum(as.numeric(d$wait_time))
+  mis = min(d$submit_time)
+
+  interpolated = approx(x= d$submit_time-mis, y=d$csum, xout = interpolation_points, method = "constant")
+
+  interpolateds<-rbind(interpolateds,data.frame(time=interpolated$x,value=interpolated$y,id=filename))
+}
+
+g_interpolateds = group_by(interpolateds,time)
[…]

### 2017-03-28 misc/visu/average_cumobj_plot.R

@@ -1,0 +1,53 @@
+#!/usr/bin/env Rscript
+
+'usage: tool.R <input_files> ...  [-o <output>]
+tool.R -h | --help
+
+options:
+<input_files>                   The input data.
+-o <output>                     Output file.
+-h , --help                     Show this screen.
+' -> doc
+
+library(docopt)
+args<- docopt(doc)
+
+#library('TTR')
+#library('gridExtra')
+library('ggplot2')
+#library('reshape2')
+#library('plyr')
+library('dplyr')
+
+theme_bw_tuned<-function()
+{
+  return(theme_bw() +theme(
+                           plot.title = element_text(face="bold", size=10),
+                           axis.title.x = element_text(face="bold", size=10),
+                           axis.title.y = element_text(face="bold", size=10, angle=90),
+                           axis.text.x = element_text(size=10),
+                           axis.text.y = element_text(size=10),
+                           panel.grid.minor = element_blank(),
+                           legend.key = element_rect(colour="white"))
+  )
+}
+
+timestamp_to_date <- function(timestamp){
+  return(as.POSIXct(timestamp, origin="1970-01-01 01:00.00", tz="Europe/Paris"))
+}
+
+dfs=data.frame()
+
+
+for (filename in args$input_files) {
+  d=read.table(filename)
+  d$name=filename
+  dfs<-rbind(dfs,d)
+}
+
+pdf(args$o,width=20)
+summary(dfs)
+ggplot(dfs,aes(x=time,y=mean_obj,ymin=min_obj,ymax=max_obj,fill=name,color=name),)+
+geom_line()+
+geom_ribbon(alpha=0.01,linetype="dashed")+
+theme_bw()

### 2017-03-28 misc/visu/compare_weekvalues.R

@@ -1,0 +1,32 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+library(reshape2)
+
+'usage: compare_weekvalues.R <input> [--out=<output>]
+        compare_weekvalues.R -h | --help
+
+options:
+ <input>         Input file  [Default: none]
+ --out=<output>  Output file [Default: tmpbars.pdf]
+ ' -> doc
+
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+
+df=read.csv(args$"<input>",sep=' ',header=FALSE, stringsAsFactors=FALSE)
+names(df)=c('id','depth','backcount')
+df$id = 1:nrow(df)
+df2=melt(df,id.vars=c('id'))
+summary(df2)
+
+plot=ggplot(df2,aes(id,value,color=variable))+
+  geom_point()+
+  geom_line()+
+  xlab("Week")+
+  ylab("Value")
+
+pdf(file=args$out,width=20,height=7)
+plot

### 2017-03-28 misc/visu/comparedatasets.R

@@ -1,0 +1,35 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+
+'usage: tool.R <input1> <input2>
+        tool.R -h | --help
+
+options:
+ <input1>        The first data file.
+ <input2>        The second tata file
+ ' -> doc
+
+args<- docopt(doc)
+
+data1=read.csv(args$input1,sep=',',header=FALSE)
+data2=read.csv(args$input2,sep=',',header=FALSE)
+
+mnm=c('q','n','y',
+      'resres',
+      'resjobq',
+      'iwk_now',
+      'iday_now',
+      'restime',
+      'resourcestate_free',
+      'l_j_running',
+      'q_running',
+      'l_waitqueue',
+      'q_waitqueue')
+
+names(data1)=mnm
+names(data2)=mnm
+
+summary(data1)
+summary(data2)

### 2017-03-28 misc/visu/comparetable.R

@@ -1,0 +1,65 @@
+#!/usr/bin/env Rscript
+library(dplyr)
+library(docopt)
+library(ggplot2)
+library(reshape2)
+library(xtable)
+
+'usage: tool.R <input1> <input2> [--o=<out> ]
+        tool.R -h | --help
+ ' -> doc
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+
+load <- function(filename,id,idsec) {
+  data=read.csv(filename,sep=' ',header=FALSE)
+
+  names(data)=c('meanwait','maxwait','meanbsld','maxbsld','meanflow','maxflow','name')
+  A <- function(x) gsub(":1","",unlist(strsplit(x,"[.]"))[as.numeric(id)])
+  B <- function(x) gsub(":1","",unlist(strsplit(x,"[.]"))[as.numeric(idsec)])
+
+  data$Reservation=sapply(data$name,A)
+  data$Backfilling=sapply(data$name,B)
+
+  df=merge(merge(aggregate(meanwait ~ Reservation * Backfilling, data = data, FUN = mean),
+                 aggregate(maxwait ~ Reservation * Backfilling, data = data, FUN = max),
+                 by=c("Reservation","Backfilling")),
+           merge(aggregate(meanbsld ~ Reservation * Backfilling, data = data, FUN = mean),
+                 aggregate(maxbsld ~ Reservation * Backfilling, data = data, FUN = max),
+                 by=c("Reservation","Backfilling")),
+           by=c("Reservation","Backfilling"))
+
+  df$meanwait=as.numeric(df$meanwait)
+  df$meanbsld=as.numeric(df$meanbsld)
+  df$maxwait=as.numeric(df$maxwait)
+  df$maxbsld=as.numeric(df$maxbsld)
+
+  return(df)
+}
+
+dfm=load(args$input1,3,2)
+dfm2=load(args$input2,3,2)
+head(dfm)
+head(dfm2)
+
+wspf=dfm[which(dfm$Backfilling=="exp" & dfm$Reservation=="exp"),]$meanwait
+wspf2=dfm2[which(dfm2$Backfilling=="exp" & dfm2$Reservation=="exp"),]$meanwait
+
+wsqf=dfm[which(dfm$Backfilling=="sqf" & dfm$Reservation=="sqf"),]$meanwait
[…]

### 2017-03-28 misc/visu/controlhist.R

@@ -1,0 +1,37 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+
+'usage: tool.R <input1> <input2> [-o=<output1>]
+        tool.R -h | --help
+
+options:
+ <input1>        The input data for truthful mcts.
+ <input2>        The input data for cheating mcts.
+ -o=<output>  Output file [Default: tmp.pdf]
+ ' -> doc
+
+args<- docopt(doc)
+
+data1=read.csv(args$input1,sep=' ',header=FALSE)
+data1$type="MCTS"
+names(data1)=c('sjbf','fcfs','ratio','type')
+data2=read.csv(args$input2,sep=' ',header=FALSE)
+data2$type="MCTS-cheat"
+names(data2)=c('sjbf','fcfs','ratio','type')
+
+if (all(data2$ratio==Inf)) {
+ data2=rbind(data2,list(0,0,99,'MCTS-cheat'))
+}
+
+df=rbind(data1,data2)
+
+p=ggplot(df,aes(x=ratio,fill=type,color=type))+
+  geom_histogram(aes(y=(..density..)),position="dodge")+
+  geom_density(alpha=0.1,aes(color=type,fill=type))+
+  xlab("Control ratio")+
+  ylab("Proportion of experiments")
+
+pdf(file=args$o,width=20,height=7)
+p

### 2017-03-28 misc/visu/cumobj_diff.R

@@ -1,0 +1,72 @@
+#!/usr/bin/env Rscript
+
+'usage: tool.R <input_files> ...  [-o <output>] [-t <type> ... ]
+tool.R -h | --help
+
+options:
+<input_files>                   The input data.
+-o <output>                     Output file.
+-t <type>                     type file.
+-h , --help                     Show this screen.
+' -> doc
+
+library(docopt)
+args<- docopt(doc)
+
+#library('TTR')
+#library('gridExtra')
+library('ggplot2')
+#library('reshape2')
+#library('plyr')
+library('dplyr')
+
+theme_bw_tuned<-function()
+{
+  return(theme_bw() +theme(
+                           plot.title = element_text(face="bold", size=10),
+                           axis.title.x = element_text(face="bold", size=10),
+                           axis.title.y = element_text(face="bold", size=10, angle=90),
+                           axis.text.x = element_text(size=10),
+                           axis.text.y = element_text(size=10),
+                           panel.grid.minor = element_blank(),
+                           legend.key = element_rect(colour="white"))
+  )
+}
+
+swf_read <- function(f)
+{
+  df <- read.table(f,comment.char=';')
+  names(df) <- c('job_id','submit_time','wait_time','run_time','proc_alloc','cpu_time_used','mem_used','proc_req','time_req','mem_req','status','user_id','group_id','exec_id','queue_id','partition_id','previous_job_id','think_time')
+  return(df)
+}
+
+timestamp_to_date <- function(timestamp){
+  return(as.POSIXct(timestamp, origin="1970-01-01 01:00.00", tz="Europe/Paris"))
+}
+
+d=swf_read(args$input_files[1])
+mis = min(d$submit_time)
+mas = max(d$submit_time)
+interpolation_points = seq(from = 0, to = mas-mis, by = 60000)
+interpolateds=data.frame(time=double(),value=double(),id=character())
+
+
+d=d[order(d$submit_time),]
+d$csum=cumsum(as.numeric(d$wait_time))
+interp1 = approx(x= d$submit_time-mis, y=d$csum, xout = interpolation_points, method = "constant")
+
+for (i in 2:length(args$input_files)) {
+  filename=args$input_files[i]
+  d=swf_read(filename)
+
+  d=d[order(d$submit_time),]
+  d$csum=cumsum(as.numeric(d$wait_time))
+  mis = min(d$submit_time)
+
[…]

### 2017-03-28 misc/visu/cumobj_diffplot.R

@@ -1,0 +1,123 @@
+#!/usr/bin/env Rscript
+
+'usage: tool.R <input_files> ...  [-o <output>] [-i <scalefile>] [--legend] [-t <trace>]
+tool.R -h | --help
+
+options:
+<input_files>                   The input data.
+-o <output>                     Output file.
+-i <scalefile>                  Scale file.
+-t <trace>                      Trace name
+-h , --help                     Show this screen.
+--legend                        Show the legend
+' -> doc
+
+library(docopt)
+args<- docopt(doc)
+
+library('ggplot2')
+library('dplyr')
+
+
+theme_bw_tuned<-function()
+{
+  return(theme_bw() +theme(
+                           plot.title = element_text(face="bold", size=10),
+                           axis.title.x = element_text(face="bold", size=10),
+                           axis.title.y = element_text(face="bold", size=10, angle=90),
+                           axis.text.x = element_text(size=10),
+                           axis.text.y = element_text(size=10),
+                           panel.grid.minor = element_blank(),
+                           legend.key = element_rect(colour="white"))
+  )
+}
+
+dfs=data.frame()
+
+
+for (filename in args$input_files) {
+  d=read.table(filename)
+  d$name=filename
+  dfs<-rbind(dfs,d)
+}
+
+
+dummyqp <- function (x) {
+  as.numeric(quantile(x,probs=c(0.9),na.rm=TRUE))
+}
+dummyqn <- function (x) {
+  as.numeric(quantile(x,probs=c(0.1),na.rm=TRUE))
+}
+
+dscale=read.table(args$i)
+dfs$value=dfs$value/dscale$mean_obj[nrow(dscale)]
+
+dfs = group_by(dfs,time,id)
+dfs=summarise(dfs,mean(value),max(value),min(value),sd(value),dummyqp(value),dummyqn(value))
+
+names(dfs)=c("time","name","mean_obj","max_obj","min_obj","sd_obj","sdn","sdp")
+
+#dfs$time=lapply(dfs$time,timestamp_to_date)
+
+timestamp_to_date <- function(timestamp){
[…]

### 2017-03-28 misc/visu/draw_3dperf.R

@@ -1,0 +1,42 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+library(scatterplot3d)
+
+'usage: tool.R <input> [--out=<output>]
+        tool.R -h | --help
+
+options:
+ <input>    Input file  [Default: none]
+ ' -> doc
+
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+
+df <- data.frame(id=double(),
+                 meanwait=double(),
+                sd=double(),
+                q1=double(),
+                q2=double(),
+                name=character(),
+                explo=double(),
+                budget=double())
+
+df=read.csv(args$"<input>",sep=' ',header=TRUE, stringsAsFactors=FALSE)
+
+names(df)=c('meanwait','maxwait','meanbsld','maxbsld','meanflow','maxflow','name','explo','budget')
+
+print(df)
+
+pdf(file=args$out,width=20,height=7)
+with(df, {
+   scatterplot3d(budget, explo, meanwait,
+                 color="blue", pch=19,
+                 type="h",
+                 main="search grid for MCTS budget vs exploration param.",
+                 xlab="MCTS search budget",
+                 ylab="Exploration parameter",
+                 zlab="Average waiting time")
+})

### 2017-03-28 misc/visu/draw_barline.R

@@ -1,0 +1,75 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+
+'usage: tool.R --in=<input>... --lab=<lab>... [--out=<output>] [--out2=<output>]
+        tool.R -h | --help
+
+options:
+ --in=<input>    Input file  [Default: none]
+ --lab=<lab>     Plot label  [Default: none]
+ --out=<output>  Output file [Default: tmpbars.pdf]
+ --out2=<output>  Output file [Default: tmpnormalized.pdf]
+ ' -> doc
+
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+
+df <- data.frame(id=double(),
+                 meanwait=double(),
+                sd=double(),
+                q1=double(),
+                q2=double(),
+                Policy=character())
+
+names(df)=c('id','meanwait','sd','q1','q2','Policy')
+
+for (i in 1:length(args$"--in")){
+  data=read.csv(args$"--in"[i],sep=',',header=TRUE, stringsAsFactors=FALSE)
+  data$Policy=args$"--lab"[i]
+  names(data)=c('truc','id','meanwait','sd','q1','q2','Policy')
+
+  df <- rbind.data.frame(data,df)
+}
+
+names(df)=c('truc','id','meanwait','sd','q1','q2','Policy')
+mww=df[which(df$Policy=="fcfs"),]$meanwait
+df$nmw=df$meanwait / mww
+
+theme_bw_tuned<-function()
+{
+	return(theme_bw() +theme(
+		plot.title = element_text(face="bold", size=10),
+		axis.title.x = element_text(face="bold", size=10),
+		axis.title.y = element_text(face="bold", size=10, angle=90),
+		axis.text.x = element_text(size=10),
+		axis.text.y = element_text(size=10),
+		panel.grid.minor = element_blank(),
+		legend.key = element_rect(colour="white"))
+         )
+}
+
+plot=ggplot(df,aes(id,meanwait,color=Policy))+
+  geom_point()+
+  geom_line()+
+  geom_errorbar(aes(ymin=q1,ymax=q2),width=0.01)+
+  xlab("Week number")+
+  ylab("AvgWait performance")+
+  scale_color_brewer(palette="Dark2")+
+  theme_bw_tuned()
+
+pdf(file=args$out,width=10,height=2)
+plot
+
+plot=ggplot(df,aes(id,nmw,color=Policy))+
[…]

### 2017-03-28 misc/visu/draw_cumulative.R

@@ -1,0 +1,230 @@
+#!/usr/bin/env Rscript
+
+'usage: tool.R <input_file> ... [-o <output_pdf>] [-l <lineinput>]
+tool.R -h | --help
+
+options:
+ <input_file>                   The input data.
+ -o <output_pdf>                Output file in case of pdf/tikz/png output.
+ -l <lineinput>                lineinput.
+ -h , --help                    Show this screen.
+ ' -> doc
+
+library(docopt)
+args<- docopt(doc)
+print(args)
+args$r=as.numeric(args$r)
+args$H=as.numeric(args$H)
+
+png(file=args$o,width=1800,height=800)
+
+library('TTR')
+library('gridExtra')
+library('ggplot2')
+library('reshape2')
+library('plyr')
+
+theme_bw_tuned<-function()
+{
+	return(theme_bw() +theme(
+		plot.title = element_text(face="bold", size=10),
+		axis.title.x = element_text(face="bold", size=10),
+		axis.title.y = element_text(face="bold", size=10, angle=90),
+		axis.text.x = element_text(size=10),
+		axis.text.y = element_text(size=10),
+		panel.grid.minor = element_blank(),
+		legend.key = element_rect(colour="white"))
+         )
+}
+
+swf_read <- function(f)
+{
+    	df <- read.table(f,comment.char=';')
+print(f)
+    	names(df) <- c('job_id','submit_time','wait_time','run_time','proc_alloc','cpu_time_used','mem_used','proc_req','time_req','mem_req','status','user_id','group_id','exec_id','queue_id','partition_id','previous_job_id','think_time')
+	return(df)
+}
+
+flow <- function(swf){
+  nbefore = nrow(data)
+  data = data[which(!is.na(data$wait_time)),]
+  if( nbefore != nrow(data))
+    print(paste("There were", nbefore-nrow(data), "jobs with corrupted wait time or run time"))
+  return(data$wait_time)
+}
+
+utilization <- function( data,utilization_start=0 )
+{
+
+  data = arrange(data, submit_time)
+  # get start time and stop time of the jobs
+  start <- data$submit_time + data$wait_time
+  stop <- data$submit_time + data$wait_time + data$run_time
+  # because jobs still running have an -1 runtime
+  stop[which(data$run_time == -1 & data$wait_time != -1)] = max(stop)
[…]

### 2017-03-28 misc/visu/draw_evo.R

@@ -1,0 +1,143 @@
+#!/usr/bin/env Rscript
+library(dplyr)
+library(docopt)
+library(ggplot2)
+library(reshape2)
+library(xtable)
+
+'usage: tool.R [ --intrain=<train>... ] [--intest=<test>... ] [--th=<th>... ] [--owait=<out>] [--obsld=<out>] [--tname=<tname>]
+
+        tool.R -h | --help
+ ' -> doc
+
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+scalar1 <- function(x) {(x-min(x))/(max(x)-min(x))}
+
+load <- function(filename,id,idsec) {
+  data=read.csv(filename,sep=' ',header=FALSE)
+
+  names(data)=c('meanwait','maxwait','meanbsld','maxbsld','meanflow','maxflow','name')
+
+  A <- function(x) gsub(":1","",unlist(strsplit(x,"[.]"))[as.numeric(id)])
+  B <- function(x) gsub(":1","",unlist(strsplit(x,"[.]"))[as.numeric(idsec)])
+  data$Reservation=sapply(data$name,A)
+  data$Backfilling=sapply(data$name,B)
+
+  print(head(data))
+
+  C <- function(x) paste(unlist(strsplit(x,"[.]"))[as.numeric(id+1)],unlist(strsplit(x,"[.]"))[as.numeric(id+2)],unlist(strsplit(x,"[.]"))[as.numeric(id+3)])
+  data$Id=sapply(data$name,C)
+
+  eid=which(data$Backfilling=="fcfs" & data$Reservation=="fcfs")
+  data$meanwait=data$meanwait/data$meanwait[eid]
+  data$meanbsld=data$meanbsld/data$meanbsld[eid]
+  data$maxbsld=data$maxbsld/data$maxbsld[eid]
+  data$maxwait=data$maxwait/data$maxwait[eid]
+
+  df=merge(merge(aggregate(meanwait ~ Reservation * Backfilling, data = data, FUN = mean),
+                 aggregate(maxwait ~ Reservation * Backfilling, data = data, FUN = mean),
+                 by=c("Reservation","Backfilling")),
+           merge(aggregate(meanbsld ~ Reservation * Backfilling, data = data, FUN = mean),
+                 aggregate(maxbsld ~ Reservation * Backfilling, data = data, FUN = mean),
+                 by=c("Reservation","Backfilling")),
[…]

### 2017-03-28 misc/visu/draw_fig_easy_no_opt.R

@@ -1,0 +1,127 @@
+#!/usr/bin/Rscript
+
+library(docopt)
+library(ggplot2)
+
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+
+
+'usage: draw_fig_easy_no_opt.R <input> ... [--mean=<output1>] [--maxbsld=<output2>] [--maxwait=<output3>]
+        draw_fig_easy_no_opt.R -h | --help
+
+options:
+ <input>                   The input data.
+ --mean=<output>           Output file [Default: tmp1.pdf]
[…]

### 2017-03-28 misc/visu/draw_histogram_wait_fromshell.R

@@ -1,0 +1,47 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+
+'usage: tool.R <input> ... [--mean=<output1>] [--max=<output2>]
+        tool.R -h | --help
+
+options:
+ <input>                   The input data.
+ --mean=<output>              Output file [Default: tmp1.pdf]
+ --max=<output>              Output file [Default: tmp2.pdf]
+ ' -> doc
+
+args<- docopt(doc)
+
+df=data.frame()
+dfm=data.frame()
+for (filename in args$input) {
+  data=read.csv(filename,sep=' ',header=FALSE)
+  names(data)=c('meanwait','maxwait','name')
+  data$type=filename
+  df=rbind(df,data)
+  dfm=rbind(dfm,data.frame(meanmax=mean(data$maxwait),meanmean=mean(data$meanwait),type=filename))
+}
+
+bwmean<-(max(df$meanwait)-min(df$meanwait))/30
+pmeanwait=ggplot(df,aes(x=meanwait,fill=type,color=type))+
+  geom_histogram(aes(y=(..density..)),position="dodge", binwidth=bwmean)+
+  geom_point(data=dfm,aes(y=-0.0001, x=meanmean,fill=type))+
+  geom_vline(data=dfm,aes(xintercept=meanmean,color=type,fill=type))+
+  geom_density(alpha=0.1,aes(color=type,fill=type))+
+  xlab("Average Waiting Time")+
+  ylab("Proportion of experiments")
+
+bwmax<-(max(df$maxwait)-min(df$maxwait))/30
+pmaxwait=ggplot(df,aes(x=maxwait,fill=type))+
+  geom_histogram(aes(y=(..density..)),position="dodge", binwidth=bwmax)+
+  geom_point(data=dfm,aes(y=-0.0001, x=meanmax,fill=type))+
+  geom_vline(data=dfm,aes(xintercept=meanmax,color=type,fill=type))+
+  xlab("Max Waiting Time")+
+  ylab("Proportion of experiments")
+
+pdf(file=args$mean,width=20,height=7)
+pmeanwait
+pdf(file=args$max,width=20,height=7)
+pmaxwait

### 2017-03-28 misc/visu/draw_intra.R

@@ -1,0 +1,70 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+
+'usage: tool.R <input> ... [--mean=<output1>] [--max=<output2>]
+        tool.R -h | --help
+
+options:
+ <input>                   The input data.
+ --mean=<output>              Output file [Default: tmp1.pdf]
+ --max=<output>              Output file [Default: tmp2.pdf]
+ ' -> doc
+
+args<- docopt(doc)
+
+df=data.frame()
+dfm=data.frame()
+size = 0
+
+for (filename in args$input) {
+  data=read.csv(filename,sep=' ',header=FALSE)
+  names(data)=c('meanwait','maxwait','meanbsld','maxbsld','meanflow','maxflow','name')
+  fstrow = data[1,]
+  data = data[2:nrow(data),]
+  size = nrow(data)
+  final_names = data$name
+  for (i in 1:nrow(data)) {
+    data$meanwait[i] = data$meanwait[i] / fstrow$meanwait[1]
+    data$maxwait[i] = data$maxwait[i] / fstrow$maxwait[1]
+  }
+  data$name = 1:nrow(data)
+  df=rbind(df,data)
+}
+
+df$name = final_names[df$name]
+
+bwmean<-(max(df$meanwait)-min(df$meanwait))/30
+
+df=df[complete.cases(df),]
+print(summary(df))
+
+dfm=data.frame()
+for (n in unique(df$name)) {
+  dfn=df[which(df$name==n),]
+  dfm=rbind(dfm,data.frame(meanmax=mean(dfn$maxwait),meanmean=mean(dfn$meanwait),name=n))
+}
+
+pmeanwait=ggplot(df,aes(x=meanwait,fill=name))+
+  geom_histogram(aes(y=(..density..)),position="dodge", binwidth=bwmean)+
+  geom_point(data=dfm,aes(y=-0.0001, x=meanmean,fill=name))+
+  geom_vline(data=dfm,aes(xintercept=meanmean,color=name,fill=name))+
+  geom_density(alpha=0.1,aes(color=name,fill=name))+
+  xlab("Average Waiting Time")+
+  ylab("Proportion of experiments")
+
+bwmax<-(max(df$maxwait)-min(df$maxwait))/30
+
+pmaxwait=ggplot(df,aes(x=maxwait,fill=name))+
[…]

### 2017-03-28 misc/visu/draw_mixedplot.R

@@ -1,0 +1,67 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+
+'usage: tool.R --in=<input>... --lab=<lab>... --type=<type>... [--out=<output>]
+        tool.R -h | --help
+
+options:
+ --in=<input>    Input file  [Default: none]
+ --lab=<lab>     Plot label  [Default: none]
+ --out=<output>  Output file [Default: tmpmixed.pdf]
+ ' -> doc
+
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+
+df <- data.frame(meanwait=double(),
+                sd=double(),
+                q1=double(),
+                q2=double(),
+                name=double(),
+                type=character(),
+                type2=character(),
+                stringsAsFactors=FALSE)
+names(df)=c('meanwait','sd','q1','q2','name','type')
+
+
+for (i in 1:length(args$"--in")){
+  data=read.csv(args$"--in"[i],sep=' ',header=FALSE, stringsAsFactors=FALSE)
+  names(data)=c('meanwait','maxwait','meanbsld','maxbsld','meanflow','maxflow','name')
+
+  if (args$lab[i] == "expansionfactor") lab <- 0
+  else if (args$lab[i] == "LF") lab <- 1
+  else lab <- as.numeric(args$lab[i])
+
+  if (args$type[i] == "train") type2 <- "training"
+  else type2 <- "testing"
+
+  quant=quantile(as.numeric(data$meanwait),probs=c(0.20,0.80))
+
+  nr = list(mean(as.numeric(data$meanwait)),
+            sd(as.numeric(data$meanwait)),
+            quant[[1]],
+            quant[[2]],
+            lab,
+            args$type[i],
+            type2)
+
+  df <- rbind.data.frame(nr,df)
+}
+
+names(df)=c('meanwait','sd','q1','q2','name','type','type2')
+summary(df)
+
+plot=ggplot(df,aes(name,meanwait,color=type))+
+  facet_grid(type2 ~ ., scales = "free_y") +
+  geom_point()+
+  geom_line()+
+  geom_errorbar(aes(ymin=q1,ymax=q2),width=0.01)+
+  #scale_x_continuous(limits=args$lab,labels=args$lab)+
+  xlab("Mixing proportion")+
+  ylab("Average waiting time performance")
+
+pdf(file=args$out,width=20,height=7)
+plot

### 2017-03-28 misc/visu/draw_pareto.R

@@ -1,0 +1,262 @@
+#!/usr/bin/env Rscript
+library(dplyr)
+library(docopt)
+library(ggplot2)
+library(reshape2)
+library(xtable)
+
+'usage: tool.R <title> <i1> <i2> <id1> <id2> <id3> <id4> [--os=<outs> ] [--outpareto=<out> ] [--outparetobsld=<out2> ] [--ot=<out3>] [--outparallel=<out3>] [--ob=<out3>]
+        tool.R -h | --help
+ ' -> doc
+
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+scalar1 <- function(x) {(x-min(x))/(max(x)-min(x))}
+
+load <- function(filename,id,idsec) {
+  data=read.csv(filename,sep=' ',header=FALSE)
+
+  names(data)=c('meanwait','maxwait','meanbsld','maxbsld','meanflow','maxflow','name')
+  A <- function(x) gsub(":1","",unlist(strsplit(x,"[.]"))[as.numeric(id)])
+  B <- function(x) gsub(":1","",unlist(strsplit(x,"[.]"))[as.numeric(idsec)])
+
+  data$Reservation=sapply(data$name,A)
+  data$Backfilling=sapply(data$name,B)
+  #d  = aggregate(meanwait ~ Reservation * Backfilling, data = data, FUN = mean)
+  #d2 = aggregate(maxwait ~ Reservation * Backfilling, data = data, FUN = max)
+       #aggregate(meanbsld ~ Reservation * Backfilling, data = data, FUN = mean),
+       #aggregate(maxbsld ~ Reservation * Backfilling, data = data, FUN = max),
+
+  df=merge(merge(aggregate(meanwait ~ Reservation * Backfilling, data = data, FUN = mean),
+                 aggregate(maxwait ~ Reservation * Backfilling, data = data, FUN = mean),
+                 by=c("Reservation","Backfilling")),
+           merge(aggregate(meanbsld ~ Reservation * Backfilling, data = data, FUN = mean),
+                 aggregate(maxbsld ~ Reservation * Backfilling, data = data, FUN = mean),
+                 by=c("Reservation","Backfilling")),
[…]

### 2017-03-28 misc/visu/flow_multiple.sh

@@ -1,0 +1,7 @@
+#!/usr/bin/env sh
+
+for f in $@;
+do
+  flow=$(cat $f | awk -f misc/visu/flow.awk)
+  echo "$flow $(basename $f)"
+done

### 2017-03-28 misc/visu/getbars.R

@@ -1,0 +1,58 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+library(ggplot2)
+
+'usage: tool.R <input>... [--out=<output>]
+        tool.R -h | --help
+
+options:
+ --in=<input>    Input file  [Default: none]
+ --lab=<lab>     Plot label  [Default: none]
+ --out=<output>  Output file [Default: tmpmixed.pdf]
+ ' -> doc
+
+args <- docopt(doc)
+
+options(stringsAsFactors=FALSE)
+
+df <- data.frame(id=double(),
+                 meanwait=double(),
+                sd=double(),
+                q1=double(),
+                q2=double())
+
+names(df)=c('id','meanwait','sd','q1','q2')
+
+
+for (i in 1:length(args$"<input>")){
+  data=read.csv(args$"<input>"[i],sep=' ',header=FALSE, stringsAsFactors=FALSE)
+  names(data)=c('meanwait','maxwait','meanbsld','maxbsld','meanflow','maxflow','name')
+
+  print(args$"<input>"[i])
+  dempt=data[which(is.na(data$meanwait)),]
+  print(dempt)
+  if (nrow(dempt)>0) {
+    print("ERROR: detected a NA in the awkstat file. Deleting the awkstats file AND the original simulated swf.")
+    system(paste("rm ",args$"<input>"[i],sep=""))
+    system(paste("rm o/perv-evolution.zmk/",data[which(is.na(data$meanwait)),2],sep=""))
+  }
+
+  quant=quantile(as.numeric(data$meanwait),probs=c(0.20,0.80))
+
+  nr = list(i,
+            mean(as.numeric(data$meanwait)),
+            sd(as.numeric(data$meanwait)),
+            quant[[1]],
+            quant[[2]])
+
+
+  df <- rbind.data.frame(nr,df)
+}
+print("EE")
+
+names(df)=c('id','meanwait','sd','q1','q2')
+
+print(df)
+
+write.csv(df,file=args$"out")

### 2017-03-28 misc/visu/knit.R

@@ -1,0 +1,19 @@
+#!/usr/bin/env Rscript
+
+library(docopt)
+
+'usage: tool.R <input>... [-o=<output1>]
+        tool.R -h | --help
+
+options:
+ <input1>        The input data for truthful mcts.
+ <input2>        The input data for cheating mcts.
+ -o=<output>  Output file [Default: tmp.pdf]
+ ' -> doc
+
+args<- docopt(doc)
+
+print(getwd())
+
+library(knitr)
+knit('misc/visu/means.Rtex',args$o)

### 2017-03-28 misc/visu/mosaic.R

@@ -1,0 +1,103 @@
+#!/usr/bin/env Rscript
+
+'usage: mosaic.R <input_files>... [-o <output>]
+tool.R -h | --help
+
+options:
+<input_files>                   The input data.
+-o <output>                     Output file.
+-h , --help                     Show this screen.
+' -> doc
+
+library(docopt)
+args<- docopt(doc)
+
+library('ggplot2')
+library('ggmosaic')
+library('dplyr')
+library('directlabels')
+
+
+theme_bw_tuned<-function()
+{
+  return(theme_bw() +theme(
+                           plot.title = element_text(face="bold", size=10),
+                           axis.title.x = element_text(face="bold", size=10),
+                           axis.title.y = element_text(face="bold", size=10, angle=90),
+                           axis.text.x = element_text(size=10),
+                           axis.text.y = element_text(size=10),
+                           panel.grid.minor = element_blank(),
+                           legend.key = element_rect(colour="black"))
+  )
+}
+
+dfs=data.frame()
+for (filename in args$input_files) {
+  d=read.delim(filename)
+  names(d)<-c("Policy")
+  d$t = seq(1,nrow(d))
+  d$t = d$t
+  dfs<-rbind(dfs,d)
+}
+
+dfs$t = floor(0.1 * dfs$t)
+dfs$t = dfs$t*10
+
+ti=table(dfs$t,dfs$Policy)
+tid=as.data.frame(prop.table(ti,1))
+
+names(tid)<-c("t","Policy","Freq")
+tid$t=as.numeric(tid$t)
+tid$Freq=as.numeric(tid$Freq)
+
+tid$t=tid$t*3600*24*7
+tmi=min(tid$t)
+tid$t=tid$t-tmi
+tma=max(tid$t)
+
+
+
+timestamp_to_date <- function(timestamp){
+  return(as.POSIXct(timestamp, origin="1970-01-01 01:00.00", tz="Europe/Paris"))
+}
+
+library(lubridate)
+mymonth <- function(x){
+  month(timestamp_to_date(x),label=TRUE)
+}
+
+ml=(tma-tmi)/6
+brkx=seq(from = tmi, to = tma, by = ml)
+labx=sapply(brkx,mymonth)
+
+pdf(args$o,width=5.5, height=4)
+df=arrange(tid,Policy,t)
+df$Policy=as.character(df$Policy)
[…]

### 2017-03-28 misc/store_annex.sh
packaging fixes

@@ -1,9 +1,0 @@
-#!/usr/bin/env bash
-
-mkdir ~/annex/experiments/$1
-cp -r o pdfs ~/annex/experiments/$1
-git log | head -n 1 > ~/annex/experiments/$1/metadata
-cd ~/annex/experiments
-echo "adding"
-git annex add $1
-git commit -m "added experiment: $1" -a

### 2017-03-28 misc/submit.sh

@@ -1,4 +1,0 @@
-#!/usr/bin/env bash
-
-jid=$(ts bash -c "$(cat $1)")
-ts -w $jid

### 2017-03-29 paper/paper.tex
added paper

\documentclass[sigconf]{acmart}
\setcopyright{rightsretained}

\usepackage{listings}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{todo}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[scientific-notation=true]{siunitx}

\usepackage{amssymb}
\usepackage{tabulary}

\usepackage{etoolbox}

\usepackage{algorithm}
\usepackage{algorithmic}

\makeatletter
\newcommand\fs@norules{\def\@fs@cfont{\bfseries}\let\@fs@capt\floatc@ruled
  \def\@fs@pre{}%
  \def\@fs@post{}%
  \def\@fs@mid{\kern3pt}%
  \let\@fs@iftopcapt\iftrue}
\makeatother
\floatstyle{norules}
\restylefloat{algorithm}

\let\tinymatrix\smallmatrix
\let\endtinymatrix\endsmallmatrix
\patchcmd{\tinymatrix}{\scriptstyle}{\scriptscriptstyle}{}{}
\patchcmd{\tinymatrix}{\scriptstyle}{\scriptscriptstyle}{}{}
\patchcmd{\tinymatrix}{\vcenter}{\vtop}{}{}
\patchcmd{\tinymatrix}{\bgroup}{\bgroup\scriptsize}{}{}

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\usepackage[draft,index]{fixme}
\fxsetup{theme=color,mode=multiuser,layout=inline,draft}

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Online (Bandit) Policy Selection for EASY-Backfilling}

\author{Eric Gaussier}
\affiliation{%
  \institution{Univ. Grenoble Alpes, CNRS, Grenoble INP, LIG}
  \country{France}}
\email{eric.gaussier@imag.fr}
\author{J\'er\^ome Lelong}
\affiliation{%
  \institution{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LJK}
  \country{France}}
\email{jerome.lelong@imag.fr}
\author{Valentin Reis}
\affiliation{%
  \institution{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LIG}
  \country{France}}
\email{valentin.reis@imag.fr}
\author{Denis Trystram}
\affiliation{%
  \institution{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LIG}
  \country{France}}
\email{denis.trystram@imag.fr}

\begin{abstract}

  The EASY-FCFS heuristic is the basic building block of job scheduling
  policies in most parallel High Performance Computing (HPC) platforms. Despite
  its good properties (simplicity, no starvation), it could still be improved
  on a per-system basis. This tuning process is difficult because of
  non-linearities in the scheduling process. The study proposed here considers
  an online approach to the automatic tuning of the EASY heuristic for HPC
  platforms. More precisely, we consider the problem of selecting a reordering
  policy of the job queue under several feedback modes. We show via a
  comprehensive experimental campaign that noisy feedback (using a weak
  simulator) recovers existing in-hindsight results that allow to divide the
  average waiting time up to a factor of 2. Moreover, we show that bandit
  feedback can be used by a simple multi-armed bandit algorithm to decrease the
  average waiting time down to 40\% of its original value without using a
  simulator.

\end{abstract}

\maketitle

\section{Introduction}
\label{sec:intro}

Plan :
\begin{itemize}
\item Context - message : les machines sont de + en + complexes, les systemes
de management restent rudimentaires face a la complexite croissante de
l'architecture et des applications.
\item les habitudes et usages existants : justification du FCFS-EASY-BF,
developper rapidement ce qui existe
\item ce que l'on propose
\end{itemize}

Providing the computing infrastuctures needed to solve actual complex problems
arising in the various fields of the modern society (including climate change,
health, green energy or security) is a strategic challenge.
In particular, High Performance Computing (HPC) systems are evolving to
extreme-scale parallel and distributed platforms.
The race for always more computing power and storage capacity does not only
lead to sophisticated specific exascale platforms, but the objective of the
community is to design efficient sustained Petascale platforms.
However, there is still a long way to go and many scientific and technical
problems to solve by the system management software in order to adapt to such
large-scale evolution and the increasing complexity
in both architecture and applications.
The existing job and resource management sofwares allow to run tens thousands
jobs on hundreds thousands cores.
They are based on robust and rather simple policies.

\bigskip
The management generates a huge amount of data.
During the last few years, there was an explosion of the number of of works at
the interface of HPC and bigdata, dealing with learning algorithms
(including the authors of this paper~\cite{Gaussier2015ImprovingBB}).
Most of these studies target the process to determine more or less accurately
the value of some specific key parameters, with the idea of
learning better estimates should improve the performances of the resource
manager.

\bigskip
Our idea within this work is to learn how the management system behave within
the classical framework of of selecting jobs of the submission queue under
EASY-backfilling.
Contrarily to most existing studies which consider only one execution trace, we
extend the learning on a multiple trace basis.
Indeed, learning for a specific trace leads to limited results, customized for
one trace on a given platform (what happens if we target another platform?).
Moreover, this is usually inefficient to deal with extreme unpredictable
events.

\subsection{Contributions}

\section{Related Works}
\label{sec:rw}

This section presents related works in the area of scheduling for batch scheduling.

\subsection{Scheduling heuristics in HPC platforms}

While parallel job scheduling is a well studied theoretical
problem~\cite{leung2004handbook}, the practical ramifications, varying
hypotheses, and inherent uncertainty of the problem in HPC have driven
practitioners and researchers alike to use and study simple heuristics.
The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} Backfilling.

While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
significant computational overhead, perhaps explaining why most of the machines
of the top500 ranking~\cite{top500} still use at the time of this publication a
variant of EASY Backfilling.

\subsection{EASY}
There is a large body of work seeking to improve EASY. Indeed, while the
heuristic is used by various resource and job management softwares (most notably
SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by system
administrators.

Several works explore how to tune EASY by reordering waiting
and/or backfilling queues~\cite{Tsafrir_easypp_2005}, sometimes even in a
randomized manner~\cite{1592720}, as well as some
implementations~\cite{Jackson2001}. However, as successful as they may be, these works
do not address the dependency~\cite{variability} of scheduling metrics on the
workload. Indeed these studies most often report \textit{post-hoc} performance since
they compare algorithms after the workload is known.

The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
method to tuning these queues, although it requires simulated scheduling runs
at decision time and therefore costs much more than the natural execution of EASY.

\subsection{Data-aware resource management}

There is a recent focus on leveraging the high amount of data available in
large scale computing systems in order to improve their behavior. Some works
use collaborative filtering to colocate tasks in clouds by estimating
application interference~\cite{7516031}.  Others are closer to the application
level and use binary classification to distinguish benign memory faults from
application errors in order to execute recovery algorithms (see
~\cite{fmodeling} for instance).

Several works use this method in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.

However, these works do not address the duality between the cumulative and
maximal scheduling costs, as mentionned in~\cite{learningruntimes}.

While these previous works intend to estimate uncertain parameters, we consider in this paper a more
pragmatic approach, which directly learn a good scheduling policy from a given policy space.

\subsection{Multi-Armed Bandits}

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards. At every round, an action (an arm) must be chosen
in a fixed set and the corresponding reward is observed. The goal of a MAB
algorithm is to maximize the total reward obtained in a number of rounds.
There are a number of works that address this problem under a variety of
constraints, The two most popular settings are the original stochastic  case,
and the adversarial case. See ~\cite{thompson} for the original work on the
stochastic case and ~\cite{Auer2002} for the UCB family of algorithms. The work
of ~\cite{Banos} is the earliest known work to us for the adversarial case
see~\cite{nonstoch} for the EXP family of algorithms. We refer to the
review~\cite{bubnow} for a comprehensive overview of the field. While these
algorithms bound the cumulative difference in loss to the best arm (the
regret), they have functional constraits such as the fact that the rewards
should be contained in a range. The\cite{Auer2002} heuristic called
Epsilon-Greedy is known to achieve good results in practice, and does not have
this requirement.

\section{Problem setting}
\label{sec:problem_setting}

This section presents the systems under study and the scheduling problem.  It
first introduces the EASY-Backfilling heuristic and gives \todo{fix} the
problem statement.

\subsection{System Description}
\label{sub:sysdesc}

The problem addressed in this work is the core logic of Resource and Job
Management Systems (RJMS) such as SLURM~\cite{SLURMdocSCHED}, PBS~\cite{PBSdoc},
OAR~\cite{capit2005batch}, Cobalt~\cite{Cobalt}, and more recently Flux~\cite{flux2014}.

The crucial part of batch scheduling software is the scheduling algorithm that determines
where and when the submitted jobs are executed. The process is as follows: jobs
are submitted by end-users and queued until the scheduler selects one of them
for running. Each job has a provided bound on the execution time and some
resource requirements (number and type of processing units). Then, the RJMS
drives the search for the resources required to execute this job. Finally, the
tasks of the job are assigned to the chosen nodes.

In the classical case, these softwares need to execute a set of concurrent
parallel jobs with rigid (known and fixed) resource requirements on a HPC
platform represented by a pool of $m$ identical resources. This is an on-line
problem since the jobs are submitted over time and their characteristics are only known when they are released.
Below is the description and the notations of the characteriscs of job $j$:

\begin{itemize}
  \item Submission date $r_j$ (also called \textit{release date})

  \item Resource requirement $q_j$ (number of processors)

  \item Actual running time $p_j$ (sometimes called \textit{processing time})

  \item Requested running time $\widetilde{p_j}$ (sometimes called \textit{walltime}), which is an upper bound of $p_j$.

\end{itemize}

The resource requirement $q_j$ of job $j$ is known when the job is submitted at
time $r_j$, while the requested running time $\widetilde{p_j}$ is given by the
user as an estimate. Its actual value $p_j$ is only known \textit{a posteriori}
when the job really completes.  Moreover, the users have incentive to
over-estimate the actual values, since jobs may be ``killed'' if they
surpass the provided value.
\subsection{EASY Backfilling}
\label{sub:easy}

The selection of the job to run is performed according to a
scheduling policy that establishes the order in which the jobs are executed.
EASY-Backfilling is the most widely used policy due to its simple and robust
implementation and known benefits such as high system
utilization~\cite{easy}. This strategy has no worst case guarantee beyond
the absence of starvation (i.e. every job will be scheduled at some moment).

The EASY-FCFS heuristic uses a job queue to select and backfill jobs.  At any
time that requires a scheduling decision (i.e. job submission or termination),
the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
and starts jobs until it finds a job that can not be started right away. It
then makes a reservation for this job at the earliest predictable time and
starts \textit{backfilling} the job queue in FCFS order, starting any job that
does not delay the unique reservation.

\subsection{Scheduling Objective}
\label{sub:scheduling_objectives}

A system administrator may use one or multiple cost metric(s). Our study of
scheduling performance relies on the waiting times of the jobs, which is one of
the more commonly used objectives.

    \begin{equation}
      \textbf{Wait}_j =  start_j-r_j
    \end{equation}

Like other cost metrics, the waiting time is usually considered in its
\textit{cumulative} version, which means that one seeks to minimize the average
waiting time (\textbf{AvgWait}). It is worth noting that the \textbf{MaxWait},
a.k.a the maximal value of the waiting time of all the jobs is also worthy of
interest. Unfortunately, these criteria can sometimes be dual in practice,
making the problem bi-objective. Subsection~\ref{sub:th} will outline our
approach to managing this aspect.

\subsection{Problem Statement}

Our problem statement is: \textbf{How to tune EASY-Backfilling in an online manner?}
\todo{introduce the difference in workloads, the duality of avg and max, and the online problem}

\section{Tuning EASY by reordering and thresholding the job queue}
\label{sec:framework}

This section presents two mechanisms for safely tuning the EASY-Backfilling:
job queue reordering and job thresholding. Together, these two building blocks
constitute a robust framework for tuning EASY.

\subsection{Reordering the job Queue}

The EASY heuristic uses a job queue to select and backfill jobs. While this job
queue is ordered in FCFS order in the original heuristic, it is possible to
reorder it at will. We settle on a reasonable search space of 10 reordering policies.

%wait mwait lpf spf sqf lqf expfact mexpfact lrf srf laf saf
\begin{enumerate}
  \item FCFS: First-Come First-Serve, which is the widely used default policy~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  %\item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
  \item LPF: Longest estimated Processing time First.
  \item LQF: Largest resource requirement $q_j$ First.
  %\item SQF: Smallest resource requirement First.
  \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
    factor is defined as follows:
  \begin{equation} \frac{start_j - r_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
  where $start_j$ is the starting time of job $j$.
  \item SEXP: Smallest Expansion Factor First
  \item LRF: Largest Ratio $\frac{p_j}{q_j}$ First
  \item SRF: Smallest Ratio First
  \item LAF: Largest Area $ p_j \times q_j$ First
  %\item SAF: Smallest Area First
\end{enumerate}

This search space is designed with the goal of being as semantically diverse as
possible without making any judgement on which policy should perform well in
practice. In the following, we denote these policies by $P_i$ with $i = 1
\ldots 10$.

\subsection{Thresholding}
As existing works point out, reordering the job queue means losing the
no-starvation guarantee and some individual jobs therefore can wait an undue
amount of time. It is possible to introduce a thresholding mechanism in order
to prevent this behavior: When a job's \textit{waiting time so far} exceeds a
fixed threshold $\Theta$, it is jumped at the head of the queue. We denote by
EASY($P,\Theta$) the scheduling policy that starts and backfill jobs according to
the (thresholded) reordering policy $P$. For the sake of completeness,
Algorithm~\ref{alg:EASY} describes the EASY($P,\Theta$) heuristic.

\begin{algorithm}[h]
  \caption{EASY($P,\Theta$) policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE Queue $Q$ of waiting jobs.
    \ENSURE None (calls to $Start()$)
    \STATE Sort $Q$ according to $P_R$
    \STATE Move all jobs of $Q$ for which $wait_j > \Theta$ ahead
    of the queue (breaking ties in FCFS order).
    \\ \textit{Starting jobs until the machine is full}
    \FOR {job $j$ in Q do}
    \IF {$j$ can be started given the current system use.}
    \STATE Pop $j$ from Q
    \STATE $Start(j)$
    \ELSE
    \STATE Reserve $j$ at the earliest
    time possible according to the estimated running times
    of the currently running jobs.
    \\ \textit{Backfilling jobs}
    \FOR {job $j'$ in $Q\setminus\{j\}$}
    \IF {$j'$ can be started without delaying the reservation on $j$.}
    \STATE Pop $j'$ from $Q$ 
    \STATE $Start(j')$
    \ENDIF
    \ENDFOR
    \STATE \textbf{break}
    \ENDIF
    \ENDFOR
  \end{algorithmic}
  \label{alg:EASY}
\end{algorithm}

\section{Online tuning}
\label{sec:online}

We present here the strategies we have retained for selecting a policy. In the remainder, we will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$. The time interval is thus divided into periods of equal length: $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period; all periods have length $\Delta$. At the beginning of each period, from the first period $\Delta_0$ to the current period $\Delta_T$, a new policy is selected. We rely here on three different strategies for this selection: complete simulation, noisy simulation, $\epsilon$-greedy bandit exploration. All these strategies are applied online.

\subsection{Online policy selection with complete simulation}
\label{sub:feedback}

Several simulators have been developed for "playing" reordering policies on a given set of jobs. Such simulations are interesting inasmuch as they provide an estimate of the cost of a given policy on a set of jobs. \textcolor{orange}{TO BE COMPLETED}

We further assume here that there is a certain regularity among periods, \textit{i.e.} that the jobs submitted between consecutive periods do not differ radically. This assumption underlies ??? and is of course all the more true than the policy period considered is not too long. It further entails that the behavior of a policy on the previous periods, in particular on the most recent ones, reflects its behavior on the current one. One can thus base the selection of a policy on its behavior on previous periods, as proposed below.

Let $l(\Delta_t)$ denote the number of jobs \textcolor{orange}{submitted?, finished?} during the period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 12$) one reordering policy (defined in Section~\ref{subsec:policies}). The cost of policy$P_i$ during the period $\Delta_t$ is defined as:
%
\[
w_t(P_i) = \sum_{j=1}^{l(\Delta_t)} Wait_j^{P_i}
\]
%
where $Wait_j^{P_i}$ denotes the waiting time for job $j$ according to policy $P_i$. The estimation of the cost of a policy over all the periods preceding the current period can then be defined as:
%
\begin{equation}
w(P_i) = \frac{1}{T} \sum_{t=0}^{T-1} (\frac{l(\Delta_t)}{\sum_{t=0}^{T-1}l(\Delta_t)})^{\alpha}  \lambda^{T-1-t} w_t(P_i)
\end{equation}
%
where $\lambda$ is a decay parameter that privileges recent periods. The fraction $l(\Delta_t)/\sum_{t=0}^{T-1}l(\Delta_t)$ allows to balance the periods according to their importance in terms of the number of jobs considered during the period, so as to penalize the influence of "small" periods wrt to "large" ones. The parameter $\alpha \in [0,1]$ serves here to control this balance (for $\alpha=0$ large and small periods are considered equal whereas for $\alpha=1$ small periods are penalized).

\subsection{Online policy selection with noisy rewards}
\label{sub:noisy}
When we have an noisy simulator.

\subsection{Bandit policy selection}
\label{sub:bandit}
When we have only bandit feedback of the accurate simulator.
We use the epsilon-greedy bandit\ref{alg:epsg}.

\begin{algorithm}[h]
  \caption{Epsilon-Greedy Bandit policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE $0 < \epsilon < 1$, $K>1$
    \FOR {$t = 1, 2, \ldots$}
    \STATE Let $i_t$ be the reordering heuristic with the highest current
    average reward.
    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
    $\epsilon$ use a random reordering heuristic.
    \ENDFOR
  \end{algorithmic}
  \label{alg:epsg}
\end{algorithm}

\section{Experiments}
\label{sec:experiments}

This section compares the different approaches to policy selection via a
comprehensive experimental campaign.  Subsection~\ref{sub:protocol} outlines
the experimental protocol used and Subsection~\ref{sub:results} contains the
experimental results and their discussion.

\subsection{Experimental Protocol}
\label{sub:protocol}

The Evaluation of computer systems performance is an intricate
task\cite{feitbook}. Here, the two main difficulties are choosing a simulation
model and taking into account the variability in the average waiting times.
This section presents our approach and provides a replicable workflow
for the interested reader.

\subsubsection{Traces}

\begin{table}[ht]
  \vspace{-0.5cm}
  \centering
  \ra{1.3}
  \caption{Workload logs used in the simulations.}
  \label{tab:logs}
  \begin{tabular}{@{}lrrrr@{}}
    \hline
    Name          & Year & \# MaxProcs & \# Jobs & Duration\\
    \hline
    KTH-SP2       & 1996 & 100         & 28k     & 11 Months\\
    CTC-SP2       & 1996 & 338         & 77k     & 11 Months\\
    SDSC-SP2      & 2000 & 128         & 59k     & 24 Months\\
    SDSC-BLUE     & 2003 & 1,152       & 243k    & 32 Months\\
    ANL-Intrepid  & 2009 & 163840      & 68k     & 9  Months\\
    CEA-Curie     & 2012 & 80,640      & 312k    & 3  Months\\
    Unilu-Gaia    & 2014 & 2,004        & 51k     & 4  Months\\
    \hline
  \end{tabular}
  \vspace{-0.5cm}
\end{table}

The experimental campaign used in this work uses a mixture of small and large
systems from different periods. Table~\ref{tab:logs} presents the 7 logs, which
can all be obtained from the Parallel Workload
Archive~\cite{Feitelson20142967}. We use the 'cleaned'\footnote{\lstinline[basicstyle=\ttfamily\color{blue}]|http://www.cs.huji.ac.il/labs/parallel/workload/logs.html#clean|} version of the logs as 
per the archive. We impose an additional
filtering step to the workloads\footnote{This filtering step is available in
the reproducible workflow\cite{repro} as shell script
\lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
clean singular data. More precisely, we apply the following modifications:

\begin{enumerate}
  \item If the number of allocated or requested cores of a job  exceeds the size of the machine, we remove the job.
  \item If the number of allocated or requested cores is negative, we use the available positive value as the request. If both are negative, we remove the job.
  \item If the runtime or submission time is negative, we remove the job.
\end{enumerate}

\subsubsection{Simulator}

The choice of a simulator is a critical part of experimental validation that
raises a tradeoff between precision and runtime. High precision can be obtained
by carefully modeling the platform and its network topology, or extracting
information from the jobs from their sources or post-morted logs. This is the
approach used by high-fidelity batch scheduling simulators such as Batsim~\cite{batsim}. In
the present work, the focus is on studying the EASY-Backfilling mechanism in
itself, without adressing the allocation problem. Moreover, the experimental
protocol used here requires many simulation runs. Therefore, we set the
precision/runtime tradeoff at the point which minimizes simulation runtimes. We
discard all topological information relative to the platform and use the job
processing times of the original workloads. In this setup, the processors are
considered to be undistinguishable from each other, and jobs can be
discontinuously mapped to any available processor on the system. We develop
a lightweight simulator\cite{ocst} and an accompanying multi-armed bandit
library\cite{obandit}. See the reproducibility paragraph below for more information.
With this simulator, we are able to replay the CEA-CURIE trace in 33 seconds on a
machine equipped with an Intel(R) Core(TM) i5-4310U CPU @ 2.00GHz, while the
same simulation with the Batsim simulator takes more than 5 hours.

\subsubsection{Resampling}

%Validating algorithms on real data is a notoriously hard
%problem~\cite{feitbook}. The case of batch scheduling is made especially
%difficult by the fluctuations in commonly used metrics~\cite{jsm}. Indeed, the
%variability in performance is known to outrank the available sample sizes:
%\textit{It is not enough to simply replay an algorithm using a trace}. For this
%reason, we clearly expose our statistical approach.

%Two main methods are available in order to generate data. On one hand, it is
%possible to model~\cite{feitperfeval} workloads. However, this often loses
%many specificities of the original data, and this work uses a second
%approach based on trace resampling~\cite{feitresampling}.

To carry our experiments, we need to generate many trace samples. More precisely, we want
to sample the job submission process, which is achieved by splitting the traces in weeks,
randomly shuffling them and picking the jobs submitted in the selected weeks.  This simple
mechanism enables us to preserve the dependency within a week. Although the system state
may not be independent from one week to an other, we can reasonably assume that it evolves
under its stationary distribution. Thus, computing an average over the sampled traces at a
given time can be related to a time average of an underlying Markov process, which
converges to the true quantity thanks to the ergodic theorem.  Moreover, the job
submission process has no long range correlation making it look like an independent and
identically distributed process when considering sufficiently time spaced weeks.

\subsubsection{Experimental campaign}

\subsubsection{Replicability}

Readers interested in replicating the experiments (or a part thereof) are
invited to peruse to the 'artifacts description' appendix of this paper.

\subsection{Results}
\label{sub:results}

\begin{figure*}[ht]
  \centering
  \includegraphics[scale=0.6]{ctc_sp.pdf}
  \includegraphics[scale=0.6]{kth_sp.pdf}\\
  \includegraphics[scale=0.6]{sdsc_s.pdf}
  \includegraphics[scale=0.6]{sdsc_b.pdf}\\
  \includegraphics[scale=0.6]{anl_in.pdf}
  \includegraphics[scale=0.6]{unilug.pdf}

  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the FullFeedback, NoisyFeedback and EpsilonGreedy
    policies. The average is obtained by resampling the original trace 100
    times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling. Each figure is a different trace, and this
figure is followed-up in figure~\ref{fig:follow} for the CEA-Curie log.}

  \label{fig:small}
\end{figure*}

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.6]{cea_cu.pdf}

  \caption{Follow-up from figure~\ref{fig:small}, plot for the CEA-Curie log.}

  \label{fig:follow}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.6]{all.pdf}
  \caption{Share of the policies chosen by the Noisy policy. The average
  choice is obtained by resampling the original trace 100 times and
  aggregating by date.}
  \label{fig:all}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.6]{mosaic_bandit.pdf}
  \caption{Share of the policies chosen by Epsilon-Greedy as a function of time.
  The average choice is obtained by resampling the original trace 100 times and
  aggregating by date.}
  \label{fig:mosb}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.6]{mosaic_noisy.pdf}
  \caption{Share of the policies chosen by the Noisy policy. The average
  choice is obtained by resampling the original trace 100 times and
  aggregating by date.}
  \label{fig:mosn}
\end{figure}

\section{Conclusion}
\label{sec:ccl}

\section*{Acknowledgments}

Authors are sorted in alphabetical order. We gracefully thank the contributors
of the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2), Travis Earheart
and Nancy Wilkins-Diehr (SDSC Blue), Lars Malinowsky (KTH SP2), Dan Dwyer and
Steve Hotovy (CTC SP2), Joseph Emeras (CEA Curie and UniLu Gaia), Susan
Coghlan, Narayan Desay, Wei Tang (ANL Intrepid), and of course Dror Feitelson.
The Metacentrum workload log was graciously provided by the Czech National Grid
Infrastructure MetaCentrum. This work has been partially supported by the LabEx
PERSYVAL-Lab (ANR-11-LABX-0025-01) funded by the French program Investissement
d'avenir. Experiments presented in this paper were carried out using the
Grid'5000 testbed. Grid'5000 is supported by a scientific interest group hosted
by Inria and including CNRS, RENATER and several Universities as well as other
organizations\footnote{https://www.grid5000.fr}.  Access to the experimental
machine(s) used in this paper was gracefully granted by research teams from
LIG\footnote{http://www.liglab.fr} and Inria\footnote{http://www.inria.fr}.

\bibliographystyle{acmst}
\bibliography{bibliography}

\clearpage
\section*{Artifacts Description}

We develop a lightweight simulator~\cite{ocst} and an accompanying multi-armed
bandit library~\cite{obandit}. Both are made available under the ISC~\cite{isc}
license and released as persistent Zenodo~\cite{zenodo} archives. Additionally,
we provide a workflow written in the zymake~\cite{breck2008zymake} format that
is able to generate all the figures from this paper and its annexes. All these
parts and their dependencies are packaged via Nix~\cite{nix} expressions.
to use the
archive at~\cite{repro} on any linux system equipped with the Nix package manager.
Generating the figures is done with:

\begin{lstlisting}
wget ".." -o banditeasy.tgz
tar -vzf banditeasy.tgz
cd banditeasy
nix-build
\end{lstlisting}

\end{document}

### 2017-03-29 paper/paper.tex
paper update

@@
 \usepackage{tikz}
-\usepackage{todo}
 \usepackage{graphicx}
@@
 
-Several works use this method in the context of HPC, in
+Several works take a similar approach in the context of HPC, in
 particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
@@
 
-While these previous works intend to estimate uncertain parameters, we consider in this paper a more
-pragmatic approach, which directly learn a good scheduling policy from a given policy space.
+A recent work~\cite{mlmem} proposes to help
+users specify better memory requirements by applying supervised learning
+techniques to job metadata.
+
+While these previous works intend to estimate uncertain parameters, they lack
+in validating the impact on the system. We consider in this paper a more
+pragmatic approach, which is to directly learn a scheduling policy from a given
+seach space.
 
@@
 This section presents the systems under study and the scheduling problem.  It
-first introduces the EASY-Backfilling heuristic and gives \todo{fix} the
-problem statement.
+first introduces the EASY-Backfilling heuristic and gives
+\textcolor{orange}{(fix ugly sentence)} the problem statement.
 
@@
 Our problem statement is: \textbf{How to tune EASY-Backfilling in an online manner?}
-\todo{introduce the difference in workloads, the duality of avg and max, and the online problem}
+\textcolor{orange}{introduce the difference in workloads, the duality of avg and max, and the online problem}
 
@@
 \subsection{Reordering the job Queue}
+\label{subsec:policies}
 
@@
 \subsection{Thresholding}
+\label{sub:th}
 As existing works point out, reordering the job queue means losing the
@@
 \begin{table}[ht]
-  \vspace{-0.5cm}
   \centering
@@
   \end{tabular}
-  \vspace{-0.5cm}
 \end{table}
@@
 information from the jobs from their sources or post-morted logs. This is the
-approach used by high-fidelity batch scheduling simulators such as Batsim~\cite{batsim}. In
-the present work, the focus is on studying the EASY-Backfilling mechanism in
-itself, without adressing the allocation problem. Moreover, the experimental
-protocol used here requires many simulation runs. Therefore, we set the
-precision/runtime tradeoff at the point which minimizes simulation runtimes. We
-discard all topological information relative to the platform and use the job
-processing times of the original workloads. In this setup, the processors are
-considered to be undistinguishable from each other, and jobs can be
-discontinuously mapped to any available processor on the system. We develop
-a lightweight simulator\cite{ocst} and an accompanying multi-armed bandit
-library\cite{obandit}. See the reproducibility paragraph below for more information.
-With this simulator, we are able to replay the CEA-CURIE trace in 33 seconds on a
-machine equipped with an Intel(R) Core(TM) i5-4310U CPU @ 2.00GHz, while the
-same simulation with the Batsim simulator takes more than 5 hours.
+approach used by high-fidelity batch scheduling simulators such as
+Batsim~\cite{batsim}. In the present work, the focus is on studying the
+EASY-Backfilling mechanism in itself, without adressing the allocation problem.
+Moreover, the experimental protocol used here requires many simulation runs.
+Therefore, we set the precision/runtime tradeoff at the point which minimizes
+simulation runtimes. We discard all topological information relative to the
+platform and use the job processing times of the original workloads. In this
+setup, the processors are considered to be undistinguishable from each other,
+and jobs can be discontinuously mapped to any available processor on the
+system. We develop a lightweight simulator\cite{ocst} and an accompanying
+multi-armed bandit library\cite{obandit}. See the reproducibility paragraph
+below for more information.  With this simulator, we are able to replay
+EASY-FCFS on the CEA-CURIE trace in 33 seconds with a machine equipped with an
+Intel(R) Core(TM) i5-4310U CPU @ 2.00GHz. As a point of comparison, the Batsim
+simulator with no network communication modeling and no resource contiguity
+takes more than 7 hours to replay the same trace on the same hardware. The complete
+simulation campaign presented below is executed on a single Dell PowerEdge R730 in 22
+hours, including input preparation and analysis code.
 
@@
 Readers interested in replicating the experiments (or a part thereof) are
-invited to peruse to the 'artifacts description' appendix of this paper.
+invited to peruse to the 'artifacts description' appendix of this paper.  We
+provide to persistent archives containing the simulator used, the analysis code
+and the associated workflow necessary to replicate all experiments and figures
+present in this paper. All these parts's dependencies are automatically managed
+via Nix~\cite{nix} and we invite the reader to replicate the results.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
 
-\section*{Acknowledgments}
+\begin{acks}
 
@@
 Infrastructure MetaCentrum. This work has been partially supported by the LabEx
-PERSYVAL-Lab (ANR-11-LABX-0025-01) funded by the French program Investissement
-d'avenir. Experiments presented in this paper were carried out using the
-Grid'5000 testbed. Grid'5000 is supported by a scientific interest group hosted
-by Inria and including CNRS, RENATER and several Universities as well as other
+PERSYVAL-Lab(\grantsponsor{ANR-11-LABX-0025-01}{ANR-11-LABX-0025-01}) funded by
+the French program Investissement d'avenir. Experiments presented in this paper
+were carried out using the Grid'5000 testbed. Grid'5000 is supported by a
+scientific interest group hosted by Inria and including CNRS, RENATER and
+several Universities as well as other
 organizations\footnote{https://www.grid5000.fr}.  Access to the experimental
@@
 LIG\footnote{http://www.liglab.fr} and Inria\footnote{http://www.inria.fr}.
+
+\end{acks}
 

### 2017-03-29 paper/paper.tex
update paper figures

@@
 the interface of HPC and bigdata, dealing with learning algorithms
-(including the authors of this paper~\cite{Gaussier2015ImprovingBB}).
+(including the authors of this paper~\cite{learningruntimes}).
 Most of these studies target the process to determine more or less accurately
@@
 \subsubsection{Traces}
+\label{sub:traces}
 
@@
 
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[scale=0.6]{variability.pdf}
+  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (pre-processing described in subsection~\ref{sub:traces}) for a few policies.}
+  \label{fig:mosn}
+\end{figure*}
+
 %Validating algorithms on real data is a notoriously hard
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[h]
   \centering
-  \includegraphics[scale=0.6]{ctc_sp.pdf}
-  \includegraphics[scale=0.6]{kth_sp.pdf}\\
-  \includegraphics[scale=0.6]{sdsc_s.pdf}
-  \includegraphics[scale=0.6]{sdsc_b.pdf}\\
-  \includegraphics[scale=0.6]{anl_in.pdf}
-  \includegraphics[scale=0.6]{unilug.pdf}
+  \includegraphics[scale=0.6]{full-CTC-SP2.pdf}
+  \includegraphics[scale=0.6]{full-KTH-SP2.pdf}\\
+  \includegraphics[scale=0.6]{full-CEA-Curi.pdf}
+  \includegraphics[scale=0.6]{full-SDSC-BLU.pdf}\\
+  \includegraphics[scale=0.6]{full-SDSC-SP2.pdf}
+  \includegraphics[scale=0.6]{full-ANL-Intr.pdf}
 
@@
   values across this resampling. Each figure is a different trace, and this
-figure is followed-up in figure~\ref{fig:follow} for the CEA-Curie log.}
+figure is followed-up in figure~\ref{fig:follow} for the UniLu-Gaia log.}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
   \centering
-  \includegraphics[scale=0.6]{cea_cu.pdf}
+  \includegraphics[scale=0.6]{full-UniLu-Ga.pdf}
 
-  \caption{Follow-up from figure~\ref{fig:small}, plot for the CEA-Curie log.}
+  \caption{Follow-up from figure~\ref{fig:small}, plot for the UniLu-Gaia log.}
 
@@
   \centering
-  \includegraphics[scale=0.6]{all.pdf}
-  \caption{Share of the policies chosen by the Noisy policy. The average
-  choice is obtained by resampling the original trace 100 times and
-  aggregating by date.}
+  \includegraphics[scale=0.6]{UniLu-Ga.pdf}
+
+  \caption{Evolution of the average cumulative waiting time improvement
+    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$. The
+    average is obtained by resampling the original trace 100 times. The dashed
+    lines represent the 10th and 90th percentiles of the values across this
+  resampling. Each figure is a different trace, and this figure is followed-up
+in figure~\ref{fig:follow} for the UniLu-Gaia log.}
+
   \label{fig:all}
@@
   \centering
-  \includegraphics[scale=0.6]{mosaic_bandit.pdf}
+  \includegraphics[scale=0.6]{mosaicbandit-UniLu-Ga.pdf}
   \caption{Share of the policies chosen by Epsilon-Greedy as a function of time.
@@
   \centering
-  \includegraphics[scale=0.6]{mosaic_noisy.pdf}
+  \includegraphics[scale=0.6]{mosaic-UniLu-Ga.pdf}
   \caption{Share of the policies chosen by the Noisy policy. The average

### 2017-03-29 paper/paper.tex
fix
Draft Artifact section

@@
 \subsubsection{Resampling}
+\label{ssub:resampling}
 
@@
     lines represent the 10th and 90th percentiles of the values across this
-  resampling. Each figure is a different trace, and this figure is followed-up
-in figure~\ref{fig:follow} for the UniLu-Gaia log.}
+  resampling.}
 
@@
 
-Authors are sorted in alphabetical order. We gracefully thank the contributors
+Authors are sorted in alphabetical order. We warmly thank Pierre Neyron for help with the experiments. We gracefully thank the contributors
 of the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2), Travis Earheart
@@
 
-We develop a lightweight simulator~\cite{ocst} and an accompanying multi-armed
-bandit library~\cite{obandit}. Both are made available under the ISC~\cite{isc}
-license and released as persistent Zenodo~\cite{zenodo} archives. Additionally,
-we provide a workflow written in the zymake~\cite{breck2008zymake} format that
-is able to generate all the figures from this paper and its annexes. All these
-parts and their dependencies are packaged via Nix~\cite{nix} expressions.
-to use the
-archive at~\cite{repro} on any linux system equipped with the Nix package manager.
-Generating the figures is done with:
+In this annex, we describe the computational artifacts associated with this
+work. We take a lightweight approach to replicability by packaging a workflow
+that generates all the figures from this paper and making it available
+publicly. Moreover, questions regarding replication can be freely directed at
+the authors via e-mail.
+
+All the code associated with this work is released under the ~\cite{ISC}
+license as a persistent Zenodo~\cite{zenodo} archive at~\cite{zenodomain}.
+
+\subsection{Packaging and Replicability}
+
+There are various approaches for making computational experiments replicable,
+among which distributing complete operating system images, containers, or
+packaging software. As our experiments are CPU bound, we decide to opt for the
+software packaging approach using the Nix~\cite{nix} packaging system.  All the
+dependencies of our experiments (including our own code, dependencies for data
+processing and visualization and workflow engine) are automatically managed by
+Nix up to the actual execution of the code.
+
+Running the experiments can be done on any platform equipped with the nix
+workflow system by running:
 
 \begin{lstlisting}
-wget ".." -o banditeasy.tgz
-tar -vzf banditeasy.tgz
-cd banditeasy
-nix-build
+nix-build (fetchTarball ...) -A banditSelection
 \end{lstlisting}
 
+\subsection{Testbed}
+
+ The experiments in this paper can run via in less than 24 hours on a large
+ machine. More precisely, we had access to a Dell PowerEdge R730 equipped with
+ a total of 56 threads @2.4GHz each and 757G of RAM.  For this reason, the
+ experimental campaign was designed to be executed on a single host.
+
+ \subsection{Workflow}
+
+ All experiments are tied together using Zymake~\cite{zymake}, a minimalistic
+ workflow system designed for computational experiments. This system is
+ analogous to a traditional build system with added workflow capabilities. The
+ entire workflow that generates this article from the input data is contained
+ in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
+ found at the root of the main archive. This workfloww is composed of the
+ following steps:
+
+ \begin{itemize}
+
+   \item[Preprocessing]: Data extraction from archives, filtering, resampling.
+     This is principally using shell scripts for data manipulation (see file
+     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter| for the
+     filtering steps used) and ocaml code that implements the resampling method
+     described in~\ref{sub:resampling}.
+
+   \item[Simulation]: This step runs the lightweight ocaml backfilling
+     simulator specially written for this work. This simulator is made
+     available under the ISC~\cite{ISC} license both as a persistent Zenodo
+     archive at~\cite{ocst} and as a git repository at ~\cite{ocstgit}.
+
+   \item[Analysis]: This step runs mostly R code that generates the figures
+     presented in this paper.
+ \end{itemize}
 \end{document}

### 2017-03-30 paper/paper.tex
incorporate nightly figures.
denis changes

@@
 
-Plan :
-\begin{itemize}
-\item Context - message : les machines sont de + en + complexes, les systemes
-de management restent rudimentaires face a la complexite croissante de
-l'architecture et des applications.
-\item les habitudes et usages existants : justification du FCFS-EASY-BF,
-developper rapidement ce qui existe
-\item ce que l'on propose
-\end{itemize}
+\subsection{Context}
 
-Providing the computing infrastuctures needed to solve actual complex problems
-arising in the various fields of the modern society (including climate change,
-health, green energy or security) is a strategic challenge.
-In particular, High Performance Computing (HPC) systems are evolving to
-extreme-scale parallel and distributed platforms.
-The race for always more computing power and storage capacity does not only
-lead to sophisticated specific exascale platforms, but the objective of the
-community is to design efficient sustained Petascale platforms.
-However, there is still a long way to go and many scientific and technical
-problems to solve by the system management software in order to adapt to such
-large-scale evolution and the increasing complexity
+Providing the computing infrastuctures needed to solve actual complex 
+problems arising in the various fields of the modern society (including 
+climate change, health, green energy or security) is a strategic challenge.
+The main pillar to address this challenge is to build extreme-scale 
+parallel and distributed platforms.
+The race for always more computing power and storage capacity does not 
+only lead to sophisticated specific exascale platforms, but the 
+objective of the community is to design efficient sustained Petascale 
+platforms.
+However, there is still a long way to go and many scientific and 
+technical problems to solve by the system management software in order 
+to adapt to such large-scale evolution and the increasing complexity
 in both architecture and applications.
-The existing job and resource management sofwares allow to run tens thousands
-jobs on hundreds thousands cores.
-They are based on robust and rather simple policies.
+The existing job and resource management softwares allow to run tens 
+thousands jobs on hundreds thousands cores.
+They are based on robust policies, however, despite their positive 
+properties of relative simplicity and their guarantee of no job starvation,
+there is still place for improvements.
+We propose here to study how to tune the classical EASY-BackFilling 
+heuristic using two new learning approaches (namely, .
 
@@
 The management generates a huge amount of data.
-During the last few years, there was an explosion of the number of of works at
-the interface of HPC and bigdata, dealing with learning algorithms
-(including the authors of this paper~\cite{learningruntimes}).
-Most of these studies target the process to determine more or less accurately
-the value of some specific key parameters, with the idea of
-learning better estimates should improve the performances of the resource
-manager.
+During the last few years, there was an explosion of the number of works 
+at the interface of HPC and BigData, dealing with learning algorithms
+(including the authors of this work~\cite{SC2015}).
+Most of these studies target the process of determining more or less 
+accurately the value of some specific key parameters, with the idea that
+learning better estimates should improve the performances of the 
+resource manager.
 
 \bigskip
-Our idea within this work is to learn how the management system behave within
-the classical framework of of selecting jobs of the submission queue under
-EASY-backfilling.
-Contrarily to most existing studies which consider only one execution trace, we
-extend the learning on a multiple trace basis.
-Indeed, learning for a specific trace leads to limited results, customized for
-one trace on a given platform (what happens if we target another platform?).
-Moreover, this is usually inefficient to deal with extreme unpredictable
-events.
+Our focus within this work is to learn how the management system behaves 
+within the usual framework of selecting jobs of the submission queue 
+under EASY-BF.
+Contrarily to most existing studies, which consider only one execution 
+trace, we extend the learning on a multiple traces basis.
+We built a simulator which generates a huge number of partial executions 
+coming from multiple users collected in various actual traces and HPC 
+platforms.
+The sampling is explained in detail later in the paper.
+This was motivated because learning for a specific trace/platform leads 
+to limited results, customized for one particular instance (then, what 
+happens if we target another platform?).
+The collected data are too small and moreover, this approach is usually 
+inefficient to deal with extreme unpredictable events.
+The sampling generation allow to smooth the data.
+
+%On se place du point de vue de l'admin system...
 
 \subsection{Contributions}
+
+Our main contribution is to investigate a new way for learning, which 
+consists in using of a simulator to generate a large data set extracted 
+from various actual traces.
+We propose two variants of the approach: the ideal situation which 
+relies on the  precise knowledge of data, and the more realistic 
+assumption of noisy data (which better reflect the uncertainties).
+We also propose a new on-line approach based on multi-armed Bandit to 
+learn very fast how to prioritize the submission and BackFilling queues.
+\bigskip
+
+Our main results are:
+
+l'approche basee sur le simulateur marche bien. meme avec des donnes 
+imprecises.
+
+Le bandit a des resultats raisonnables, il ne necessite que peu de code 
+supplementaire...
+
+Dans les details :
+on montre que une gestion aleatoire des queues est meilleure que FCFS, 
+et on identifie des politiques qui marchent bien dans plusieurs traces, 
+etc..
 
@@
 
-This section presents related works in the area of scheduling for batch scheduling.
+This section presents and discuss the most significant results related 
+to job scheduling and learning algorithms in this context.
+Let us start by recalling some basics about scheduling heuristics in HPC 
+platforms:
 
-\subsection{Scheduling heuristics in HPC platforms}
+While parallel job scheduling is a old studied theoretical
+problem~\cite{leung2004handbook} \texbf{PREFER DROZDOWSKI HERE + SURVEY 
+FEITELSON-RUDOLPH-SWIEGELSHON}, the practical ramifications, varying
+hypotheses, and inherent uncertainty of the problem applied on the HPC 
+field have driven
+practitioners to use \textit{simple} heuristics (and researchers to 
+study their behavior).
+The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
+Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.
 
-While parallel job scheduling is a well studied theoretical
-problem~\cite{leung2004handbook}, the practical ramifications, varying
-hypotheses, and inherent uncertainty of the problem in HPC have driven
-practitioners and researchers alike to use and study simple heuristics.
-The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
-Conservative~\cite{Mu'alem:2001:UPW:380314.380315} Backfilling.
-
-While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
-significant computational overhead, perhaps explaining why most of the machines
-of the top500 ranking~\cite{top500} still use at the time of this publication a
-variant of EASY Backfilling.
-
-\subsection{EASY}
-There is a large body of work seeking to improve EASY. Indeed, while the
-heuristic is used by various resource and job management softwares (most notably
-SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by system
+While Conservative Backfilling offers many advantages~\cite{bfchar}, it 
+has a
+significant computational overhead, which mainly explains why most of 
+the machines
+of the top500 ranking~\cite{top500} rather use a variant of EASY-BF 
+instead.
+There is a large body of work seeking to improve/tune EASY. Indeed, 
+while the
+basic mechanism is used by some actual resource and job management 
+softwares (most notably
+SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by 
+system
 administrators.
 
+The original EASY mechanism refers to a First-Come-First-Serve basis.
 Several works explore how to tune EASY by reordering waiting
@@
 randomized manner~\cite{1592720}, as well as some
-implementations~\cite{Jackson2001}. However, as successful as they may be, these works
-do not address the dependency~\cite{variability} of scheduling metrics on the
-workload. Indeed these studies most often report \textit{post-hoc} performance since
+implementations~\cite{Jackson2001}. However, as successful as they may 
+be, these works
+do not address the dependency of scheduling metrics on the
+workload~\cite{variability}. Indeed these studies most often report 
+\textit{post-hoc} performance since
 they compare algorithms after the workload is known.
@@
 The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
-method to tuning these queues, although it requires simulated scheduling runs
-at decision time and therefore costs much more than the natural execution of EASY.
+method to tuning these queues, although it requires simulated scheduling 
+runs
+at decision time and therefore, it costs much more than the natural 
+execution of EASY.
 
-\subsection{Data-aware resource management}
+\paragraph
+\textbf{Data-aware resource management.}
 
-There is a recent focus on leveraging the high amount of data available in
+There was a recent focus on leveraging the huge amount of data available in
 large scale computing systems in order to improve their behavior. Some works
 use collaborative filtering to colocate tasks in clouds by estimating
-application interference~\cite{7516031}.  Others are closer to the application
+application interference~\cite{7516031}.  Some others are closer to the 
+application
 level and use binary classification to distinguish benign memory faults from
-application errors in order to execute recovery algorithms (see
-~\cite{fmodeling} for instance).
+application errors in order to execute recovery algorithms 
+(see~\cite{fmodeling} for instance).
 
-Several works take a similar approach in the context of HPC, in
-particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
+Several works use \textbf{this method} (LAQUELLE?) in the context of HPC, in
+particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that 
+better job
 runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
@@
 probabilistic integration procedures~\cite{Nissimov2008}.
+However, these works do not address the duality between the cumulative and
+maximal scheduling costs, as mentioned in~\cite{learningruntimes}.
+While these previous works intend to estimate uncertain parameters, we 
+consider here a more
+pragmatic approach, which directly learn a good scheduling policy from a 
+given policy space.
 
-However, these works do not address the duality between the cumulative and
-maximal scheduling costs, as mentionned in~\cite{learningruntimes}.
-
-A recent work~\cite{mlmem} proposes to help
-users specify better memory requirements by applying supervised learning
-techniques to job metadata.
-
-While these previous works intend to estimate uncertain parameters, they lack
-in validating the impact on the system. We consider in this paper a more
-pragmatic approach, which is to directly learn a scheduling policy from a given
-seach space.
-
-\subsection{Multi-Armed Bandits}
+\paragraph
+\textbf{Multi-Armed Bandits.}
 
 A multi-armed bandit (MAB) problem is a sequential allocation problem with
-partially observable rewards. At every round, an action (an arm) must be chosen
+partially observable rewards. At every round, an action (represented by 
+an arm of the Bandit) must be chosen
 in a fixed set and the corresponding reward is observed. The goal of a MAB
-algorithm is to maximize the total reward obtained in a number of rounds.
-There are a number of works that address this problem under a variety of
-constraints, The two most popular settings are the original stochastic  case,
-and the adversarial case. See ~\cite{thompson} for the original work on the
-stochastic case and ~\cite{Auer2002} for the UCB family of algorithms. The work
-of ~\cite{Banos} is the earliest known work to us for the adversarial case
+algorithm is to maximize the total reward obtained in a successive 
+number of rounds.
+There exist a bunch of works that address this problem under a variety of
+constraints, The two most popular settings are the original stochastic  case
+and the adversarial case. See~\cite{thompson} for the original work on the
+stochastic case and~\cite{Auer2002} for the UCB family of algorithms 
+\textbf{EXPLAIN HERE THE ACRONYM}.
+According to our knowledge, the work
+of~\cite{Banos} is the earliest known work to us for the adversarial case
 see~\cite{nonstoch} for the EXP family of algorithms. We refer to the
@@
 algorithms bound the cumulative difference in loss to the best arm (the
-regret), they have functional constraits such as the fact that the rewards
-should be contained in a range. The\cite{Auer2002} heuristic called
-Epsilon-Greedy is known to achieve good results in practice, and does not have
+regret), they have functional constraints such as the fact that the rewards
+should be contained in a range. The heuristic called
+\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to 
+achieve good results in practice, and does not have
 this requirement.
@@
   \item LCFS: Last-Come First-Serve.
-  %\item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
+  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
   \item LPF: Longest estimated Processing time First.
   \item LQF: Largest resource requirement $q_j$ First.
-  %\item SQF: Smallest resource requirement First.
+  \item SQF: Smallest resource requirement First.
   \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
@@
   \item LAF: Largest Area $ p_j \times q_j$ First
-  %\item SAF: Smallest Area First
+  \item SAF: Smallest Area First
 \end{enumerate}
@@
   \includegraphics[scale=0.6]{variability.pdf}
-  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (pre-processing described in subsection~\ref{sub:traces}) for a few policies.}
-  \label{fig:mosn}
+  \caption{Variability in the weekly average waiting time of EASY($P,20h$) using the KTH-SP2 trace (pre-processing described in subsection~\ref{sub:traces}) for
+  a few select policies from~\ref{sub:th}.}
+  \label{fig:variability}
 \end{figure*}
@@
 invited to peruse to the 'artifacts description' appendix of this paper.  We
-provide to persistent archives containing the simulator used, the analysis code
-and the associated workflow necessary to replicate all experiments and figures
-present in this paper. All these parts's dependencies are automatically managed
-via Nix~\cite{nix} and we invite the reader to replicate the results.
+provide to persistent archives containing the simulator, the analysis code
+and an automated workflow that replicates all experiments and figures
+present in this paper.
 
@@
   \centering
-  \includegraphics[scale=0.6]{full-CTC-SP2.pdf}
-  \includegraphics[scale=0.6]{full-KTH-SP2.pdf}\\
-  \includegraphics[scale=0.6]{full-CEA-Curi.pdf}
-  \includegraphics[scale=0.6]{full-SDSC-BLU.pdf}\\
-  \includegraphics[scale=0.6]{full-SDSC-SP2.pdf}
-  \includegraphics[scale=0.6]{full-ANL-Intr.pdf}
+  \includegraphics[scale=0.6]{CTC-SP2.pdf}
+  \includegraphics[scale=0.6]{KTH-SP2.pdf}\\
+  \includegraphics[scale=0.6]{CEA-Curi.pdf}
+  \includegraphics[scale=0.6]{SDSC-BLU.pdf}\\
+  \includegraphics[scale=0.6]{SDSC-SP2.pdf}
+  \includegraphics[scale=0.6]{ANL-Intr.pdf}
 
@@
   \centering
-  \includegraphics[scale=0.6]{full-UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{UniLu-Ga.pdf}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
   \centering
-  \includegraphics[scale=0.6]{UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{full-UniLu-Ga.pdf}
 
   \caption{Evolution of the average cumulative waiting time improvement
-    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$. The
+    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
+    UniLu-Gaia trace. The
     average is obtained by resampling the original trace 100 times. The dashed
@@
 \clearpage
-\section*{Artifacts Description}
 
-In this annex, we describe the computational artifacts associated with this
-work. We take a lightweight approach to replicability by packaging a workflow
-that generates all the figures from this paper and making it available
+\section*{Artifacts Description Appendix}
+
+This appendix describes the computational artifacts associated with this work. We
+take a lightweight approach to replicability by packaging a declarative
+workflow that generates all the figures from this paper and making it available
 publicly. Moreover, questions regarding replication can be freely directed at
@@
 packaging software. As our experiments are CPU bound, we decide to opt for the
-software packaging approach using the Nix~\cite{nix} packaging system.  All the
-dependencies of our experiments (including our own code, dependencies for data
-processing and visualization and workflow engine) are automatically managed by
-Nix up to the actual execution of the code.
+software packaging approach using the Nix~\cite{nix} packaging system. This
+allows for replicating the experiments on clusters where virtualization is not
+available or kernels are old and decreases runtime. All the dependencies of our
+experiments (including our own code, dependencies for data processing and
+visualization and workflow engine) are automatically managed by Nix up to the
+actual execution of the code.
 
 Running the experiments can be done on any platform equipped with the nix
-workflow system by running:
+package manager by running:
 
 \begin{lstlisting}
-nix-build (fetchTarball ...) -A banditSelection
+nix-build (fetchTarball ...)
 \end{lstlisting}
@@
 
- The experiments in this paper can run via in less than 24 hours on a large
- machine. More precisely, we had access to a Dell PowerEdge R730 equipped with
+ The experiments in this paper can run via in less than 24 hours on a moderately parallel
+ host. More precisely, we had access to a Dell PowerEdge R730 equipped with
  a total of 56 threads @2.4GHz each and 757G of RAM. For this reason, the
- experimental campaign was designed to be executed on a single host.
+ experimental campaign workflow was designed to be executed on a single host.
 
@@
  All experiments are tied together using Zymake~\cite{zymake}, a minimalistic
- workflow system designed for computational experiments. This system is
+ workflow system designed for computational experiments\footnote{The zymake system
+ is also packaged by our Nix expressions.}. This system is
  analogous to a traditional build system with added workflow capabilities. The
@@
  in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
- found at the root of the main archive. This workfloww is composed of the
+ found at the root of the main archive. This workflow is composed of the
  following steps:
@@
  \begin{itemize}
-
    \item[Preprocessing]: Data extraction from archives, filtering, resampling.
@@
      filtering steps used) and ocaml code that implements the resampling method
-     described in~\ref{sub:resampling}.
+     described in~\ref{ssub:resampling}.
 
@@
 
-   \item[Analysis]: This step runs mostly R code that generates the figures
+   \item[Analysis]: This step runs R code that generates the figures
      presented in this paper.

### 2017-03-30 paper/paper.tex
fix denischanges

@@
 While parallel job scheduling is a old studied theoretical
-problem~\cite{leung2004handbook} \texbf{PREFER DROZDOWSKI HERE + SURVEY 
+problem~\cite{leung2004handbook} \textbf{PREFER DROZDOWSKI HERE + SURVEY 
 FEITELSON-RUDOLPH-SWIEGELSHON}, the practical ramifications, varying

### 2017-03-30 paper/paper.tex
all

--- 2017-03-29 paper/paper.tex
@@
 At the beginning of each period, from the first period $\Delta_0$ to the current period $\Delta_T$, a new policy is selected.
-We rely here on three different strategies for this selection: complete simulation, noisy simulation, $\epsilon$-greedy bandit exploration.
-All these strategies are applied online.
 
-\subsection{Online policy selection with complete simulation}
+We assume here that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods.
+This assumption is validated in the study presented in~\ref{???}.
+It further entails that the behavior of a policy on the previous periods reflects its behavior on the current one, so that the selection of a policy can be based on its behavior on previous periods.
+However, there usually is an important variability on the cost metric between different periods, for any policy.
+This is illustrated in Figure~\ref{fig:variability} where \textcolor{orange}{???}.
+When the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
+
+The selection of a policy is of course reminiscent of reinforcement learning.
+It is important to note, however, that a pure reinforcement learning approach is difficult to develop in our context as \textcolor{orange}{???}.
+We rely here on simpler, yet we believe more effective, strategies to solve this problem.
+These strategies are applied online and rely on exact simulation, noisy simulation and $\epsilon$-greedy bandit exploration.
+
+\subsection{Online policy selection with exact simulation}
 \label{sub:feedback}
@@
 
-We further assume here that there is a certain regularity among periods, \textit{i.e.} that the jobs submitted between consecutive periods do not differ radically.
-This assumption underlies ???
-and is of course all the more true than the policy period considered is not too long.
-It further entails that the behavior of a policy on the previous periods, in particular on the most recent ones, reflects its behavior on the current one.
-One can thus base the selection of a policy on its behavior on previous periods, as proposed below.
-
-Let $l(\Delta_t)$ denote the number of jobs \textcolor{orange}{submitted?, finished?} during the period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 12$) one reordering policy (defined in Section~\ref{subsec:policies}).
-The cost of policy$P_i$ during the period $\Delta_t$ is defined as:
-%
-\[
-w_t(P_i) = \sum_{j=1}^{l(\Delta_t)} Wait_j^{P_i}
-\]
-%
-where $Wait_j^{P_i}$ denotes the waiting time for job $j$ according to policy $P_i$. The estimation of the cost of a policy over all the periods preceding the current period can then be defined as:
+Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 10$) one reordering policy (defined in Section~\ref{subsec:policies}).
+The cost of policy $P_i$ during the period $\Delta_t$ is defined as:
 %
 \begin{equation}
-w(P_i) = \frac{1}{T} \sum_{t=0}^{T-1} (\frac{l(\Delta_t)}{\sum_{t=0}^{T-1}l(\Delta_t)})^{\alpha}  \lambda^{T-1-t} w_t(P_i)
+\label{eq:cost-exact}
+w_{\mbox{exact}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \mbox{Wait}_j^{P_i}
 \end{equation}
 %
-where $\lambda$ is a decay parameter that privileges recent periods.
-The fraction $l(\Delta_t)/\sum_{t=0}^{T-1}l(\Delta_t)$ allows to balance the periods according to their importance in terms of the number of jobs considered during the period, so as to penalize the influence of "small" periods wrt to "large" ones.
-The parameter $\alpha \in [0,1]$ serves here to control this balance (for $\alpha=0$ large and small periods are considered equal whereas for $\alpha=1$ small periods are penalized).
+where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of the waiting time for job $j$ according to policy $P_i$.
+The estimation of the cost of a policy over all the periods preceding the current period can then be defined as:
+%
+\begin{equation}
+\label{eq:tot-cost-exact}
+w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{exact}}(t;P_i)
+\end{equation}
+%
+$\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent history (\textit{i.e.} recent periods).
+Note that the above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.
 
-\subsection{Online policy selection with noisy rewards}
+In the context of this study, as detailed in Section~\ref{sec:experiments}, we rely on metrics averaged over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies.
+Such traces are typically obtained by simulation.
+Using the same simulator for generating the traces and estimating the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact} would however be too optimistic and would represent an upper bound on what can be achieved by a selection strategy based on simulation.
+
+In order to have a more realistic estimate of the behavior of a selection strategy based on simulation, we introduce noise in the simulator, as described below.
+
+\subsection{Online policy selection with noisy simulation}
 \label{sub:noisy}
-When we have an noisy simulator.
+
+In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact}
 
@@
   \centering
-  \includegraphics[scale=0.6]{variability.pdf}
+  \includegraphics[scale=0.6]{figures/variability.pdf}
   \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (pre-processing described in subsection~\ref{sub:traces}) for a few policies.}
@@
   \centering
-  \includegraphics[scale=0.6]{full-CTC-SP2.pdf}
-  \includegraphics[scale=0.6]{full-KTH-SP2.pdf}\\
-  \includegraphics[scale=0.6]{full-CEA-Curi.pdf}
-  \includegraphics[scale=0.6]{full-SDSC-BLU.pdf}\\
-  \includegraphics[scale=0.6]{full-SDSC-SP2.pdf}
-  \includegraphics[scale=0.6]{full-ANL-Intr.pdf}
+  \includegraphics[scale=0.6]{figures/CTC-SP2.pdf}
+  \includegraphics[scale=0.6]{figures/KTH-SP2.pdf}\\
+  \includegraphics[scale=0.6]{figures/CEA-Curi.pdf}
+  \includegraphics[scale=0.6]{figures/SDSC-BLU.pdf}\\
+  \includegraphics[scale=0.6]{figures/SDSC-SP2.pdf}
+  \includegraphics[scale=0.6]{figures/ANL-Intr.pdf}
 
@@
   \centering
-  \includegraphics[scale=0.6]{full-UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{figures/UniLu-Ga.pdf}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
   \centering
-  \includegraphics[scale=0.6]{UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{figures/full-UniLu-Ga.pdf}
 
   \caption{Evolution of the average cumulative waiting time improvement
-    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$. The
+    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
+    UniLu-Gaia trace. The
     average is obtained by resampling the original trace 100 times. The dashed
@@
   \centering
-  \includegraphics[scale=0.6]{mosaicbandit-UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}
   \caption{Share of the policies chosen by Epsilon-Greedy as a function of time.
@@
   \centering
-  \includegraphics[scale=0.6]{mosaic-UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}
   \caption{Share of the policies chosen by the Noisy policy. The average

### 2017-03-30 paper/paper.tex
all

@@
 
-In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact}
+In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling, either dow or up, the waiting time of jobs obtained by simulation:
+%
+\begin{equation}
+\label{eq:cost-exact}
+w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho \mbox{Wait}_j^{P_i}
+\end{equation}
+%
 

### 2017-03-30 paper/paper.tex
all

--- 2017-03-30 paper/paper.tex
@@
 
-In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact}
+In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either down or up, for each job:
+%
+\begin{enumerate}
+\item For each job $j$ submitted during $\Delta_t$: $\rho_j \sim U[0.85,1.15]$
+\item Then:
+%
+\begin{equation}
+\label{eq:cost-noisy}
+w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i}
+\end{equation}
+%
+\end{enumerate}
+%
+$\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a +/-15\% noise on the waiting time estimated by the simulator.
+The overall cost is then defined in the same way as before, leading to:
+%
+\begin{equation}
+\label{eq:tot-cost-noisy}
+w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i)
+\end{equation}
+%
+Instead of introducing noise in the output of the simulator when selecting the policy, one could have introduced it when generating the execution traces.
+This would however have prevented us from comparing the exact and noisy selection methods, the first one representing an upper bound on the second one.
+As we will see in Section~\ref{sec:experiments}, there is little difference between the two selection methods.
 
@@
 \label{sub:bandit}
+
 When we have only bandit feedback of the accurate simulator.

### 2017-03-30 paper/run-latex.sh
arch change. pdflatex works in src/paper/

@@ -1,0 +1,166 @@
+source $stdenv/setup
+
+mkdir -p $out
+
+export VARTEXFONTS=$TMPDIR/texfonts
+export TEXMFCNF=$TMPDIR:
+echo 'max_print_line = 8192' >> $TMPDIR/texmf.cnf
+
+mkdir root
+cd root
+
+startDir=$(perl $copyIncludes $includes)
+cd $startDir
+
+for i in $extraFiles; do
+  if test -d $i; then
+      ln -sf $i/* .
+  else
+      ln -sf $i $(stripHash $i)
+  fi
+done
+
+rootName=$(basename $(stripHash "$rootFile"))
+
+rootNameBase=$(echo "$rootName" | sed 's/\..*//')
+
+if test -n "$generatePDF"; then
+    latex=pdflatex
+else
+    latex=latex
+fi
+
+latexFlags="-file-line-error"
+tmpFile=$out/log
+
+showError() {
+    echo
+    echo "LATEX ERROR (LAST LOG LINES SHOWN):"
+    tail -n 20 $tmpFile
+    bzip2 $tmpFile
+    exit 1
+}
+
+runLaTeX() {
+    if ! $latex $latexFlags $rootName >$tmpFile 2>&1; then showError; fi
+    runNeeded=
+    if fgrep -q \
+        -e "LaTeX Warning: Label(s) may have changed." \
+        -e "Rerun to get citations correct." \
+        "$tmpFile"; then
+        runNeeded=1
+    fi
+}
+
+echo
+
+
+if test -n "$copySources"; then
+    cp -prd $TMPDIR/root $out/tex-srcs
+fi
+
+
+echo "PASS 1..."
+runLaTeX
+echo
+
+
+for auxFile in $(find . -name "*.aux"); do
+    # Run bibtex to process all bibliographies.  There may be several
+    # when we're using the multibib package.
+    if grep -q '\\citation' $auxFile; then
+        auxBase=$(basename $auxFile .aux)
+        if [ -e $auxBase.bbl ]; then
+            echo "SKIPPING BIBTEX ON $auxFile!"
+        else
+            echo "RUNNING BIBTEX ON $auxFile..."
+            bibtex --terse $auxBase
+            cp $auxBase.bbl $out
+            runNeeded=1
+        fi
+        echo
+    fi
+
+    # "\pgfsyspdfmark" in the aux file seems to indicate that PGF/TikZ
[…]

### 2017-03-30 paper/src/paper.tex
re-include erased modifications.

--- 2017-03-30 paper/paper.tex
@@
 
-This section presents related works in the area of scheduling for batch scheduling.
+This section presents and discuss the most significant results related 
+to job scheduling and learning algorithms in this context.
+Let us start by recalling some basics about scheduling heuristics in HPC 
+platforms:
 
-\subsection{Scheduling heuristics in HPC platforms}
+While parallel job scheduling is a old studied theoretical
+problem~\cite{leung2004handbook} \textbf{PREFER DROZDOWSKI HERE + SURVEY 
+FEITELSON-RUDOLPH-SWIEGELSHON}, the practical ramifications, varying
+hypotheses, and inherent uncertainty of the problem applied on the HPC 
+field have driven
+practitioners to use \textit{simple} heuristics (and researchers to 
+study their behavior).
+The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
+Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.
 
-While parallel job scheduling is a well studied theoretical
-problem~\cite{leung2004handbook}, the practical ramifications, varying
-hypotheses, and inherent uncertainty of the problem in HPC have driven
-practitioners and researchers alike to use and study simple heuristics.
-The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
-Conservative~\cite{Mu'alem:2001:UPW:380314.380315} Backfilling.
-
-While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
-significant computational overhead, perhaps explaining why most of the machines
-of the top500 ranking~\cite{top500} still use at the time of this publication a
-variant of EASY Backfilling.
-
-\subsection{EASY}
-There is a large body of work seeking to improve EASY. Indeed, while the
-heuristic is used by various resource and job management softwares (most notably
-SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by system
+While Conservative Backfilling offers many advantages~\cite{bfchar}, it 
+has a
+significant computational overhead, which mainly explains why most of 
+the machines
+of the top500 ranking~\cite{top500} rather use a variant of EASY-BF 
+instead.
+There is a large body of work seeking to improve/tune EASY. Indeed, 
+while the
+basic mechanism is used by some actual resource and job management 
+softwares (most notably
+SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by 
+system
 administrators.
 
+The original EASY mechanism refers to a First-Come-First-Serve basis.
 Several works explore how to tune EASY by reordering waiting
@@
 However, these works do not address the duality between the cumulative and
-maximal scheduling costs, as mentionned in~\cite{learningruntimes}.
+maximal scheduling costs, as mentioned in~\cite{learningruntimes}.
+While these previous works intend to estimate uncertain parameters, we 
+consider here a more
+pragmatic approach, which directly learn a good scheduling policy from a 
+given policy space.
 
-A recent work~\cite{mlmem} proposes to help
-users specify better memory requirements by applying supervised learning
-techniques to job metadata.
-
-While these previous works intend to estimate uncertain parameters, they lack
-in validating the impact on the system. We consider in this paper a more
-pragmatic approach, which is to directly learn a scheduling policy from a given
-seach space.
-
-\subsection{Multi-Armed Bandits}
+\paragraph
+\textbf{Multi-Armed Bandits.}
 
 A multi-armed bandit (MAB) problem is a sequential allocation problem with
-partially observable rewards. At every round, an action (an arm) must be chosen
+partially observable rewards. At every round, an action (represented by 
+an arm of the Bandit) must be chosen
 in a fixed set and the corresponding reward is observed. The goal of a MAB
-algorithm is to maximize the total reward obtained in a number of rounds.
-There are a number of works that address this problem under a variety of
-constraints, The two most popular settings are the original stochastic  case,
-and the adversarial case. See ~\cite{thompson} for the original work on the
-stochastic case and ~\cite{Auer2002} for the UCB family of algorithms. The work
-of ~\cite{Banos} is the earliest known work to us for the adversarial case
+algorithm is to maximize the total reward obtained in a successive 
+number of rounds.
+There exist a bunch of works that address this problem under a variety of
+constraints, The two most popular settings are the original stochastic  case
+and the adversarial case. See~\cite{thompson} for the original work on the
+stochastic case and~\cite{Auer2002} for the UCB family of algorithms 
+\textbf{EXPLAIN HERE THE ACRONYM}.
+According to our knowledge, the work
+of~\cite{Banos} is the earliest known work to us for the adversarial case
 see~\cite{nonstoch} for the EXP family of algorithms. We refer to the

### 2017-03-30 paper/src/paper.tex
re-include the non-pushed changes.

@@
 \usepackage{listings}
+\usepackage{hyperref}
 \usepackage{booktabs}
@@
 probabilistic integration procedures~\cite{Nissimov2008}.
-
 However, these works do not address the duality between the cumulative and
@@
 
-Authors are sorted in alphabetical order. We warmly thank Pierre Neyron for help with the experiments. We gracefully thank the contributors
-of the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2), Travis Earheart
+Authors are sorted in alphabetical order. The authors would like to warmly
+thank Pierre Neyron for help with computational experiments and Arnaud Legrand
+for discussions about reproducibility. We gracefully thank the contributors of
+the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2), Travis Earheart
 and Nancy Wilkins-Diehr (SDSC Blue), Lars Malinowsky (KTH SP2), Dan Dwyer and
@@
 
-\section*{Artifacts Description Appendix}
+\section{artifact description: [online (bandit) policy selection for easy-backfilling]}
 
-This appendix describes the computational artifacts associated with this work. We
+\subsection{Abstract}
+
+this appendix describes the computational artifacts associated with this work. we
 take a lightweight approach to replicability by packaging a declarative
 workflow that generates all the figures from this paper and making it available
-publicly. Moreover, questions regarding replication can be freely directed at
+publicly. moreover, questions regarding replication can be freely directed at
 the authors via e-mail.
 
-All the code associated with this work is released under the ~\cite{ISC}
-license as a persistent Zenodo~\cite{zenodo} archive at~\cite{zenodomain}.
+all the code associated with this work is released under the ~\cite{isc}
+license as a persistent zenodo~\cite{zenodo} archive at~\cite{zenodomain}.
 
-\subsection{Packaging and Replicability}
+\subsection{Description}
 
@@
 among which distributing complete operating system images, containers, or
-packaging software. As our experiments are CPU bound, we decide to opt for the
-software packaging approach using the Nix~\cite{nix} packaging system. This
+packaging software. as our experiments are cpu bound, we decide to opt for the
+software packaging approach using the nix~\cite{nix} packaging system. This
 allows for replicating the experiments on clusters where virtualization is not
-available or kernels are old and decreases runtime. All the dependencies of our
+available or kernels are old and decreases runtime. all the dependencies of our
 experiments (including our own code, dependencies for data processing and
@@
 actual execution of the code.
+
+\subsubsection{check-list (artifact meta information)}
+
+{\small
+\begin{itemize}
+  \item {\bf algorithm: Scheduling Simulation Workflow}
+  \item {\bf program: Ocaml, R, Bash, Zymake(all sources and dependencies included)}
+  \item {\bf compilation: Ocaml 4.02.3, GNU bash version 4.4.5(1)-release, R version 3.3.2}
+  \item {\bf transformations: Patched nix standard environment.}
+  \item {\bf binary: Most binaries are cached by Nix channels.}
+  \item {\bf data set: Traces from the parallel workload archive, included.}
+  \item {\bf run-time environment: Nixpkgs 17.03}
+  \item {\bf hardware: Dell PowerEdge R730}
+  \item {\bf output: Files}
+  \item {\bf experiment workflow: \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|, see below}
+  \item {\bf publicly available?: Yes}
+\end{itemize}
+}
+
+\subsubsection{Obtaining the code.}
 
@@
 
-\subsection{Testbed}
+\subsubsection{Hardware}
 
@@
 
- \subsection{Workflow}
+\subsubsection{Software dependencies}
 
- All experiments are tied together using Zymake~\cite{zymake}, a minimalistic
- workflow system designed for computational experiments\footnote{The zymake system
- is also packaged by our Nix expressions.}. This system is
+Managed by Nix.
+
+\subsubsection{Datasets}
+
+From Swf Archive. Cached.
+
+\subsection{Installation}
+
+nix-build
+
+\subsection{Experiment workflow}
+
+ All experiments are tied together using zymake~\cite{zymake}, a minimalistic
+ workflow system designed for computational experiments\footnote{the zymake system
+ is also packaged by our nix expressions.}. This system is
  analogous to a traditional build system with added workflow capabilities. The
@@
  \begin{itemize}
-   \item[Preprocessing]: Data extraction from archives, filtering, resampling.
-     This is principally using shell scripts for data manipulation (see file
+   \item[data]: Data extraction from archives, filtering, resampling.
+     this is principally using shell scripts for data manipulation (see file
      \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter| for the
@@
 
-   \item[Simulation]: This step runs the lightweight ocaml backfilling
-     simulator specially written for this work. This simulator is made
-     available under the ISC~\cite{ISC} license both as a persistent Zenodo
+   \item[simulation]: This step runs the lightweight ocaml backfilling
+     simulator specially written for this work. this simulator is made
+     available under the isc~\cite{isc} license both as a persistent zenodo
      archive at~\cite{ocst} and as a git repository at ~\cite{ocstgit}.
 
-   \item[Analysis]: This step runs R code that generates the figures
+   \item[analysis]: This step runs r code that generates the figures
      presented in this paper.
  \end{itemize}
+
+\subsection{Evaluation and expected result}
+
+o/zymakefile/pdfs
+
+\subsection{Experiment customization}
+
+modify zymakefile
+
+\subsection{Notes}
+This artefact description was prepared according to the guidelines located at \url{http://ctuning.org/ae/submission.html}
 \end{document}

### 2017-03-30 paper/src/paper.tex
modif artifact

@@
 \usepackage{listings}
+\usepackage{dirtree}
 \usepackage{hyperref}
@@
 at the interface of HPC and BigData, dealing with learning algorithms
-(including the authors of this work~\cite{SC2015}).
+(including the authors of this work~\cite{learningruntimes}).
 Most of these studies target the process of determining more or less
@@
 application interference~\cite{7516031}.  Some others are closer to the
-application
-level and use binary classification to distinguish benign memory faults from
-application errors in order to execute recovery algorithms 
+application level and use binary classification to distinguish benign memory
+faults from application errors in order to execute recovery algorithms
 (see~\cite{fmodeling} for instance).
@@
 \begin{equation}
-\label{eq:cost-exact}
+\label{eq:cost-noisy}
 w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho \mbox{Wait}_j^{P_i}
@@
 PERSYVAL-Lab(\grantsponsor{ANR-11-LABX-0025-01}{ANR-11-LABX-0025-01}) funded by
-the French program Investissement d'avenir. Experiments presented in this paper
-were carried out using the Grid'5000 testbed. Grid'5000 is supported by a
-scientific interest group hosted by Inria and including CNRS, RENATER and
-several Universities as well as other
-organizations\footnote{https://www.grid5000.fr}.  Access to the experimental
-machine(s) used in this paper was gracefully granted by research teams from
-LIG\footnote{http://www.liglab.fr} and Inria\footnote{http://www.inria.fr}.
-
+the French program Investissement d'avenir.
 \end{acks}
@@
 
-this appendix describes the computational artifacts associated with this work. we
+This appendix describes the computational artifacts associated with this work. We
 take a lightweight approach to replicability by packaging a declarative
 workflow that generates all the figures from this paper and making it available
-publicly. moreover, questions regarding replication can be freely directed at
+publicly. Moreover, questions regarding replication can be freely directed at
 the authors via e-mail.
 
-all the code associated with this work is released under the ~\cite{isc}
+All the code associated with this work is released under the ~\cite{isc}
 license as a persistent zenodo~\cite{zenodo} archive at~\cite{zenodomain}.
@@
 among which distributing complete operating system images, containers, or
-packaging software. as our experiments are cpu bound, we decide to opt for the
-software packaging approach using the nix~\cite{nix} packaging system. This
-allows for replicating the experiments on clusters where virtualization is not
-available or kernels are old and decreases runtime. all the dependencies of our
+packaging software. As our experiments are cpu bound, we decide to opt for the
+software packaging approach. This allows for replicating the experiments on clusters where virtualization is not
+available or kernels are too old for containers old and decreases runtime.
+
+All the dependencies of our
 experiments (including our own code, dependencies for data processing and
@@
 
-\subsubsection{check-list (artifact meta information)}
+\subsubsection{Check-list (artifact meta information)}
 
@@
 
- The experiments in this paper can run via in less than 24 hours on a moderately parallel
- host. More precisely, we had access to a Dell PowerEdge R730 equipped with
- a total of 56 threads @2.4GHz each and 757G of RAM. For this reason, the
- experimental campaign workflow was designed to be executed on a single host.
+ Experiments presented in this paper were carried out using the Grid'5000
+ testbed. Grid'5000 is supported by a scientific interest group hosted by Inria
+ and including CNRS, RENATER and several Universities as well as other
+ organizations\footnote{https://www.grid5000.fr}.  Access to the experimental
+ machine(s) used in this paper was gracefully granted by research teams from
+ LIG\footnote{http://www.liglab.fr} and Inria\footnote{http://www.inria.fr}.
+
+ The experiments in this paper can run via in less than 24 hours on a
+ moderately parallel host. More precisely, we had access to a Dell PowerEdge
+ R730 equipped with a total of 56 threads @2.4GHz each and 757G of RAM. For
+ this reason, the experimental campaign workflow was designed to be executed on
+ a single host.
 
@@
 
+All the software dependencies of our code
 Managed by Nix.
+
+\dirtree{%
+.1 root .
+.2 zymakefile .
+.2 default.nix .
+.2 pkgs .
+.2 gz .
+.3 ANL-Intrepid.swf.gz .
+.3 CEA-Curie.swf.gz .
+.3 CTC-SP2.swf.gz .
+.3 KTH-SP2.swf.gz .
+.3 SDSC-BLUE.swf.gz .
+.3 SDSC-SP2.swf.gz .
+.3 UniLu-Gaia.swf.gz .
+.2 misc .
+.3 strongvisu .
+.3 visu .
+.2 ocst .
+.2 paper .
+}
 

### 2017-03-30 paper/src/paper.tex
artifact change
small fixes to the artifacts.

@@
   \item {\bf experiment workflow: \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|, see below}
-  \item {\bf publicly available?: Yes}
+  \item {\bf publicly available?: Open-Sourced under the ISC~\cite{isc} license.}
 \end{itemize}
@@
 \subsubsection{Obtaining the code.}
+
+\textbf{one-liner : replication}
 
@@
 
-All the software dependencies of our code
-Managed by Nix.
-
-\dirtree{%
-.1 root .
-.2 zymakefile .
-.2 default.nix .
-.2 pkgs .
-.2 gz .
-.3 ANL-Intrepid.swf.gz .
-.3 CEA-Curie.swf.gz .
-.3 CTC-SP2.swf.gz .
-.3 KTH-SP2.swf.gz .
-.3 SDSC-BLUE.swf.gz .
-.3 SDSC-SP2.swf.gz .
-.3 UniLu-Gaia.swf.gz .
-.2 misc .
-.3 strongvisu .
-.3 visu .
-.2 ocst .
-.2 paper .
-}
+All the software dependencies of our experiments are automatically managed by Nix.
 
@@
 
-From Swf Archive. Cached.
+All data from this paper was obtained from the Swf Parallel Workloads
+Archive~\cite{Feitelson20142967}. The data is cached in gzipped format in the
+persistent archive associated with this workflow, should the Parallel
+Workdloads Archive become unavailable.
 
@@
 
-nix-build
+The experiment dependencies are loaded into the environment by the
+\lstinline[basicstyle=\ttfamily]|nix-build| command.
+
+Nevertheless, the reader interested in using the tools outside the workflow can
+obtain an environment by running the
+\lstinline[basicstyle=\ttfamily]|nix-shell| command:
+
+\begin{lstlisting}
+nix-shell (fetchTarball ...)
+\end{lstlisting}
+
+Or alternatively, if the user is at the root of the extracted archive:
+
+\begin{lstlisting}
+nix-shell -A banditSelection
+\end{lstlisting}
+
+This allows to use the following command-line tools:
+
+\begin{itemize}
+
+  \item[ocs] The scheduling simulator used in the experiments. Use
+    \lstinline[basicstyle=\ttfamily]|ocs --help| to see all available options.
+
+  \item[ocs-sampler] The workload resampler used in the experiments. Use
+    \lstinline[basicstyle=\ttfamily]|ocs-sampler --help| for usage.
+
+  \item[zymake] The zymake~\cite{zymake} tool.
+    The workflow can be executed in the current environment in parallel
+    via \lstinline[basicstyle=\ttfamily]|zymake -l localhost zymakefile|.
+
+\end{itemize}
 
@@
 
-o/zymakefile/pdfs
+All resulting figures will be situated in the simlinked folder named
+\lstinline[basicstyle=\ttfamily\color{blue}]|result| positionned at the root of
+the archive.
 
@@
 
-modify zymakefile
+Readers interested with modifying the experiments can obtain the zenodo archive at \url{https://...}
+
+The experiments are replicated with the same procedure as before, this time invoking
+the build command with the local folder:
+
+\begin{lstlisting}
+nix-build -A banditSelection
+\end{lstlisting}
+
+The file tree has the following structure:
+
+\dirtree{%
+  .1 root .
+.2 zymakefile \textcolor{brown}{Zymake workflow} .
+.2 default.nix \textcolor{brown}{Nix packaging} .
+.2 pkgs/ \textcolor{brown}{Nix packaging} .
+.2 gz/ \textcolor{brown}{Data archive} .
+.3 ANL-Intrepid.swf.gz .
+.3 $\cdots$ .
+.2 misc/ \textcolor{brown}{Visualization and filtering code} .
+.2 ocst/ \textcolor{brown}{Simulator Code} .
+.2 paper/ \textcolor{brown}{This article.} .
+}
+
+Most simple workflow modifications (using different traces, changing the sample
+sizes) can be obtained by modifying the
+\lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|. For modification
 

### 2017-03-30 paper/paper.tex
all

@@
 
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[scale=0.6]{figures/variability.pdf}
+  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (see Subsection~\ref{sub:traces}) for 7 different policies.}
+  \label{fig:mosn}
+\end{figure*}
+
 We present here the strategies we have retained for selecting a policy.
-In the remainder, we will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$.
-The time interval is thus divided into periods of equal length: $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period; all periods have length $\Delta$.
-At the beginning of each period, from the first period $\Delta_0$ to the current period $\Delta_T$, a new policy is selected.
+We will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$.
+The time interval is thus divided into periods of equal lengths ($\Delta$): $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period.
+A new policy is selected at the beginning of each period and applied during the whole period.
 
-We assume here that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods.
+We further assume that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods.
 This assumption is validated in the study presented in~\ref{???}.
 It further entails that the behavior of a policy on the previous periods reflects its behavior on the current one, so that the selection of a policy can be based on its behavior on previous periods.
-However, there usually is an important variability on the cost metric between different periods, for any policy.
-This is illustrated in Figure~\ref{fig:variability} where \textcolor{orange}{???}.
-When the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
+However, there may be some variability between different periods for certain cost metrics.
+This is illustrated in Figure~\ref{fig:mosn} that displays the average waiting time for a trace using weekly periods encompassing one year.
+As one can note, the average waiting time varies a lot from one week to the other, for all the 7 policies considered.
+THis shows that when when the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
+We will come back to this issue below.
 
@@
 Several simulators have been developed for "playing" reordering policies on a given set of jobs.
-Such simulations are interesting inasmuch as they provide an estimate of the cost of a given policy on a set of jobs.
-\textcolor{orange}{TO BE COMPLETED}
+We rely in this study on a lightweight simulator (see Section~\ref{sec:experiments}) that can efficiently simulate different policies.
+Such simulators are interesting inasmuch as they provide an estimate of the cost of a given policy on a set of jobs, as described below.
 
@@
 \label{ssub:resampling}
-
-\begin{figure*}[ht]
-  \centering
-  \includegraphics[scale=0.6]{figures/variability.pdf}
-  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (pre-processing described in subsection~\ref{sub:traces}) for a few policies.}
-  \label{fig:mosn}
-\end{figure*}
 

### 2017-03-30 paper/paper.tex
commit du jeudi soir

@@
 \usepackage{amsfonts}
+\usepackage{bbm}
 \usepackage[scientific-notation=true]{siunitx}
@@
 
-\section{Related Works}
+\section{Related Work}
 \label{sec:rw}
@@
 As one can note, the average waiting time varies a lot from one week to the other, for all the 7 policies considered.
-THis shows that when when the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
+This indicates that when when the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
 We will come back to this issue below.
@@
 
-\subsection{Online policy selection with exact simulation}
+\subsection{Policy selection with exact simulation}
 \label{sub:feedback}
@@
 $\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent history (\textit{i.e.} recent periods).
-Note that the above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.
+One then selects the policy $P$ that minimizes the above cost for the period $\Delta_T$:
+%
+\begin{equation}
+\label{eq:select-exact}
+P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
+\end{equation}
 
-In the context of this study, as detailed in Section~\ref{sec:experiments}, we rely on metrics averaged over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies.
+The above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.
+
+Furthermore, in the context of this study and as detailed in Section~\ref{sec:experiments}, we rely on averages over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies.
 Such traces are typically obtained by simulation.
 Using the same simulator for generating the traces and estimating the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact} would however be too optimistic and would represent an upper bound on what can be achieved by a selection strategy based on simulation.
-
 In order to have a more realistic estimate of the behavior of a selection strategy based on simulation, we introduce noise in the simulator, as described below.
 
-\subsection{Online policy selection with noisy simulation}
+\subsection{Policy selection with noisy simulation}
 \label{sub:noisy}
@@
 %
+As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above cost is selected for the period $\Delta_T$.
+
 Instead of introducing noise in the output of the simulator when selecting the policy, one could have introduced it when generating the execution traces.
 This would however have prevented us from comparing the exact and noisy selection methods, the first one representing an upper bound on the second one.
-As we will see in Section~\ref{sec:experiments}, there is little difference between the two selection methods.
+As we will see in Section~\ref{sec:experiments}, there is little difference between the two selection methods, which is an important result of our study.
 
-\subsection{Bandit policy selection}
+Lastly, as before, the length of the period has no impact here as the cost metric corresponds to the cumulative simulated waiting time for each policy.
+
+\subsection{Policy selection with $\epsilon$-greedy exploration}
 \label{sub:bandit}
 
-When we have only bandit feedback of the accurate simulator.
-We use the epsilon-greedy bandit\ref{alg:epsg}.
+The previous strategies estimate the cost of all policies over all previous periods; the best policy according to this cost is then selected for the current period, the costs of all policies being updated for the next period.
+This is interesting as one maintains a complete knowledge of all policies over time.
+However, this requires computing many estimates at each period (as many as there are policies), which can be time consuming even with lightweight simulators.
 
-\begin{algorithm}[h]
-  \caption{Epsilon-Greedy Bandit policy}
-  \begin{algorithmic}[1]
-    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
-    \renewcommand{\algorithmicensure}{\textbf{Output:}}
-    \REQUIRE $0 < \epsilon < 1$, $K>1$
-    \FOR {$t = 1, 2, \ldots$}
-    \STATE Let $i_t$ be the reordering heuristic with the highest current
-    average reward.
-    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
-    $\epsilon$ use a random reordering heuristic.
-    \ENDFOR
-  \end{algorithmic}
-  \label{alg:epsg}
-\end{algorithm}
+We thus explore here a more efficient strategy that dispenses with estimating the cost of all policies at a given time and only updates the cost of the policy that is currently being used.
+This strategy makes use of the $\epsilon$-greedy exploration method, standard in reinforcement learning \cite{Watkins:1989}, to tradeoff exploitation and exploration, and relies, as before, on past estimates of the cost to select the best policy in the exploitation mode.
+
+Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during the period $\Delta_t$. We define the cost of policy $P_i$ during the period $\Delta_t$ as:
+%
+\[
+w_{\epsilon}(t;P_i) =
+\left\{
+\begin{array}{ll}
+ \sum_{j=1}^{l(\Delta_t)} \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t \nonumber\\
+ 0 & \mbox{otherwise}
+ \end{array}
+\right.
+\]
+%
+As one can note, the actual waiting time is used here, so that one dispenses with the use of a simulator for efficiency considerations.
+This indeed leads to faster and more accurate estimate of the cumulative waiting time, however only for the policy that is used during $\Delta_t$.
+This also explains the difference between the definitions of $l(\Delta_t)$: in the previous selection strategies, one relies on a simulated estimate so that jobs need not be finished during $\Delta_t$, but only submitted; in contrast, to know their actual waiting time, jobs need be finished during $\Delta_t$.
+
+The cost over all previous period of a policy $P_i$ is then defined as:
+%
+\begin{equation}
+\label{eq:tot-cost-greedy}
+w(\rightarrow T;P_i) = \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i)
+\end{equation}
+%
+where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period $\Delta_t$ and $0$ otherwise.
+The normalization by $\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary here to ensure that policies remain comparable over time, as not normalizing would favor the policy that was first selected, even if this choice was random.
+This normalization however entails that the size of the policy period is important: it should not be too large so as to avoid relying on too few points for estimating the cost, and it should not be too small to avoid extreme variations between periods.
+As explained in Section~\ref{sec:experiments}, we rely in this study on periods of one day and one week.
+
+The selection of the policy to be used during the current time period $\Delta_T$ is then based on the standard $\epsilon$-greedy exploration strategy:
+%
+\begin{itemize}
+\item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
+%
+\begin{equation}
+\label{eq:select-greedy}
+P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
+\end{equation}
+%
+\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 10$, at random (exploration mode).
+\end{itemize}
+
+Several studies propose to update $\epsilon$ over time, the exploration being less important once the estimates for the different policies are reliable \cite{Tokic:2010}.
+This strategy has however not been beneficial in our case.
+We believe this is due to the stochastic nature of the traces we use, as explained in Section~\ref{sec:experiments}.
+
+%\begin{algorithm}[h]
+%  \caption{Epsilon-Greedy Bandit policy}
+%  \begin{algorithmic}[1]
+%    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
+%    \renewcommand{\algorithmicensure}{\textbf{Output:}}
+%    \REQUIRE $0 < \epsilon < 1$, $K>1$
+%    \FOR {$t = 1, 2, \ldots$}
+%    \STATE Let $i_t$ be the reordering heuristic with the highest current
+%    average reward.
+%    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
+%    $\epsilon$ use a random reordering heuristic.
+%    \ENDFOR
+%  \end{algorithmic}
+%  \label{alg:epsg}
+%\end{algorithm}
 

### 2017-03-31 paper/src/paper.tex
include Eric's modifications.

@@
 However, these works do not address the duality between the cumulative and
-maximal scheduling costs, as mentioned in~\cite{learningruntimes}.
-While these previous works intend to estimate uncertain parameters, we
-consider here a more
-pragmatic approach, which directly learn a good scheduling policy from a
-given policy space.
+maximal scheduling costs, as mentionned in~\cite{learningruntimes}.
+
+A recent work~\cite{mlmem} proposes to help
+users specify better memory requirements by applying supervised learning
+techniques to job metadata.
+
+While these previous works intend to estimate uncertain parameters, they lack
+in validating the impact on the system. We consider in this paper a more
+pragmatic approach, which is to directly learn a scheduling policy from a given
+seach space.
 
@@
 
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[scale=0.6]{figures/variability.pdf}
+  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (see Subsection~\ref{sub:traces}) for 7 different policies.}
+  \label{fig:mosn}
+\end{figure*}
+
 We present here the strategies we have retained for selecting a policy.
-In the remainder, we will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$.
-The time interval is thus divided into periods of equal length: $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period; all periods have length $\Delta$.
-At the beginning of each period, from the first period $\Delta_0$ to the current period $\Delta_T$, a new policy is selected.
+We will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$.
+The time interval is thus divided into periods of equal lengths ($\Delta$): $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period.
+A new policy is selected at the beginning of each period and applied during the whole period.
 
-We assume here that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods.
+We further assume that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods.
 This assumption is validated in the study presented in~\ref{???}.
 It further entails that the behavior of a policy on the previous periods reflects its behavior on the current one, so that the selection of a policy can be based on its behavior on previous periods.
-However, there usually is an important variability on the cost metric between different periods, for any policy.
-This is illustrated in Figure~\ref{fig:variability} where \textcolor{orange}{???}.
-When the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
+However, there may be some variability between different periods for certain cost metrics.
+This is illustrated in Figure~\ref{fig:mosn} that displays the average waiting time for a trace using weekly periods encompassing one year.
+As one can note, the average waiting time varies a lot from one week to the other, for all the 7 policies considered.
+This indicates that when when the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
+We will come back to this issue below.
 
@@
 $\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent history (\textit{i.e.} recent periods).
-Note that the above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.
+One then selects the policy $P$ that minimizes the above cost for the period $\Delta_T$:
+%
+\begin{equation}
+\label{eq:select-exact}
+P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
+\end{equation}
 
-In the context of this study, as detailed in Section~\ref{sec:experiments}, we rely on metrics averaged over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies.
+The above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.
+
+Furthermore, in the context of this study and as detailed in Section~\ref{sec:experiments}, we rely on averages over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies.
 Such traces are typically obtained by simulation.
 Using the same simulator for generating the traces and estimating the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact} would however be too optimistic and would represent an upper bound on what can be achieved by a selection strategy based on simulation.
-
 In order to have a more realistic estimate of the behavior of a selection strategy based on simulation, we introduce noise in the simulator, as described below.
 
-\subsection{Online policy selection with noisy simulation}
+\subsection{Policy selection with noisy simulation}
 \label{sub:noisy}
 
-In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling, either dow or up, the waiting time of jobs obtained by simulation:
+In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either down or up, for each job:
+%
+\begin{enumerate}
+\item For each job $j$ submitted during $\Delta_t$: $\rho_j \sim U[0.85,1.15]$
+\item Then:
 %
@@
 \label{eq:cost-noisy}
-w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho \mbox{Wait}_j^{P_i}
+w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i}
 \end{equation}
 %
+\end{enumerate}
+%
+$\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a +/-15\% noise on the waiting time estimated by the simulator.
+The overall cost is then defined in the same way as before, leading to:
+%
+\begin{equation}
+\label{eq:tot-cost-noisy}
+w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i)
+\end{equation}
+%
+As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above cost is selected for the period $\Delta_T$.
 
-\subsection{Bandit policy selection}
+Instead of introducing noise in the output of the simulator when selecting the policy, one could have introduced it when generating the execution traces.
+This would however have prevented us from comparing the exact and noisy selection methods, the first one representing an upper bound on the second one.
+As we will see in Section~\ref{sec:experiments}, there is little difference between the two selection methods, which is an important result of our study.
+
+Lastly, as before, the length of the period has no impact here as the cost metric corresponds to the cumulative simulated waiting time for each policy.
+
+\subsection{Policy selection with $\epsilon$-greedy exploration}
 \label{sub:bandit}
-When we have only bandit feedback of the accurate simulator.
-We use the epsilon-greedy bandit\ref{alg:epsg}.
 
-\begin{algorithm}[h]
-  \caption{Epsilon-Greedy Bandit policy}
-  \begin{algorithmic}[1]
-    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
-    \renewcommand{\algorithmicensure}{\textbf{Output:}}
-    \REQUIRE $0 < \epsilon < 1$, $K>1$
-    \FOR {$t = 1, 2, \ldots$}
-    \STATE Let $i_t$ be the reordering heuristic with the highest current
-    average reward.
-    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
-    $\epsilon$ use a random reordering heuristic.
-    \ENDFOR
-  \end{algorithmic}
-  \label{alg:epsg}
-\end{algorithm}
+The previous strategies estimate the cost of all policies over all previous periods; the best policy according to this cost is then selected for the current period, the costs of all policies being updated for the next period.
+This is interesting as one maintains a complete knowledge of all policies over time.
+However, this requires computing many estimates at each period (as many as there are policies), which can be time consuming even with lightweight simulators.
+
+We thus explore here a more efficient strategy that dispenses with estimating the cost of all policies at a given time and only updates the cost of the policy that is currently being used.
+This strategy makes use of the $\epsilon$-greedy exploration method, standard in reinforcement learning \cite{Watkins:1989}, to tradeoff exploitation and exploration, and relies, as before, on past estimates of the cost to select the best policy in the exploitation mode.
+
+Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during the period $\Delta_t$. We define the cost of policy $P_i$ during the period $\Delta_t$ as:
+%
+\[
+w_{\epsilon}(t;P_i) =
+\left\{
+\begin{array}{ll}
+ \sum_{j=1}^{l(\Delta_t)} \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t \nonumber\\
+ 0 & \mbox{otherwise}
+ \end{array}
+\right.
+\]
+%
+As one can note, the actual waiting time is used here, so that one dispenses with the use of a simulator for efficiency considerations.
+This indeed leads to faster and more accurate estimate of the cumulative waiting time, however only for the policy that is used during $\Delta_t$.
+This also explains the difference between the definitions of $l(\Delta_t)$: in the previous selection strategies, one relies on a simulated estimate so that jobs need not be finished during $\Delta_t$, but only submitted; in contrast, to know their actual waiting time, jobs need be finished during $\Delta_t$.
+
+The cost over all previous period of a policy $P_i$ is then defined as:
+%
+\begin{equation}
+\label{eq:tot-cost-greedy}
+w(\rightarrow T;P_i) = \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i)
+\end{equation}
+%
+where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period $\Delta_t$ and $0$ otherwise.
+The normalization by $\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary here to ensure that policies remain comparable over time, as not normalizing would favor the policy that was first selected, even if this choice was random.
+This normalization however entails that the size of the policy period is important: it should not be too large so as to avoid relying on too few points for estimating the cost, and it should not be too small to avoid extreme variations between periods.
+As explained in Section~\ref{sec:experiments}, we rely in this study on periods of one day and one week.
+
+The selection of the policy to be used during the current time period $\Delta_T$ is then based on the standard $\epsilon$-greedy exploration strategy:
+%
+\begin{itemize}
+\item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
+%
+\begin{equation}
+\label{eq:select-greedy}
+P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
+\end{equation}
+%
+\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 10$, at random (exploration mode).
+\end{itemize}
+
+Several studies propose to update $\epsilon$ over time, the exploration being less important once the estimates for the different policies are reliable \cite{Tokic:2010}.
+This strategy has however not been beneficial in our case.
+We believe this is due to the stochastic nature of the traces we use, as explained in Section~\ref{sec:experiments}.
+
+%\begin{algorithm}[h]
+%  \caption{Epsilon-Greedy Bandit policy}
+%  \begin{algorithmic}[1]
+%    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
+%    \renewcommand{\algorithmicensure}{\textbf{Output:}}
+%    \REQUIRE $0 < \epsilon < 1$, $K>1$
+%    \FOR {$t = 1, 2, \ldots$}
+%    \STATE Let $i_t$ be the reordering heuristic with the highest current
+%    average reward.
+%    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
+%    $\epsilon$ use a random reordering heuristic.
+%    \ENDFOR
+%  \end{algorithmic}
+%  \label{alg:epsg}
+%\end{algorithm}
 

### 2017-03-31 paper/src/paper.tex
paper changes

@@
 
-Providing the computing infrastuctures needed to solve actual complex
-problems arising in the various fields of the modern society (including
-climate change, health, green energy or security) is a strategic challenge.
-The main pillar to address this challenge is to build extreme-scale
-parallel and distributed platforms.
-The race for always more computing power and storage capacity does not
-only lead to sophisticated specific exascale platforms, but the
-objective of the community is to design efficient sustained Petascale
-platforms.
-However, there is still a long way to go and many scientific and
-technical problems to solve by the system management software in order
-to adapt to such large-scale evolution and the increasing complexity
-in both architecture and applications.
-The existing job and resource management softwares allow to run tens
-thousands jobs on hundreds thousands cores.
-They are based on robust policies, however, despite their positive
-properties of relative simplicity and their guarantee of no job starvation,
-there is still place for improvements.
-We propose here to study how to tune the classical EASY-BackFilling
-heuristic using two new learning approaches (namely, .
+Providing the computing infrastuctures needed to solve actual complex problems
+arising in the various fields of the modern society (including climate change,
+health, green energy or security) is a strategic challenge.  The main pillar to
+address this challenge is to build extreme-scale parallel and distributed
+platforms.  The race for always more computing power and storage capacity does
+not only lead to sophisticated specific exascale platforms, but the objective
+of the community is to design efficient sustained Petascale platforms.
+However, there is still a long way to go and many scientific and technical
+problems to solve by the system management software in order to adapt to such
+large-scale evolution and the increasing complexity in both architecture and
+applications.  The existing job and resource management softwares allow to run
+tens thousands jobs on hundreds thousands cores.  They are based on robust
+policies, however, despite their positive properties of relative simplicity and
+their guarantee of no job starvation, there is still place for improvements.
+We propose here to study how to tune the ordering of the submission queues in
+the frame of the classical EASY-BackFilling family of heuristics using two new
+online learning approaches (an approach based on a simulator and an approach
+using multi-armed bandit).
+
+Most common work in the literature consider \textit{a posteriori} optimization
+of a scheduling algorithm on a (actual) data set.  The increasing amount and
+diversity of data generated by large scale platforms motivates the community to
+study more sophisticated policies that could leverage these data.  During the
+last few years, there was an explosion of the number of works at the interface
+of HPC and BigData fileds, dealing with learning algorithms .  Recent
+sophisticated approaches developed \textit{ad hoc} supervised machine learning
+algorithms for reducing uncertainty in problem parameters (e.g. job execution
+times, memory requirements, etc.) via prediction.  However, only few works
+study the impact of such techniques on the performances of the resource
+manager.
+
+Contrarily to most existing studies, which consider the learning of specific
+parameters of a trace, we provide a complete system that uses feedback from the
+system for adapting the whole behavior online.  Our focus within this work is
+to learn how the management system behaves within the usual framework of
+EASY-BF.  More precisely, we provide strategies for choosing the most
+appropriate queue reordering policy used for job selection.  This approach is
+assessed by an extensive resampling-based experimental campaign that uses 7
+actual traces.
+
+\subsection{Contributions}
+
+Our main contribution is to investigate two original methods for learning a
+good scheduling policy on a per-system basis.  Each method aims to choose
+the best queue reordering policy in a fixed search set of semantically diverse
+options as fast as possible.
+
+The first method (full feedback) continuously uses simulation on data gathered
+from the past and current system workload. In order to reflect the
+uncertainties in the data, we also evaluate a "noisy" variant in which
+simulations are imprecise.
+
+The second method (bandit feedback) is a simulator-less approach based on a
+multi-armed bandit algorithm that derives feedback by measuring system
+performance.
+
+Both these methods are evaluated with a specially developped simulator that is
+open-sourced and packaged for the community.
 
 \bigskip
-The management generates a huge amount of data.
-During the last few years, there was an explosion of the number of works
-at the interface of HPC and BigData, dealing with learning algorithms
-(including the authors of this work~\cite{learningruntimes}).
-Most of these studies target the process of determining more or less
-accurately the value of some specific key parameters, with the idea that
-learning better estimates should improve the performances of the
-resource manager.
-
-\bigskip
-Our focus within this work is to learn how the management system behaves
-within the usual framework of selecting jobs of the submission queue
-under EASY-BF.
-Contrarily to most existing studies, which consider only one execution
-trace, we extend the learning on a multiple traces basis.
-We built a simulator which generates a huge number of partial executions
-coming from multiple users collected in various actual traces and HPC
-platforms.
-The sampling is explained in detail later in the paper.
-This was motivated because learning for a specific trace/platform leads
-to limited results, customized for one particular instance (then, what
-happens if we target another platform?).
-The collected data are too small and moreover, this approach is usually
-inefficient to deal with extreme unpredictable events.
-The sampling generation allow to smooth the data.
-
-%On se place du point de vue de l'admin system...
-
-\subsection{Contributions}
-
-Our main contribution is to investigate a new way for learning, which
-consists in using of a simulator to generate a large data set extracted
-from various actual traces.
-We propose two variants of the approach: the ideal situation which
-relies on the  precise knowledge of data, and the more realistic
-assumption of noisy data (which better reflect the uncertainties).
-We also propose a new on-line approach based on multi-armed Bandit to
-learn very fast how to prioritize the submission and BackFilling queues.
-\bigskip
-
-Our main results are:
-
-l'approche basee sur le simulateur marche bien. meme avec des donnes
-imprecises.
-
-Le bandit a des resultats raisonnables, il ne necessite que peu de code
-supplementaire...
-
-Dans les details :
-on montre que une gestion aleatoire des queues est meilleure que FCFS,
-et on identifie des politiques qui marchent bien dans plusieurs traces,
-etc..
+
+Our main results are as follows: The approach based on the simulator provides
+very good results, close to the best policy in the search set. In essence, this
+means that we are able to recover in an online fashion existing results that
+divide average waiting times by factor of \textcolor{orange}{X} to
+\textcolor{orange}{X}. Indeed, we show that the required sample sizes for
+policy selection in batch schedulers are small enough for workload resampling to
+be avoided in practice. The variant based on noisy data, which is closer to
+actual conditions of uncertainties on both jobs and platform, behaves also very
+well. The alternate approach based on a multi-armed bandit does also provide
+reasonable results at a much lower cost, dividing the average waiting times by
+a factor of \textcolor{orange}{X} to \textcolor{orange}{X}.
+
+Let us emphasize the main results of this study: While the First-Come,
+First-Serve approach to minimizing the waiting times is a common default
+strategy, it can absolutely be outperformed.  We show how to automate selection
+of a good alternative policy in two cases (depending whether a simulator is
+available or not), and provide the difference in performance to be expected
+between both situations.
 
@@
 
-This section presents and discuss the most significant results related
-to job scheduling and learning algorithms in this context.
-Let us start by recalling some basics about scheduling heuristics in HPC
-platforms:
-
-While parallel job scheduling is a old studied theoretical
-problem~\cite{leung2004handbook} \textbf{PREFER DROZDOWSKI HERE + SURVEY
-FEITELSON-RUDOLPH-SWIEGELSHON}, the practical ramifications, varying
-hypotheses, and inherent uncertainty of the problem applied on the HPC
-field have driven
-practitioners to use \textit{simple} heuristics (and researchers to
-study their behavior).
-The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
-Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.
-
-While Conservative Backfilling offers many advantages~\cite{bfchar}, it
-has a
-significant computational overhead, which mainly explains why most of
-the machines
-of the top500 ranking~\cite{top500} rather use a variant of EASY-BF
-instead.
-There is a large body of work seeking to improve/tune EASY. Indeed,
-while the
-basic mechanism is used by some actual resource and job management
-softwares (most notably
-SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by
-system
-administrators.
-
-The original EASY mechanism refers to a First-Come-First-Serve basis.
-Several works explore how to tune EASY by reordering waiting
-and/or backfilling queues~\cite{Tsafrir_easypp_2005}, sometimes even in a
-randomized manner~\cite{1592720}, as well as some
-implementations~\cite{Jackson2001}. However, as successful as they may
-be, these works
-do not address the dependency of scheduling metrics on the
-workload~\cite{variability}. Indeed these studies most often report
-\textit{post-hoc} performance since
-they compare algorithms after the workload is known.
+This section presents and discuss the most significant results related to job
+scheduling and learning algorithms in this context.  Let us start by recalling
+some basics about scheduling heuristics in HPC platforms:
+
+Parallel job scheduling is a old studied theoretical
+problem~\cite{Frachtenberg:2009:JSS:1692356} \textcolor{orange}{(cite
+feitelson-rudolph-swiegelshon)} whose practical ramifications, varying
+hypotheses, and inherent uncertainty of the problem applied on the HPC field
+have driven practitioners to use simple heuristics (and researchers to
+study their behavior). The two most popular heuristics for HPC platforms are
+EASY~\cite{easy} and Conservative~\cite{Mu'alem:2001:UPW:380314.380315}
+backfillings.
+
+While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
+significant computational overhead, which mainly explains why most of the
+machines of the top500 ranking~\cite{top500} rather use a variant of EASY-BF
+instead.  There is a large body of work seeking to improve/tune EASY. Indeed,
+while the basic mechanism is used by some actual resource and job management
+softwares (most notably SLURM~\cite{SLURMdocSCHED}), this is rarely done
+without fine tunings by system administrators.
+
+The original EASY mechanism refers to a First-Come-First-Serve basis.  Several
+works explore how to tune EASY by reordering waiting and/or backfilling
+queues~\cite{Tsafrir_easypp_2005}, sometimes even in a randomized
+manner~\cite{1592720}, as well as some implementations~\cite{Jackson2001}.
+However, as successful as they may be, these works do not address the
+dependency of scheduling metrics on the workload~\cite{variability}. Indeed
+these studies most often report \textit{post-hoc} performance since they
+compare algorithms after the workload is known.
 
 The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
-method to tuning these queues, although it requires simulated scheduling
-runs
-at decision time and therefore, it costs much more than the natural
-execution of EASY.
-
-\paragraph
-\textbf{Data-aware resource management.}
+method to tuning these queues, although it requires simulated scheduling runs
+at decision time and therefore, it costs much more than the natural execution of EASY.
+
+\subsection{Data-aware resource management.}
 
 There was a recent focus on leveraging the huge amount of data available in
-large scale computing systems in order to improve their behavior. Some works
-use collaborative filtering to colocate tasks in clouds by estimating
+large scale computing platforms in order to improve system performance. Some
+works use collaborative filtering to colocate tasks in clouds by estimating
 application interference~\cite{7516031}. Some others are closer to the
-application level and use binary classification to distinguish benign memory
-faults from application errors in order to execute recovery algorithms
-(see~\cite{fmodeling} for instance).
-
-Several works use \textbf{this method} (LAQUELLE?) in the context of HPC, in
-particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that
-better job
+application level. For instance, ~\cite{fmodeling} uses binary classification
+to distinguish benign memory faults from application errors in order to execute
+recovery algorithms.
+
+Several works use similar techniques in the context of HPC, in
+particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
 runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
 Some algorithms estimate runtime distributions model and choose jobs using
-probabilistic integration procedures~\cite{Nissimov2008}.
-However, these works do not address the duality between the cumulative and
-maximal scheduling costs, as mentionned in~\cite{learningruntimes}.
-
-A recent work~\cite{mlmem} proposes to help
-users specify better memory requirements by applying supervised learning
-techniques to job metadata.
-
-While these previous works intend to estimate uncertain parameters, they lack
-in validating the impact on the system. We consider in this paper a more
-pragmatic approach, which is to directly learn a scheduling policy from a given
-seach space.
-
-\paragraph
-\textbf{Multi-Armed Bandits.}
+probabilistic integration procedures~\cite{Nissimov2008}.  However, these works
+do not address the duality between the cumulative and maximal scheduling costs,
+as mentioned in~\cite{learningruntimes}. While these previous works intend to
+estimate uncertain parameters, we consider here a more pragmatic approach,
+which directly learn a good scheduling policy from a given policy space.
+
+Existing work~\cite{jsspp17} takes this approach in an offline manner, by
+splitting workload data in a training/testing fashion. This work studied
+EASY-Backfilling and shows that the relative performance of some well-known
+priority orders for starting jobs differs between workloads, but is relatively
+stable troughout time for a given workload.
+
+\subsection{Multi-Armed Bandits.}
 
 A multi-armed bandit (MAB) problem is a sequential allocation problem with
-partially observable rewards. At every round, an action (represented by
-an arm of the Bandit) must be chosen
-in a fixed set and the corresponding reward is observed. The goal of a MAB
-algorithm is to maximize the total reward obtained in a successive
-number of rounds.
-There exist a bunch of works that address this problem under a variety of
-constraints, The two most popular settings are the original stochastic  case
-and the adversarial case. See~\cite{thompson} for the original work on the
-stochastic case and~\cite{Auer2002} for the UCB family of algorithms
-\textbf{EXPLAIN HERE THE ACRONYM}.
-According to our knowledge, the work
+partially observable rewards. At every round, an action (represented by an arm
+of the Bandit) must be chosen in a fixed set and the corresponding reward is
+observed. The goal of a MAB algorithm is to maximize the total reward obtained
+in a successive number of rounds.  There exist a bunch of works that address
+this problem under a variety of constraints, The two most popular settings are
+the original stochastic  case and the adversarial case. See~\cite{thompson} for
+the original work on the stochastic case and~\cite{Auer2002} for the "Upper
+Confidence Bound" family of algorithms. According to our knowledge, the work
 of~\cite{Banos} is the earliest known work to us for the adversarial case
-see~\cite{nonstoch} for the EXP family of algorithms. We refer to the
-review~\cite{bubnow} for a comprehensive overview of the field. While these
-algorithms bound the cumulative difference in loss to the best arm (the
+see~\cite{nonstoch} for the "Exponential-weight" family of algorithms. We refer
+to the review~\cite{bubnow} for a comprehensive overview of the field. While
+these algorithms bound the cumulative difference in loss to the best arm (the
 regret), they have functional constraints such as the fact that the rewards
-should be contained in a range. The heuristic called
-\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to
-achieve good results in practice, and does not have
-this requirement.
+should be contained in a range. A simple heuristic called
+\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to achieve good
+results in practice, and does not have this requirement.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-This section presents the systems under study and the scheduling problem.  It
-first introduces the EASY-Backfilling heuristic and gives
-\textcolor{orange}{(fix ugly sentence)} the problem statement.
+%This section presents the systems under study and the scheduling problem.  It
+%first introduces the EASY-Backfilling heuristic and gives \todo{fix} the
+%problem statement.
 
@@
 
-The problem addressed in this work is the core logic of Resource and Job
-Management Systems (RJMS) such as SLURM~\cite{SLURMdocSCHED}, PBS~\cite{PBSdoc},
-OAR~\cite{capit2005batch}, Cobalt~\cite{Cobalt}, and more recently Flux~\cite{flux2014}.
-
-The crucial part of batch scheduling software is the scheduling algorithm that determines
-where and when the submitted jobs are executed. The process is as follows: jobs
-are submitted by end-users and queued until the scheduler selects one of them
-for running. Each job has a provided bound on the execution time and some
-resource requirements (number and type of processing units). Then, the RJMS
-drives the search for the resources required to execute this job. Finally, the
-tasks of the job are assigned to the chosen nodes.
-
-In the classical case, these softwares need to execute a set of concurrent
-parallel jobs with rigid (known and fixed) resource requirements on a HPC
-platform represented by a pool of $m$ identical resources. This is an on-line
-problem since the jobs are submitted over time and their characteristics are only known when they are released.
-Below is the description and the notations of the characteriscs of job $j$:
+The crucial part of batch scheduling software is the scheduling algorithm that
+determines where and when the submitted jobs are executed. The process is as
+follows: jobs are submitted by end-users and queued until the scheduler selects
+one of them for running. Each job has a provided bound on the execution time
+and some resource requirements (number and type of processing units). Then, the
+RJMS drives the search for the resources required to execute this job. Finally,
+the tasks of the job are assigned to the chosen nodes.
+
+In the classical case, the management software needs to execute a set of
+concurrent parallel jobs with rigid (known and fixed) resource requirements on
+a HPC platform represented by a pool of $m$ identical resources. This is an
+on-line problem since the jobs are submitted over time and their
+characteristics are only known when they are released.  Below is the brief
+description (with the corresponding notations) of the characteristics of job
+$j$:
 
 \begin{itemize}
+
    \item Submission date $r_j$ (also called \textit{release date})
@@
 
-  \item Requested running time $\widetilde{p_j}$ (sometimes called \textit{walltime}), which is an upper bound of $p_j$.
+   \item Requested running time $\widetilde{p_j}$ (sometimes called
+     \textit{walltime}), which is an upper bound of $p_j$.
 
@@
 when the job really completes.  Moreover, the users have incentive to
-over-estimate the actual values, since jobs may be ``killed'' if they
-surpass the provided value.
-\subsection{EASY Backfilling}
+over-estimate the actual values, since jobs may be ``killed'' if they surpass
+the provided value.
+
+\subsection{Brief description of EASY Backfilling}
 \label{sub:easy}
 
-The selection of the job to run is performed according to a
-scheduling policy that establishes the order in which the jobs are executed.
-EASY-Backfilling is the most widely used policy due to its simple and robust
-implementation and known benefits such as high system
-utilization~\cite{easy}. This strategy has no worst case guarantee beyond
-the absence of starvation (i.e. every job will be scheduled at some moment).
+The selection of the job to run is performed according to a scheduling policy
+that establishes the order in which the jobs are executed.  EASY-Backfilling is
+the most widely used policy due to its simple and robust implementation and
+known benefits such as high system utilization~\cite{easy}. This strategy has
+no worst case guarantee beyond the absence of starvation (i.e. every job will
+be scheduled at some moment).
 
@@
 a.k.a the maximal value of the waiting time of all the jobs is also worthy of
-interest. Unfortunately, these criteria can sometimes be dual in practice,
-making the problem bi-objective. Subsection~\ref{sub:th} will outline our
-approach to managing this aspect.
-
-\subsection{Problem Statement}
-
-Our problem statement is: \textbf{How to tune EASY-Backfilling in an online manner?}
-\textcolor{orange}{introduce the difference in workloads, the duality of avg and max, and the online problem}
+interest. Unfortunately, these criteria are antagonistic in nature. Section~\ref{sub:th} will outline our
+approach to address the bi-objective aspect of this problem.
 
@@
 can all be obtained from the Parallel Workload
-Archive~\cite{Feitelson20142967}. We use the 'cleaned'\footnote{\lstinline[basicstyle=\ttfamily\color{blue}]|http://www.cs.huji.ac.il/labs/parallel/workload/logs.html#clean|} version of the logs as
-per the archive. We impose an additional
-filtering step to the workloads\footnote{This filtering step is available in
-the reproducible workflow\cite{repro} as shell script
+Archive~\cite{Feitelson20142967}. We use the
+'cleaned'\footnote{\lstinline[basicstyle=\ttfamily\color{blue}]|http://www.cs.huji.ac.il/labs/parallel/workload/logs.html#clean|}
+version of the logs as per the archive. We impose an additional filtering step
+to the workloads\footnote{This filtering step is available in the reproducible
+workflow\cite{repro} as shell script
 \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
@@
 \begin{enumerate}
-  \item If the number of allocated or requested cores of a job  exceeds the size of the machine, we remove the job.
-  \item If the number of allocated or requested cores is negative, we use the available positive value as the request. If both are negative, we remove the job.
+
+  \item If the number of allocated or requested cores of a job  exceeds the
+    size of the machine, we remove the job.
+
+  \item If the number of allocated or requested cores is negative, we use the
+    available positive value as the request. If both are negative, we remove
+    the job.
+
   \item If the runtime or submission time is negative, we remove the job.
+
 \end{enumerate}
@@
 simulator with no network communication modeling and no resource contiguity
-takes more than 7 hours to replay the same trace on the same hardware. The complete
-simulation campaign presented below is executed on a single Dell PowerEdge R730 in 22
-hours, including input preparation and analysis code.
+takes more than 7 hours to replay the same trace on the same hardware. The
+complete simulation campaign presented below is executed on a single Dell
+PowerEdge R730 in 22 hours, including input preparation and analysis code.
 
@@
 
-To carry our experiments, we need to generate many trace samples. More precisely, we want
-to sample the job submission process, which is achieved by splitting the traces in weeks,
-randomly shuffling them and picking the jobs submitted in the selected weeks.  This simple
-mechanism enables us to preserve the dependency within a week. Although the system state
-may not be independent from one week to an other, we can reasonably assume that it evolves
-under its stationary distribution. Thus, computing an average over the sampled traces at a
-given time can be related to a time average of an underlying Markov process, which
-converges to the true quantity thanks to the ergodic theorem.  Moreover, the job
-submission process has no long range correlation making it look like an independent and
-identically distributed process when considering sufficiently time spaced weeks.
+To carry our experiments, we need to generate many trace samples. More
+precisely, we want to sample the job submission process, which is achieved by
+splitting the traces in weeks, randomly shuffling them and picking the jobs
+submitted in the selected weeks.  This simple mechanism enables us to preserve
+the dependency within a week. Although the system state may not be independent
+from one week to an other, we can reasonably assume that it evolves under its
+stationary distribution. Thus, computing an average over the sampled traces at
+a given time can be related to a time average of an underlying Markov process,
+which converges to the true quantity thanks to the ergodic theorem.  Moreover,
+the job submission process has no long range correlation making it look like an
+independent and identically distributed process when considering sufficiently
+time spaced weeks.
 
@@
 invited to peruse to the 'artifacts description' appendix of this paper.  We
-provide to persistent archives containing the simulator, the analysis code
-and an automated workflow that replicates all experiments and figures
-present in this paper.
+provide to persistent archives containing the simulator, the analysis code and
+an automated workflow that replicates all experiments and figures present in
+this paper.
 
@@
     compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
-    UniLu-Gaia trace. The
-    average is obtained by resampling the original trace 100 times. The dashed
-    lines represent the 10th and 90th percentiles of the values across this
-  resampling.}
+    UniLu-Gaia trace. The average is obtained by resampling the original trace
+    100 times. The dashed lines represent the 10th and 90th percentiles of the
+  values across this resampling.}
 
@@
   \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}
-  \caption{Share of the policies chosen by Epsilon-Greedy as a function of time.
-  The average choice is obtained by resampling the original trace 100 times and
-  aggregating by date.}
+
+  \caption{Share of the policies chosen by Epsilon-Greedy as a function of
+  time.  The average choice is obtained by resampling the original trace 100
+times and aggregating by date.}
+
   \label{fig:mosb}
@@
   \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}
-  \caption{Share of the policies chosen by the Noisy policy. The average
-  choice is obtained by resampling the original trace 100 times and
-  aggregating by date.}
+
+  \caption{Share of the policies chosen by the Noisy policy. The average choice
+  is obtained by resampling the original trace 100 times and aggregating by
+date.}
+
   \label{fig:mosn}

### 2017-03-31 paper/src/paper.tex
u

@@
 Parallel job scheduling is a old studied theoretical
-problem~\cite{Frachtenberg:2009:JSS:1692356} \textcolor{orange}{(cite
-feitelson-rudolph-swiegelshon)} whose practical ramifications, varying
-hypotheses, and inherent uncertainty of the problem applied on the HPC field
-have driven practitioners to use simple heuristics (and researchers to
-study their behavior). The two most popular heuristics for HPC platforms are
-EASY~\cite{easy} and Conservative~\cite{Mu'alem:2001:UPW:380314.380315}
-backfillings.
+problem~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865}
+whose practical ramifications, varying hypotheses, and inherent uncertainty of
+the problem applied on the HPC field have driven practitioners to use simple
+heuristics (and researchers to study their behavior). The two most popular
+heuristics for HPC platforms are EASY~\cite{easy} and
+Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.
 
@@
 
-Readers interested with modifying the experiments can obtain the zenodo archive at \url{https://...}
-
-The experiments are replicated with the same procedure as before, this time invoking
-the build command with the local folder:
+Readers interested with modifying the experiments can obtain the zenodo archive
+at \url{https://...}
+
+The experiments are replicated with the same procedure as before, this time
+invoking the build command with the local folder:
 
@@
 \subsection{Notes}
-This artefact description was prepared according to the guidelines located at \url{http://ctuning.org/ae/submission.html}
+
+This artefact description was prepared according to the guidelines located at
+\url{http://ctuning.org/ae/submission.html}
+
 \end{document}

### 2017-03-31 paper/src/paper.tex
changes to eric's section.

@@
-\documentclass[sigconf]{acmart}
+\documentclass[sigconf,review]{acmart}
 \setcopyright{rightsretained}
@@
 Providing the computing infrastuctures needed to solve actual complex problems
-arising in the various fields of the modern society (including climate change,
+arising in the various fields of modern society (including climate change,
 health, green energy or security) is a strategic challenge. The main pillar to
@@
 not only lead to sophisticated specific exascale platforms, but the objective
-of the community is to design efficient sustained Petascale platforms.
-However, there is still a long way to go and many scientific and technical
-problems to solve by the system management software in order to adapt to such
+of the community is to design efficient sustained Petascale platforms. This
 large-scale evolution and the increasing complexity in both architecture and
-applications.  The existing job and resource management softwares allow to run
-tens thousands jobs on hundreds thousands cores.  They are based on robust
-policies, however, despite their positive properties of relative simplicity and
-their guarantee of no job starvation, there is still place for improvements.
-We propose here to study how to tune the ordering of the submission queues in
-the frame of the classical EASY-BackFilling family of heuristics using two new
-online learning approaches (an approach based on a simulator and an approach
-using multi-armed bandit).
+applications create many scientific and technical problems. Accordingly, system
+management software still has a long way to go. The existing job and resource
+management softwares allow to run tens thousands jobs on hundreds thousands
+cores. They are based on robust policies, however, despite their positive
+properties of relative simplicity and their guarantee of no job starvation,
+there is still room for improvement. We propose here to study how to tune the
+ordering of the submission queues in the frame of the classical
+EASY-BackFilling family of heuristics using two new online learning approaches
+(an approach based on a simulator and an approach using multi-armed bandit).
 
@@
 of a scheduling algorithm on a (actual) data set.  The increasing amount and
-diversity of data generated by large scale platforms motivates the community to
+diversity of data generated by large scale platforms motivate the community to
 study more sophisticated policies that could leverage these data.  During the
@@
 Our main contribution is to investigate two original methods for learning a
-good scheduling policy on a per-system basis.  Each method aims to choose
-the best queue reordering policy in a fixed search set of semantically diverse
-options as fast as possible.
+good scheduling policy on a per-system basis. We focus in this work on the
+average waiting time of the jobs. Each method aims to choose the best queue
+reordering policy in a fixed search set of semantically diverse options.
 
@@
 uncertainties in the data, we also evaluate a "noisy" variant in which
-simulations are imprecise.
+simulations are imprecise\footnote{This mitigates the fact that we are
+using the same batch scheduling simulator to learn and validate our policies}.
 
@@
   \includegraphics[scale=0.6]{figures/variability.pdf}
-  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (see Subsection~\ref{sub:traces}) for 7 different policies.}
+
+  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace
+  (see Subsection~\ref{sub:traces}) for 7 different policies. The policy set is
+  reduced as not to obstruct the figure.}
+
   \label{fig:mosn}
@@
 
-We present here the strategies we have retained for selecting a policy.
-We will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$.
-The time interval is thus divided into periods of equal lengths ($\Delta$): $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period.
-A new policy is selected at the beginning of each period and applied during the whole period.
-
-We further assume that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods.
-This assumption is validated in the study presented in~\ref{???}.
-It further entails that the behavior of a policy on the previous periods reflects its behavior on the current one, so that the selection of a policy can be based on its behavior on previous periods.
-However, there may be some variability between different periods for certain cost metrics.
-This is illustrated in Figure~\ref{fig:mosn} that displays the average waiting time for a trace using weekly periods encompassing one year.
-As one can note, the average waiting time varies a lot from one week to the other, for all the 7 policies considered.
-This indicates that when when the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.
-We will come back to this issue below.
+We present here the strategies we have retained for selecting a policy. We will
+refer to the period during which a selected policy is applied as the
+\textit{policy period} and will denote the length of this period as $\Delta$.
+The time interval is thus divided into periods of equal lengths ($\Delta$):
+$\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy
+period. A new policy is selected at the beginning of each period and applied
+during the whole period.
+
+We further assume that there is a certain regularity among periods,
+\textit{i.e.} that the distributions underlying the jobs submitted do not
+radically differ between consecutive periods. This assumption is validated in
+the study presented in~\cite{jsspp17}. It further entails that the behavior of
+a policy on the previous periods reflects its behavior on the current one, so
+that the selection of a policy can be based on its behavior on previous
+periods. However, there may be some variability between different periods for
+certain cost metrics~\cite{feitelson2001metrics}. This is illustrated in
+Figure~\ref{fig:mosn} which displays the average waiting time of various
+policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
+workloads) using weekly periods encompassing one year. As one can note, the
+average waiting time varies a lot from one week to the other, for all the 7
+policies considered. This indicates that when when the cost metric is averaged
+over different periods, there is a tradeoff to find in between longer periods
+that would somehow limit the variability, and shorter ones that yield more
+values for the estimation. We will come back to this issue below.
 
 The selection of a policy is of course reminiscent of reinforcement learning.
-It is important to note, however, that a pure reinforcement learning approach is difficult to develop in our context as \textcolor{orange}{???}.
-We rely here on simpler, yet we believe more effective, strategies to solve this problem.
-These strategies are applied online and rely on exact simulation, noisy simulation and $\epsilon$-greedy bandit exploration.
+It is important to note, however, that a pure reinforcement learning approach
+is difficult to develop in our context. Indeed, while we have outlined a
+reduced action space, the state space to consider is infinite. This is not
+prohibitive \textit{per se}, as modern methods~\ref{approximatereinforcement}
+can bypass dimensionality issues via function approximation. However, these
+methods rely on large amounts of data, and we are studying online (often called
+\textit{on-policy} methods in reinforcement learning) methods. We rely here on
+simpler, yet we believe more effective, strategies to solve this problem.
+These strategies are applied online and rely on exact simulation, noisy
+simulation and $\epsilon$-greedy bandit exploration.
 
@@
 
-Several simulators have been developed for "playing" reordering policies on a given set of jobs.
-We rely in this study on a lightweight simulator (see Section~\ref{sec:experiments}) that can efficiently simulate different policies.
-Such simulators are interesting inasmuch as they provide an estimate of the cost of a given policy on a set of jobs, as described below.
-
-Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 10$) one reordering policy (defined in Section~\ref{subsec:policies}).
-The cost of policy $P_i$ during the period $\Delta_t$ is defined as:
-%
-\begin{equation}
-\label{eq:cost-exact}
-w_{\mbox{exact}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \mbox{Wait}_j^{P_i}
-\end{equation}
-%
-where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of the waiting time for job $j$ according to policy $P_i$.
-The estimation of the cost of a policy over all the periods preceding the current period can then be defined as:
-%
-\begin{equation}
-\label{eq:tot-cost-exact}
-w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{exact}}(t;P_i)
-\end{equation}
-%
-$\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent history (\textit{i.e.} recent periods).
-One then selects the policy $P$ that minimizes the above cost for the period $\Delta_T$:
-%
-\begin{equation}
-\label{eq:select-exact}
-P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
-\end{equation}
-
-The above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.
-
-Furthermore, in the context of this study and as detailed in Section~\ref{sec:experiments}, we rely on averages over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies.
-Such traces are typically obtained by simulation.
-Using the same simulator for generating the traces and estimating the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact} would however be too optimistic and would represent an upper bound on what can be achieved by a selection strategy based on simulation.
-In order to have a more realistic estimate of the behavior of a selection strategy based on simulation, we introduce noise in the simulator, as described below.
+Several simulators have been developed for "playing" reordering policies on a
+given set of jobs. We rely in this study on a lightweight simulator (see
+Section~\ref{sec:experiments}) that can efficiently simulate different
+policies. Such simulators are interesting inasmuch as they provide an estimate
+of the cost of a given policy on a set of jobs, as described below.
+
+Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the
+period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 10$) one reordering
+policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$
+during the period $\Delta_t$ is defined as:
+%
+\begin{equation} \label{eq:cost-exact} w_{\mbox{exact}}(t;P_i) =
+\sum_{j=1}^{l(\Delta_t)} \mbox{Wait}_j^{P_i} \end{equation}
+%
+where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of
+the waiting time for job $j$ according to policy $P_i$. The estimation of the
+cost of a policy over all the periods preceding the current period can then be
+defined as:
+%
+\begin{equation} \label{eq:tot-cost-exact} w(\rightarrow T;P_i) =
+\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{exact}}(t;P_i) \end{equation}
+%
+$\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent
+history (\textit{i.e.} recent periods). One then selects the policy $P$ that
+minimizes the above cost for the period $\Delta_T$:
+%
+\begin{equation} \label{eq:select-exact} P = \mbox{argmin}_{P_i, 1 \le i \le
+10} w(\rightarrow T;P_i) \end{equation}
+
+The above cost directly corresponds to the cumulative waiting time of the
+policy over the preceding periods (more precisely to the cumulative simulated
+estimate of the waiting time of the policy over the preceding periods) so that
+the length of the period has little impact here. The only bias comes from the
+boundary states of the simulation. Indeed, we run simulations from an empty
+system and wait for the system to be empty when job submissions cease at the
+end of the submission period.
+
+Furthermore, in the context of this study and as detailed in
+Section~\ref{sec:experiments}, we rely on averages over several execution
+traces in order to obtain reliable estimate of the behavior of different
+policies and policy selection strategies. Such traces are typically obtained by
+simulation. Using the same simulator for generating the traces and estimating
+the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact}
+would however be too optimistic and would represent an upper bound on what can
+be achieved by a selection strategy based on simulation. In order to have a
+more realistic estimate of the behavior of a selection strategy based on
+simulation, we introduce noise in the simulator, as described below.
 
@@
 
-In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either down or up, for each job:
-%
-\begin{enumerate}
-\item For each job $j$ submitted during $\Delta_t$: $\rho_j \sim U[0.85,1.15]$
-\item Then:
-%
-\begin{equation}
-\label{eq:cost-noisy}
-w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i}
-\end{equation}
-%
-\end{enumerate}
-%
-$\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a +/-15\% noise on the waiting time estimated by the simulator.
-The overall cost is then defined in the same way as before, leading to:
-%
-\begin{equation}
-\label{eq:tot-cost-noisy}
-w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i)
-\end{equation}
-%
-As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above cost is selected for the period $\Delta_T$.
-
-Instead of introducing noise in the output of the simulator when selecting the policy, one could have introduced it when generating the execution traces.
-This would however have prevented us from comparing the exact and noisy selection methods, the first one representing an upper bound on the second one.
-As we will see in Section~\ref{sec:experiments}, there is little difference between the two selection methods, which is an important result of our study.
-
-Lastly, as before, the length of the period has no impact here as the cost metric corresponds to the cumulative simulated waiting time for each policy.
+In order to simulate how the simulation strategy for selecting policies would
+work on non simulated traces, we randomly introduce noise in the estimate of
+the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either
+down or up, for each job:
+%
+\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
+\sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i} \end{equation}
+%
+where $\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a
++/-15\% noise on the waiting time estimated by the simulator. The overall cost
+is then defined in the same way as before, leading to:
+%
+\begin{equation} \label{eq:tot-cost-noisy} w(\rightarrow T;P_i) =
+\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
+%
+As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above
+cost is selected for the period $\Delta_T$.
+
+Instead of introducing noise in the output of the simulator when selecting the
+policy, one could have introduced it when collecting the execution traces. Our choice
+to add the noise after the simulation step is motivated by simplicity. As we
+will see in Section~\ref{sec:experiments}, there is little difference between
+the two selection methods, which is an important result of our study.
+
+Lastly, as before, the length of the period has little impact here as the cost
+metric corresponds to the cumulative simulated waiting time for each policy.
 
@@
 
-The previous strategies estimate the cost of all policies over all previous periods; the best policy according to this cost is then selected for the current period, the costs of all policies being updated for the next period.
-This is interesting as one maintains a complete knowledge of all policies over time.
-However, this requires computing many estimates at each period (as many as there are policies), which can be time consuming even with lightweight simulators.
-
-We thus explore here a more efficient strategy that dispenses with estimating the cost of all policies at a given time and only updates the cost of the policy that is currently being used.
-This strategy makes use of the $\epsilon$-greedy exploration method, standard in reinforcement learning \cite{Watkins:1989}, to tradeoff exploitation and exploration, and relies, as before, on past estimates of the cost to select the best policy in the exploitation mode.
-
-Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during the period $\Delta_t$. We define the cost of policy $P_i$ during the period $\Delta_t$ as:
-%
-\[
-w_{\epsilon}(t;P_i) =
-\left\{
-\begin{array}{ll}
- \sum_{j=1}^{l(\Delta_t)} \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t \nonumber\\
- 0 & \mbox{otherwise}
- \end{array}
-\right.
-\]
-%
-As one can note, the actual waiting time is used here, so that one dispenses with the use of a simulator for efficiency considerations.
-This indeed leads to faster and more accurate estimate of the cumulative waiting time, however only for the policy that is used during $\Delta_t$.
-This also explains the difference between the definitions of $l(\Delta_t)$: in the previous selection strategies, one relies on a simulated estimate so that jobs need not be finished during $\Delta_t$, but only submitted; in contrast, to know their actual waiting time, jobs need be finished during $\Delta_t$.
+The previous strategies estimate the cost of all policies over all previous
+periods; the best policy according to this cost is then selected for the
+current period, the costs of all policies being updated for the next period.
+This is interesting as one maintains a complete knowledge of all policies over
+time. However, this requires computing many estimates at each period (as many
+as there are policies), which can be time consuming or cumbersome even with
+lightweight simulators.
+
+We thus explore here a more efficient strategy that dispenses with estimating
+the cost of all policies at a given time and only updates the cost of the
+policy that is currently being used. This strategy makes use of the
+$\epsilon$-greedy exploration method, standard in reinforcement learning
+\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to tradeoff
+exploitation and exploration, and relies, as before, on past estimates of the
+cost to select the best policy in the exploitation mode.
+
+Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during
+the period $\Delta_t$. We define the cost of policy $P_i$ during the period
+$\Delta_t$ as:
+%
+\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta_t)}
+  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t
+  \nonumber\\ 0 & \mbox{otherwise} \end{array} \right.  \]
+%
+As one can note, the actual waiting time is used here, so that one dispenses
+with the use of a simulator for efficiency considerations. This leads to a
+faster and easier estimate of the cumulative waiting time, however only for the
+policy that is used during $\Delta_t$. Note that there is as previously a bias
+due to boundary effects in the measurement method, as we do include jobs that
+had been started before $\Delta_t$, as well as disregard jobs that were
+submitted during $\Delta_t$ but finish later.
+
+%This also explains the difference
+%between the definitions of $l(\Delta_t)$: in the previous selection strategies,
+%one relies on a simulated estimate so that jobs need not be finished during
+%$\Delta_t$, but only submitted; in contrast, to know their actual waiting time,
+%jobs need be finished during $\Delta_t$.
 
@@
 %
-\begin{equation}
-\label{eq:tot-cost-greedy}
-w(\rightarrow T;P_i) = \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i)
-\end{equation}
-%
-where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period $\Delta_t$ and $0$ otherwise.
-The normalization by $\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary here to ensure that policies remain comparable over time, as not normalizing would favor the policy that was first selected, even if this choice was random.
-This normalization however entails that the size of the policy period is important: it should not be too large so as to avoid relying on too few points for estimating the cost, and it should not be too small to avoid extreme variations between periods.
-As explained in Section~\ref{sec:experiments}, we rely in this study on periods of one day and one week.
-
-The selection of the policy to be used during the current time period $\Delta_T$ is then based on the standard $\epsilon$-greedy exploration strategy:
+\begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
+\frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1}
+\mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
+%
+where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
+$\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
+\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary here to ensure that policies
+remain comparable over time, as not normalizing would favor the policy that was
+first selected, even if this choice was random. This normalization however
+entails that the size of the policy period is important: it should not be too
+large so as to avoid relying on too few points for estimating the cost, and it
+should not be too small to avoid extreme variations between periods. As
+explained in Section~\ref{sec:experiments}, we rely in this study on periods of
+one day and one week.
+
+The selection of the policy to be used during the current time period
+$\Delta_T$ is then based on the standard $\epsilon$-greedy exploration
+strategy:
+
 %
@@
 
-Several studies propose to update $\epsilon$ over time, the exploration being less important once the estimates for the different policies are reliable \cite{Tokic:2010}.
-This strategy has however not been beneficial in our case.
-We believe this is due to the stochastic nature of the traces we use, as explained in Section~\ref{sec:experiments}.
+Several studies propose to decay $\epsilon$ over time, the exploration being
+less important once the estimates for the different policies are reliable
+\cite{Tokic:2010}. This strategy has however not been beneficial in our case.
+We believe this is due to the properties of the traces we use, as
+explained in Section~\ref{sec:experiments}.
 

### 2017-03-31 paper/src/paper.tex
add paper preview.

@@
 arising in the various fields of modern society (including climate change,
-health, green energy or security) is a strategic challenge. The main pillar to
-address this challenge is to build extreme-scale parallel and distributed
-platforms. The race for always more computing power and storage capacity does
-not only lead to sophisticated specific exascale platforms, but the objective
+health, green energy or security) is a strategic challenge. The main pillar of
+the answer to this challenge is to build extreme-scale parallel and distributed
+platforms. The never-ending race for more computing power and storage capacity does
+not only lead to sophisticated specific Exascale platforms, but the objective
 of the community is to design efficient sustained Petascale platforms. This
@@
 management software still has a long way to go. The existing job and resource
-management softwares allow to run tens thousands jobs on hundreds thousands
-cores. They are based on robust policies, however, despite their positive
-properties of relative simplicity and their guarantee of no job starvation,
+management softwares allow to run tens of thousands of jobs on hundreds of thousands
+of cores. They are based on robust policies, however, despite their positive
+properties such as relative simplicity and prevention of job starvation,
 there is still room for improvement. We propose here to study how to tune the
@@
 last few years, there was an explosion of the number of works at the interface
-of HPC and BigData fileds, dealing with learning algorithms .  Recent
+of HPC and BigData fields, dealing with learning algorithms. Recent
 sophisticated approaches developed \textit{ad hoc} supervised machine learning
 algorithms for reducing uncertainty in problem parameters (e.g. job execution
-times, memory requirements, etc.) via prediction.  However, only few works
-study the impact of such techniques on the performances of the resource
+times, memory requirements, etc.) via prediction. However, only few studies
+evaluate the impact of such techniques on the performances of the resource
 manager.
@@
 EASY-BF. More precisely, we provide strategies for choosing the most
-appropriate queue reordering policy used for job selection.  This approach is
-assessed by an extensive resampling-based experimental campaign that uses 7
+appropriate queue reordering policy to be used for job selection. This approach
+is assessed by an extensive resampling-based experimental campaign that uses 7
 actual traces.
@@
 
-Both these methods are evaluated with a specially developped simulator that is
+Both of these methods are evaluated with a specially developped simulator that is
 open-sourced and packaged for the community.
@@
 be avoided in practice. The variant based on noisy data, which is closer to
-actual conditions of uncertainties on both jobs and platform, behaves also very
-well. The alternate approach based on a multi-armed bandit does also provide
+actual conditions of uncertainties on both jobs and platform, also behaves very
+well. The alternate approach based on a multi-armed bandit also provides
 reasonable results at a much lower cost, dividing the average waiting times by

### 2017-03-31 paper/src/paper.tex
paper and last compiled version

@@
 from the past and current system workload. In order to reflect the
-uncertainties in the data, we also evaluate a "noisy" variant in which
-simulations are imprecise\footnote{This mitigates the fact that we are
-using the same batch scheduling simulator to learn and validate our policies}.
+uncertainties in the data, we also evaluate a more realistic "noisy" variant in
+which simulations are imprecise.
 

### 2017-03-31 paper/src/paper.tex
ccl draft

@@
 
+The scheduling of parallel jobs on a given HPC platform is a very hard
+optimization problem with many uncertain parameters. Even thought these
+uncertainties could be reduced, determining efficient strategies remains
+difficult.
+
+We presented in this work a new way of addressing this problem.  The idea was
+to look at the level of the scheduling process instead of trying to change its
+parameters.
+
+More precisely, we showed that it is worth to learn on the scheduling process
+itself. Indeed, reordering the submission queues under EASY-BackFilling leads
+to considerable gains in performance.
+
+This approach was used in two methods: a method based on a simulator and a
+method based on multi-armed bandits. The first one provides very good results,
+improving the average waiting times of the baseline FCFS ordering policy by a
+factor of \textcolor{orange}{X} to \textcolor{orange}{X} The second one is
+slighly less efficient with an improvement factor of  \textcolor{orange}{X} to
+\textcolor{orange}{X}. However, it is more easy to use and less costly to run.
+
+There are two directions for extensions of this work.  First, the question of
+the existence of a more efficient selection strategy for bandit feedback for
+these systems is not addressed by this work. Is a simulator really necessary?
+%@Denis: a essayer.. 1 mois de travail pour tester?
+Second, one can help but wonder what performance could be achieved by extending
+the search space to arbitrary policies. This means doing a complete study
+including lower bounding the cost and studying learning performance using some
+function approximator.
+%@Denis: c'est ce que font vasily et salah de deux manieres differentes.
+As these directions are data-bound, they will require careful application of
+trace resampling techniques in order to be successful.
+
 \begin{acks}

### 2017-03-31 paper/src/paper.tex
lambda value 0

@@
 
+We use devise the following experimental campaign. For each workload trace
+considered, we use the original data to resample 100 traces, each of a length
+of two years. For each resampled trace, we run simulations using the
+EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
+threshold $\Theta$ at 40 hours, a sane default. The choice of this default
+stems from the analysis developped in ~\cite{jsspp17}. This value is high
+enough for queue reorderings to be leveraged while being low enough to provide
+the functional requirement of preventing starvation. We simulate
+EASY($P,\Theta=40h$) using the following policies:
+
+\begin{itemize}
+  \item All fixed $P_i$ for $1<i<10$.
+  \item The simulation-based ("full feedback") strategy.
+  \item The noisy simulation ("noisy feedback") variant.
+  \item The bandit feedback strategy with a fixed value of $\lambda=0$.
+\end{itemize}
+
+This campaign of 7 (traces) * 100 (resampling) * 13 Policies
+
 \subsubsection{Replicability}
@@
 Readers interested in replicating the experiments (or a part thereof) are
-invited to peruse to the 'artifacts description' appendix of this paper.  We
+invited to peruse the 'artifacts description' appendix of this paper. We
 provide to persistent archives containing the simulator, the analysis code and
 an automated workflow that replicates all experiments and figures present in
-this paper.
+this paper with automatic dependency management.
 

### 2017-04-01 paper/src/paper.tex
changes to expe section.

@@
 
-  \label{fig:mosn}
+  \label{fig:variability}
 \end{figure*}
@@
 reduced action space, the state space to consider is infinite. This is not
-prohibitive \textit{per se}, as modern methods~\ref{approximatereinforcement}
+prohibitive \textit{per se}, as modern methods~\cite{aprl}
 can bypass dimensionality issues via function approximation. However, these
@@
 explained in Section~\ref{sec:experiments}.
-
-%\begin{algorithm}[h]
-%  \caption{Epsilon-Greedy Bandit policy}
-%  \begin{algorithmic}[1]
-%    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
-%    \renewcommand{\algorithmicensure}{\textbf{Output:}}
-%    \REQUIRE $0 < \epsilon < 1$, $K>1$
-%    \FOR {$t = 1, 2, \ldots$}
-%    \STATE Let $i_t$ be the reordering heuristic with the highest current
-%    average reward.
-%    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
-%    $\epsilon$ use a random reordering heuristic.
-%    \ENDFOR
-%  \end{algorithmic}
-%  \label{alg:epsg}
-%\end{algorithm}
 
@@
   \item The noisy simulation ("noisy feedback") variant.
-  \item The bandit feedback strategy with a fixed value of $\lambda=0$.
+  \item The bandit feedback strategy with hyperparameters $\lambda=0$, $\Delta=1 \text{day}$
+  \item The bandit feedback strategy with hyperparameters $\lambda=0$, $\Delta=1 \text{week}$.
 \end{itemize}
 
-This campaign of 7 (traces) * 100 (resampling) * 13 Policies
+This campaign of 7 traces * 100 resampling * 13 algorithms is repeated
 
@@
 
-\begin{figure*}[h]
+\begin{figure*}[ht]
   \centering

### 2017-04-01 paper/src/paper.tex
add experiment description, draft result.

@@
 
-\bigskip
-
 Our main results are as follows: The approach based on the simulator provides
@@
 practice. In the following, we denote these policies by $P_i$ with $i = 1
-\ldots 10$.
+\ldots 12$.
 
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[]
   \caption{EASY($P,\Theta$) policy}
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[]
   \centering
@@
 Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the
-period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 10$) one reordering
+period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 12$) one reordering
 policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$
@@
 \begin{equation} \label{eq:select-exact} P = \mbox{argmin}_{P_i, 1 \le i \le
-10} w(\rightarrow T;P_i) \end{equation}
+12} w(\rightarrow T;P_i) \end{equation}
 
@@
 \label{eq:select-greedy}
-P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
+P = \mbox{argmin}_{P_i, 1 \le i \le 12} w(\rightarrow T;P_i)
 \end{equation}
 %
-\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 10$, at random (exploration mode).
+\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 12$, at random (exploration mode).
 \end{itemize}
@@
 
-\begin{table}[ht]
+\begin{table}[]
   \centering
@@
 
-We use devise the following experimental campaign. For each workload trace
+We use devise the following experimental campaign. For each of the 7 workloads
 considered, we use the original data to resample 100 traces, each of a length
@@
 the functional requirement of preventing starvation. We simulate
-EASY($P,\Theta=40h$) using the following policies:
+EASY($P,\Theta=40h$) using the following strategies:
 
 \begin{itemize}
-  \item All fixed $P_i$ for $1<i<10$.
-  \item The simulation-based ("full feedback") strategy.
-  \item The noisy simulation ("noisy feedback") variant.
-  \item The bandit feedback strategy with hyperparameters $\lambda=0$, $\Delta=1 \text{day}$
-  \item The bandit feedback strategy with hyperparameters $\lambda=0$, $\Delta=1 \text{week}$.
+
+  \item All fixed $P_i$ for $1<i<12$ from the search set defined in
+    ~\ref{subsec:policies}.  This is done for two reasons. First, it allows to
+    see which policy works well for which platform. Second, it allows to
+    compare how well strategies proposed in this work with the best
+    \textit{a-posteriori} known policy.
+
+  \item The simulation-based ("full feedback") strategy with hyperparameter
+    $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. This choice of
+    hyperparameter is a consequence of the natural rythm of human activity:
+    daily or weekly simulation periods seems to facilitate comparability of
+    performance between two periods. Indeed, arbitrary period values would
+    needlessly increase the variability of the cost metric.
+
+  \item The noisy simulation ("noisy feedback") variant with hyperparameter
+    $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
+
+  \item The bandit feedback strategy with hyperparameters $\lambda=1, \Delta
+    \in \left\{ 1 \text{day},1 \text{week} \right\}$. The choice of $\lambda=1$
+    means that we use no decay in the estimation of the overall cost. Choosing
+    to eliminate decay in this experimental campaign is natural since our
+    resampled data is in a stationary regime \textit{by construction}, as
+    explained in~\ref{ssub:resampling}.
+
 \end{itemize}
 
-This campaign of 7 traces * 100 resampling * 13 algorithms is repeated
+The experiment campaign consists of 7 traces $\times$ 100 resamples $\times$
+(12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
+two years.  This means a total simulated timespan of 25200 years. Moreover,
+both the full-feedback and noisy feedback strategies will also be calling
+simulations as part of their procedure. Indeed, these strategies re-simulate
+the system 12 times every at every period $\Delta$, which bumps the total
+effective simulation timespan to 92400 years. This motivates the use of a
+lightweight simulator in the context of this research.
 
@@
 
-\begin{figure*}[ht]
+We first attract the attention of the reader to Table~\ref{tab:full}
+
+\begin{table*}[]
+  \centering
+  \ra{1.3}
+  \caption{Average Cumulative waiting time improvement with respect to EASY-FCFS of fixed policies.}
+  \label{tab:full}
+  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
+    \hline
+    Trace        & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF \\
+    \hline
+    KTH-SP2      & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    CTC-SP2      & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    SDSC-SP2     & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    SDSC-BLUE    & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    ANL-Intrepid & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    CEA-Curie    & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    Unilu-Gaia   & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    \hline
+  \end{tabular}
+\end{table*}
+
+\begin{figure*}[]
   \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[]
   \centering
@@
 
-\begin{figure}[H]
+\begin{figure}[]
   \centering
@@
 This approach was used in two methods: a method based on a simulator and a
-method based on multi-armed bandits. The first one provides very good results,
+method based on a multi-armed bandit algorithm. The first one provides very good results,
 improving the average waiting times of the baseline FCFS ordering policy by a
@@
 %@Denis: c'est ce que font vasily et salah de deux manieres differentes.
+%TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
 As these directions are data-bound, they will require careful application of
@@
     The workflow can be executed in the current environment in parallel
-    via \lstinline[basicstyle=\ttfamily]|zymake -l localhost zymakefile|.
+    via \\ \lstinline[basicstyle=\ttfamily]|zymake -l localhost zymakefile|.
 
@@
 Readers interested with modifying the experiments can obtain the zenodo archive
-at \url{https://...}
+at \url{https://truc.tgz}
 

### 2017-04-01 paper/src/paper.tex
tex update

@@
 
-%wait mwait lpf spf sqf lqf expfact mexpfact lrf srf laf saf
 \begin{enumerate}
@@
 
+  \item An additional "Random" strategy to be used as a reference point, with
+    hyperparameter $ \Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
+    This strategy corresponds to choosing $P_i$ uniformly at random in $\left\{
+      P_1 , \cdots, P_{12} \right\}$ during every period $\Delta_t$.
+
 \end{itemize}
@@
 \label{sub:results}
-
-We first attract the attention of the reader to Table~\ref{tab:full}
 
@@
 
+We begin by attracting the attention of the reader to Table~\ref{tab:full},
+which reports the average cumulative waiting time performance improvement over
+EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. More
+precisely, let us denote the resampled traces of a given workload by $T_k$,
+where k is the index of the resampled trace ($1 \le k \le 100$). We also denote
+the number of jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by
+$\bar{w}(T_k,S)$ the cumulative waiting time obtained by a simulation run for a
+strategy $S$:
+
+\begin{equation}
+  \bar{w}(T_k,P) = \sum_{j=1}^{l(T_k)} \text{Wait}_j^{P}
+\end{equation}
+
+So that the values appearing in Table~\ref{tab:full} are exactly:
+
+\begin{equation}
+  \frac{\sum_{k=1}^{k=100}\bar{w}(T_k,\text{EASY}(P_i,40h))-
+  \bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}{\sum_{k=1}^{k=100}\bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}
+\end{equation}
+
+Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[scale=0.6]{figures/full-UniLu-Ga.pdf}
+  \caption{Evolution of the average cumulative waiting time improvement
+    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
+    UniLu-Gaia trace. The average is obtained by resampling the original trace
+    100 times. The dashed lines represent the 10th and 90th percentiles of the
+  values across this resampling.}
+  \label{fig:all}
+\end{figure}
+
+Figure~\ref{fig:all} reports the evolution of this quantity for the UniLu-Gaia
+trace, which means that it shows how the average improvement in cumulative
+waiting time evolves as a function of time. In other words, this is the average
+curve obtained if one were to plot the cumulative sum of job waiting times for
+every resampled trace, incrementing the sum when a job finishes. Accordingly,
+the final values attained by every curve correspond exactly to the values
+reported in Table~\ref{tab:full}. The dashed values represent the 10th and 90th
+percentile of the values.
+
+Observe that \textcolor{orange}{TODO}
+
 \begin{figure*}[]
@@
 
-\begin{figure}[]
+Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
+result concerning the strategies proposed in this paper. These figures show the average cumulative improvement
+of over EASY(FCFS,40h) of the three strategies we propose in the same format as Figure~\ref{fig:all}.
+There are four strategies displayed on the graph:
+
+\begin{itemize}
+  \item Full (simulated) Feedback ($\Delta=$1 day)
+  \item Noisy (simulated) Feedback ($\Delta=$1 day)
+  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 day, $\epsilon=0.5$)
+  \item Random $P \in P_i, 1 \le k \le 100$. ($\Delta=$1 day)
+\end{itemize}
+
+Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+
+\begin{table*}[]
   \centering
-  \includegraphics[scale=0.6]{figures/full-UniLu-Ga.pdf}
-
-  \caption{Evolution of the average cumulative waiting time improvement
-    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
-    UniLu-Gaia trace. The average is obtained by resampling the original trace
-    100 times. The dashed lines represent the 10th and 90th percentiles of the
-  values across this resampling.}
-
-  \label{fig:all}
-\end{figure}
+  \ra{1.3}
+  \caption{Average Cumulative waiting time improvement with respect to EASY-FCFS of fixed policies.}
+  \label{tab:strat}
+  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
+    \hline
+    Trace        & Random &      & Full &      & Noisy &      & Bandit &     \\
+    \hline
+    $\Delta$     & day    & week & day  & week & day   & week & day    & week\\
+    CTC-SP2      & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
+    SDSC-SP2     & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
+    SDSC-BLUE    & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
+    ANL-Intrepid & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
+    CEA-Curie    & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
+    Unilu-Gaia   & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
+    \hline
+  \end{tabular}
+\end{table*}
+
+Table~\ref{tab:strat} gives the performance of all variants for $\Delta=$ 1
+day and $\Delta=$ 1 week.
+
+Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
 
@@
 
+Figure~\ref{fig:mosb} gives insight regarding the behavior of the
+Noisy (simulated) Feedback policy. This figure reports the share of the policies chosen
+by the strategy on average as a function of time. This is an area plot, which means
+that the height of the colored region represents the value. Moreover, the proportion of
+times each policies are chosen by the strategy is aggregated across resampled traces.
+
+Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+
 \begin{figure}[]
@@
 \end{figure}
+
+Figure~\ref{fig:mosn} is the analogue of Figure~\ref{fig:mosb}, this time for
+the Epsilon-Greedy policy.
+
+Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
 

### 2017-04-02 paper/src/paper.tex
rework resampling section.

@@
 
-The Evaluation of computer systems performance is an intricate
-task\cite{feitbook}. Here, the two main difficulties are choosing a simulation
+The Evaluation of computer systems performance is a complex
+task. Here, the two main difficulties are choosing a simulation
 model and taking into account the variability in the average waiting times.
-This section presents our approach and provides a replicable workflow
-for the interested reader.
+This section presents our approach, which is based on lightweight simulation
+and trace resampling.
 
@@
 
-%Validating algorithms on real data is a notoriously hard
-%problem~\cite{feitbook}. The case of batch scheduling is made especially
-%difficult by the fluctuations in commonly used metrics~\cite{jsm}. Indeed, the
-%variability in performance is known to outrank the available sample sizes:
-%\textit{It is not enough to simply replay an algorithm using a trace}. For this
-%reason, we clearly expose our statistical approach.
-
-%Two main methods are available in order to generate data. On one hand, it is
-%possible to model~\cite{feitperfeval} workloads. However, this often loses
-%many specificities of the original data, and this work uses a second
-%approach based on trace resampling~\cite{feitresampling}.
-
-To carry our experiments, we need to generate many trace samples. More
-precisely, we want to sample the job submission process, which is achieved by
-splitting the traces in weeks, randomly shuffling them and picking the jobs
-submitted in the selected weeks.  This simple mechanism enables us to preserve
-the dependency within a week. Although the system state may not be independent
-from one week to an other, we can reasonably assume that it evolves under its
-stationary distribution. Thus, computing an average over the sampled traces at
-a given time can be related to a time average of an underlying Markov process,
-which converges to the true quantity thanks to the ergodic theorem.  Moreover,
-the job submission process has no long range correlation making it look like an
-independent and identically distributed process when considering sufficiently
-time spaced weeks.
+Comparing algorithms in batch scheduling is made especially difficult by the
+fluctuations in commonly used metrics~\cite{jsm}. The variability in
+performance is known to outrank the available sample sizes: \textit{It is not
+enough to simply replay an algorithm using a workload trace}. Indeed, as
+presented in Figure~\ref{fig:variability}, the objective function we are using
+in this work is subject to high variability. This variability can be so extreme
+that even a timeframe of a fews years can be insufficient to compare scheduling
+algorithms. In our experience inverting some months in the trace at random can
+lead to reversal in the results. This motivated decades of research in workload
+modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} proposes to
+resample data directly from the original traces. In this work, we use a simple
+implementation of these ideas.  More precisely, we want to sample the job
+submission process, which is achieved in the following manner: For evey week in
+a resampled trace, we iterate over every existing user in the original trace.
+For each of these users, we choose a week of the original trace uniformly at
+random. The jobs from this user in this original week are chosen to be present
+in the week being resampled.  This simple mechanism enables us to preserve the
+dependency within a week.  Although the system state may not be independent
+from one week to an other, we rely on existing work~\ref{jsspp17} and assume
+for these experiments that the system is in a stationary state. While we
+do provide guidelines as how to incorporate tracking variant of our algorithms
+in~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
+this will not be studied in the experiments from this paper, and is left for
+future research.
+
+%Thus,
+%computing an average over the sampled traces at a given time can be related to
+%a time average of an underlying Markov process, which converges to the true
+%quantity thanks to the ergodic theorem.  Moreover, the job submission process
+%has no long range correlation making it look like an independent and
+%identically distributed process when considering sufficiently time spaced
+%weeks.
 

### 2017-04-02 paper/src/paper.tex
paper modif.
fix paper
paper change

@@
-\documentclass[sigconf,review]{acmart}
+\documentclass[sigconf,review,anonymous]{acmart}
 \setcopyright{rightsretained}
@@
 \fxsetup{theme=color,mode=multiuser,layout=inline,draft}
+
+\acmConference[SC’15]{Supercomputing}{May 29 -- June 2}{Orlando, FL, USA}
 
@@
   \ra{1.3}
-  \caption{Average Cumulative waiting time improvement with respect to EASY-FCFS of fixed policies.}
+  \caption{Average total waiting time diminution of fixed policies with respect to EASY(FCFS).}
+
+  %DELTA WEEK
+
   \label{tab:full}
@@
     \hline
-    KTH-SP2      & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
-    CTC-SP2      & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
-    SDSC-SP2     & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
-    SDSC-BLUE    & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
-    ANL-Intrepid & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
-    CEA-Curie    & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
-    Unilu-Gaia   & LCFS & LPF & SPF & LQF & SQF & SEXP & LEXP & LRF & SRF & LAF & SAF\\
+    ANL-Intrepid & -25\% & -9\%           & -3\%  & \textbf{-28\%} & -2\%  & -21\% & -24\%          & -21\%          & -10\% & -19\%        & +5\%  \\
+    KTH-SP2      & -13\% & \textbf{-16\%} & +5\%  & \textbf{-16\%} & +3\%  & -8\%  & -15\%          & -8\%           & -13\% & -12\%        & +15\% \\
+    CTC-SP2      & -40\% & -19\%          & -14\% & \textbf{-47\%} & +15\% & -36\% & -19\%          & -38\%          & -12\% & -44\%        & +22\% \\
+    SDSC-SP2     & -8\%  & \textbf{-15\%} & +4\%  & -8\%           & +4\%  & -3\%  & \textbf{-15\%} & -3\%           & -10\% & -11\%        & +18\% \\
+    SDSC-BLUE    & -26\% & -29\%          & 0\%   & -29\%          & +17\% & -17\% & -26\%          & -15\%          & -15\% & \textbf{-32\%} & +23\% \\
+    CEA-Curie    & -46\% & -23\%          & -45\% & \textbf{-57\%} & -19\% & -51\% & -23\%          & \textbf{-57\%} & -24\% & -40\%        & -17\% \\
+    Unilu-Gaia   & -58\% & -33\%          & -10\% & \textbf{-62\%} & +16\% & -49\% & -27\%          & -56\%          & -16\% & -62\%        & +20\% \\
     \hline
@@
 We begin by attracting the attention of the reader to Table~\ref{tab:full},
-which reports the average cumulative waiting time performance improvement over
+which reports the average total waiting time performance improvement over
 EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. More
@@
 
-Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+The best values are outlined in bold in the table. These values give the best
+attainable performance in the search set defined in~\ref{subsec:policies}.
+Note that there seems to be some regularity; the SAF policy is remarkably
+inefficient for instance, and the LQF policy works very well. However, note
+that there are traces for which LQF is not optimal. Moreover, these
+are~\textit{a-posteriori} results, which means that we only know post-factum
+that this policy would have been better. In the present work, we are
+studying adaptive policies that obtain a good performance on any given system
+by selecting one of these policies.
 
@@
   \centering
-  \includegraphics[scale=0.6]{figures/full-UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{figures/full-ANL-Intr.pdf}
   \caption{Evolution of the average cumulative waiting time improvement
@@
 
-Figure~\ref{fig:all} reports the evolution of this quantity for the UniLu-Gaia
-trace, which means that it shows how the average improvement in cumulative
-waiting time evolves as a function of time. In other words, this is the average
-curve obtained if one were to plot the cumulative sum of job waiting times for
-every resampled trace, incrementing the sum when a job finishes. Accordingly,
-the final values attained by every curve correspond exactly to the values
-reported in Table~\ref{tab:full}. The dashed values represent the 10th and 90th
-percentile of the values.
-
-Observe that \textcolor{orange}{TODO}
+Figure~\ref{fig:all} reports the evolution of the average cumulative waiting
+time for the UniLu-Gaia tis quantityrace, which means that it shows how the
+average improvement in cumulative waiting time evolves as a function of time.
+In other words, this is the average curve obtained if one were to plot the
+cumulative sum of job waiting times for every resampled trace, incrementing the
+sum when a job finishes. Accordingly, the final values attained by every curve
+correspond exactly to the values reported in Table~\ref{tab:full}. The dashed
+values represent the 10th and 90th percentile of the values.
+
+Observe that while the 10th and 90th percentiles of the improvement with
+respect to EASY-FCFS are large, the values are still contained enough to make
+definite statements about how these policies compare to the FCFS baseline.
+This figure does not report the variability of the policies against one another,
+and does not allow to compare them beyong comparing their average performance.
+However, the fact that the band of percentiles are roughly of the size of half
+the best attainable improvement was our original motivation for this work, as
+one can expect to be able to distinguish between these policies in an
+online manner.
 
@@
 Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
-result concerning the strategies proposed in this paper. These figures show the average cumulative improvement
-of over EASY(FCFS,40h) of the three strategies we propose in the same format as Figure~\ref{fig:all}.
-There are four strategies displayed on the graph:
+result concerning the strategies proposed in this paper. These figures show the
+average cumulative improvement of over EASY(FCFS,40h) of the three strategies
+we propose in the same format as Figure~\ref{fig:all}.  There are four
+strategies displayed on the graph:
 

### 2017-04-02 misc/visu/cumobj_diffplot.R
dotted line

@@ -93,7 +93,7 @@ print(momi)
 pdf(args$o,width=5.5, height=4)
 p=ggplot(dfs,aes(x=time-tmi,y=mean_obj,ymin=sdn,ymax=sdp,fill=name,color=name),)+
   geom_line()+
-  geom_ribbon(alpha=0.03,linetype="dashed")+
+  geom_ribbon(alpha=0.03,linetype="dotted")+
   expand_limits(x = c(tmi, tma+(tma-tmi)*0.15))+
   #scale_x_continuous(limits=c(min(timestamp_to_date(dfs$time)),max(timestamp_to_date((dfs$time))-min(timestamp_to_date(dfs$time)))))+
   scale_x_continuous(breaks = brkx, labels = labx) +

### 2017-04-02 paper/src/paper.tex
expe fix
advance paper

@@
-\documentclass[sigconf,review,anonymous]{acmart}
+\documentclass[sigconf,anonymous]{acmart}
 \setcopyright{rightsretained}
@@
 
-Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+Table~\ref{tab:strat} reports the average total waiting time performance
+improvement over EASY(FCFS,40h) of all the proposed strategies along with the
+random variant.
+
+We make the following observations.
+
+\paragraph
+\textbf{FCFS vs Random.} Note that randomly taking a policy in the search
+set defined in ~\ref{subsec:policies} gives already a much better performance
+than EASY(FCFS). This is due
+
+\paragraph
+\textbf{Simulated Feedback finds the best policy}
+\paragraph
+\textbf{Noisy simulations are useful.}
+\paragraph
+\textbf{Full vs Bandit feedback}
 
@@
   \centering
-  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{figures/mosaicbandit-ANL-Intr.pdf}
 
@@
   \centering
-  \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}
+  \includegraphics[scale=0.6]{figures/mosaic-ANL-Intr.pdf}
 

### 2017-04-02 paper/src/paper.tex
paper update

@@
 \cite{Tokic:2010}. This strategy has however not been beneficial in our case.
-We believe this is due to the properties of the traces we use, as
-explained in Section~\ref{sec:experiments}.
+We believe this is due to the properties of the traces we use, as explained in
+Section~\ref{sec:experiments}. The $\epsilon$-greedy strategy is known to be
+outperformed in terms of regret by other classical algorithms such as the UCB
+or EXP family of algorithms. See~\cite{bubnow} for a precise definition and
+thourough treatment of this subject. These algorithms however all rely on
+pre-normalized rewards. Here, the scale of $w_{\epsilon}(t;P_i)$ is not known
+in advance. While the usual doubling tricks can be devised to go around
+sequential issues of scale with good asymptotical performance, they ruin
+performance in practice. Accordingly, we only perform experiments using the
+simple $\epsilon$-greedy method.
 
@@
 dependency within a week.  Although the system state may not be independent
-from one week to an other, we rely on existing work~\ref{jsspp17} and assume
+from one week to an other, we rely on existing work~\cite{jsspp17} and assume
 for these experiments that the system is in a stationary state. While we
@@
 average cumulative improvement of over EASY(FCFS,40h) of the three strategies
-we propose in the same format as Figure~\ref{fig:all}.  There are four
+we propose in the same format as Figure~\ref{fig:all}. There are five
 strategies displayed on the graph:
@@
 \begin{itemize}
-  \item Full (simulated) Feedback ($\Delta=$1 day)
-  \item Noisy (simulated) Feedback ($\Delta=$1 day)
-  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 day, $\epsilon=0.5$)
-  \item Random $P \in P_i, 1 \le k \le 100$. ($\Delta=$1 day)
+  \item Random $P \in P_i, 1 \le k \le 100$. ($\Delta=$1 week)
+  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 week)
+  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 day)
+  \item Full (simulated) Feedback ($\Delta=$1 week)
+  \item Noisy (simulated) Feedback ($\Delta=$1 week)
 \end{itemize}
-
-Table~\ref{tab:strat} reports the average total waiting time performance
-improvement over EASY(FCFS,40h) of all the proposed strategies along with the
-random variant.
-
-We make the following observations.
-
-\paragraph
-\textbf{FCFS vs Random.} Note that randomly taking a policy in the search
-set defined in ~\ref{subsec:policies} gives already a much better performance
-than EASY(FCFS). This is due
-
-\paragraph
-\textbf{Simulated Feedback finds the best policy}
-\paragraph
-\textbf{Noisy simulations are useful.}
-\paragraph
-\textbf{Full vs Bandit feedback}
 
@@
 
-Table~\ref{tab:strat} gives the performance of all variants for $\Delta=$ 1
-day and $\Delta=$ 1 week.
-
-Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+Table~\ref{tab:strat} reports the average total waiting time performance
+improvement over EASY(FCFS,40h) of all the proposed strategies along with the
+random variant.
+
+We make the following observations.
+
+\textbf{FCFS vs Random.} Randomly taking a policy in the search set defined in
+~\ref{subsec:policies} is better than using the FCFS order. This is due to the
+known fact that the FCFS order is not optimal for optimizing the average
+waiting time objective, and confirms that the search set is well chosen.  The
+value of the period length $\Delta$ has an impact on the performance of the
+random policy. Indeed, the shorter period leads to better performance. We
+suggest that this is due to the properties of the system under study. Indeed,
+when a bad policy is used, the job queue builds up. By switching more
+frequently from bad to good policies, the system never has time to build up the
+job queue. In other words, while bad decisions impact the system through an
+accumulation of waiting jobs, there is less contribution to system inertia from
+good decisions, as the number of jobs in queue is below bounded by zero. This
+interpretation also means that randomized bandit strategies are robust by
+design, and needs more subsequent experimenting to be validated.
+
+\textbf{The Simulated Feedback strategy consistently performs as well as the
+best fixed policy.} The values attained by the simulation-based strategy match
+the highlighted values from Table~\ref{tab:full}. This corresponds to the curve
+labeled "Full" in Figures~\ref{fig:small} and~\ref{fig:follow}. Moreover,
+these average cumulative waiting time curves exhibit linearity (as opposed to
+being concave), which further shows that the choice of the best policy can be
+made rapidly.  Note that the value of $\Delta$ has no impact over the
+performance of this strategy.
+
+\textbf{Noisy simulations can be used.} The performance attained by the noisy
+variant almost matches that of the simulation-based strategy. This indicates
+that it is not necessary to have access to a precise simulator. Rather, we may
+hope that an imprecise but unbiased reporting of the performance is enough to
+rapidly select a reordering policy .
+
+\textbf{Full vs Bandit feedback.} As expected, the bandit strategy evolves
+between the random strategy and the full feedback strategy.
+It is able to recover between \textcolor{orange}{X\%} to \textcolor{orange}{X\%}
+of the Random-FullFeedback gap.
 
@@
 
-Figure~\ref{fig:mosb} gives insight regarding the behavior of the
-Noisy (simulated) Feedback policy. This figure reports the share of the policies chosen
-by the strategy on average as a function of time. This is an area plot, which means
-that the height of the colored region represents the value. Moreover, the proportion of
-times each policies are chosen by the strategy is aggregated across resampled traces.
-
-Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+Figure~\ref{fig:mosb} gives insight regarding the behavior of the Noisy
+(simulated) Feedback policy with $\Delta=$ 1 week. This figure reports the
+share of the policies chosen by the strategy on average as a function of time.
+This is an area plot, which means that the height of the colored region
+represents the value. Moreover, the proportion of times each policies are
+chosen by the strategy is aggregated across resampled traces.
+
+While this policy obtains an excellent performance, the convergence to a fixed
+policy is slow.  This is not harmful \textit{per se} and just means that the
+variability in the waiting times is indeed high. Additionally, we observe that
+the increasingly dominant policy, LCFS, is the second best overall policy. This
+is perhaps due to the estimation bias stemming from the boundary conditions of
+the weekly simulations.
 
@@
 Figure~\ref{fig:mosn} is the analogue of Figure~\ref{fig:mosb}, this time for
-the Epsilon-Greedy policy.
-
-Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.
+the Epsilon-Greedy policy with $\Delta$ = 1 week.
+
+Observe that this time the choice is much less stable. The convergence is
+extremely slow, which is expected given the variability of the reward we give
+to the bandit algorithm.
 

### 2017-04-02 paper/src/paper.tex
table input.

@@
 work on non simulated traces, we randomly introduce noise in the estimate of
-the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either
-down or up, for each job:
+the policy cost defined by Eq.~\ref{eq:cost-exact} by rescaling it, either
+down or up:
 %
 \begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
-\sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i} \end{equation}
-%
-where $\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a
-+/-15\% noise on the waiting time estimated by the simulator. The overall cost
+ \rho \sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
+%
+where $\rho$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
++/-20\% noise on the waiting time estimated by the simulator. The overall cost
 is then defined in the same way as before, leading to:
@@
     $\Delta$     & day    & week & day  & week & day   & week & day    & week\\
-    CTC-SP2      & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
-    SDSC-SP2     & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
-    SDSC-BLUE    & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
-    ANL-Intrepid & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
-    CEA-Curie    & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
-    Unilu-Gaia   & LCFS   & LPF  & SPF  & LQF  & SQF   & SEXP & LEXP   & LRF \\
+    CTC-SP2      & -20    &      & -44  &      & -44   &      & -26    & \\
+    KTH-SP2      & -6     &      & -12  &      & -12   &      & -7     & \\
+    SDSC-SP2     & -4     &      & -13  &      & -13   &      & -5     & \\
+    SDSC-BLUE    & -12    &      & -31  &      & -30   &      & -12    & \\
+    ANL-Intrepid & -13    &      & -22  &      & -19   &      & -13    & \\
+    CEA-Curie    & -35    &      & -53  &      & -53   &      & -36    & \\
+    Unilu-Gaia   & -29    &      & -59  &      & -58   &      & -33    & \\
     \hline

### 2017-04-02 paper/src/paper.tex
changes for release.

@@
 
-\acmConference[SC’15]{Supercomputing}{May 29 -- June 2}{Orlando, FL, USA}
+\acmConference[SC’17]{Supercomputing}{May 29 -- June 2}{Orlando, FL, USA}
 
@@
 means that we are able to recover in an online fashion existing results that
-divide average waiting times by factor of \textcolor{orange}{X} to
-\textcolor{orange}{X}. Indeed, we show that the required sample sizes for
-policy selection in batch schedulers are small enough for workload resampling to
-be avoided in practice. The variant based on noisy data, which is closer to
-actual conditions of uncertainties on both jobs and platform, also behaves very
-well. The alternate approach based on a multi-armed bandit also provides
-reasonable results at a much lower cost, dividing the average waiting times by
-a factor of \textcolor{orange}{X} to \textcolor{orange}{X}.
+reduce average waiting times between -11\% and -60\% compared to EASY-FCFS.
+Indeed, we show that the required sample sizes for policy selection in batch
+schedulers are small enough for workload resampling to be avoided in practice.
+The variant based on noisy data, which is closer to actual conditions of
+uncertainties on both jobs and platform, also behaves very well. The alternate
+approach based on a multi-armed bandit also provides reasonable results at a
+much lower cost, reducing the average waiting times down to between -8\% and
+-46\% compared to EASY-FCFS.
 
@@
   \ra{1.3}
-  \caption{Average Cumulative waiting time improvement with respect to EASY-FCFS of fixed policies.}
+  \caption{Average Cumulative waiting time improvement of the proposed policies with respect to EASY(FCFS).}
   \label{tab:strat}
@@
     \hline
-    $\Delta$     & day    & week & day  & week & day   & week & day    & week\\
-    CTC-SP2      & -20    &      & -44  &      & -44   &      & -26    & \\
-    KTH-SP2      & -6     &      & -12  &      & -12   &      & -7     & \\
-    SDSC-SP2     & -4     &      & -13  &      & -13   &      & -5     & \\
-    SDSC-BLUE    & -12    &      & -31  &      & -30   &      & -12    & \\
-    ANL-Intrepid & -13    &      & -22  &      & -19   &      & -13    & \\
-    CEA-Curie    & -35    &      & -53  &      & -53   &      & -36    & \\
-    Unilu-Gaia   & -29    &      & -59  &      & -58   &      & -33    & \\
+    $\Delta$     & week   & day   & week  & day   & week  & day   & week   & day \\
+    CTC-SP2      & -20\%  & -23\% & -44\% & -44\% & -44\% & -44\% & -26\%  & -31\% \\
+    KTH-SP2      & -6\%   & -8\%  & -12\% & -11\% & -12\% & -12\% & -7\%   & -10\% \\
+    SDSC-SP2     & -4\%   & -7\%  & -13\% & -13\% & -13\% & -13\% & -5\%   & -8\% \\
+    SDSC-BLUE    & -12\%  & -14\% & -31\% & -31\% & -30\% & -29\% & -12\%  & -17\% \\
+    ANL-Intrepid & -13\%  & -20\% & -22\% & -26\% & -19\% & -24\% & -13\%  & -20\% \\
+    CEA-Curie    & -35\%  & -42\% & -53\% & -47\% & -53\% & -48\% & -36\%  & -46\% \\
+    Unilu-Gaia   & -29\%  & -31\% & -59\% & -60\% & -58\% & -58\% & -33\%  & -34\% \\
     \hline
@@
 \textbf{Full vs Bandit feedback.} As expected, the bandit strategy evolves
-between the random strategy and the full feedback strategy.
-It is able to recover between \textcolor{orange}{X\%} to \textcolor{orange}{X\%}
-of the Random-FullFeedback gap.
+between the random strategy and the full feedback strategy. The bandit-based
+strategy always fares better with a shorter period $\Delta$, recovering between
+a quarter to the whole gap between the Random and Full Feedback strategies.
 
@@
 method based on a multi-armed bandit algorithm. The first one provides very good results,
-improving the average waiting times of the baseline FCFS ordering policy by a
-factor of \textcolor{orange}{X} to \textcolor{orange}{X} The second one is
-slighly less efficient with an improvement factor of  \textcolor{orange}{X} to
-\textcolor{orange}{X}. However, it is more easy to use and less costly to run.
+reducing the average waiting times of the baseline FCFS ordering policy by a
+factor of -11\% to -60\%. The second one is
+slighly less efficient with an improvement factor of -8\% to
+-46\%. However, it is easier to use and less costly to run.
 
@@
      simulator specially written for this work. this simulator is made
-     available under the isc~\cite{isc} license both as a persistent zenodo
-     archive at~\cite{ocst} and as a git repository at ~\cite{ocstgit}.
+     available under the isc~\cite{isc} license as a persistent zenodo
+     archive at~\cite{ocst}.
 

### 2017-04-03 paper/src/paper.tex
fix paper

@@
 and jobs can be discontinuously mapped to any available processor on the
-system. We develop a lightweight simulator\cite{ocst} and an accompanying
-multi-armed bandit library\cite{obandit}. See the reproducibility paragraph
-below for more information.  With this simulator, we are able to replay
-EASY-FCFS on the CEA-CURIE trace in 33 seconds with a machine equipped with an
-Intel(R) Core(TM) i5-4310U CPU @ 2.00GHz. As a point of comparison, the Batsim
-simulator with no network communication modeling and no resource contiguity
-takes more than 7 hours to replay the same trace on the same hardware. The
-complete simulation campaign presented below is executed on a single Dell
-PowerEdge R730 in 22 hours, including input preparation and analysis code.
+system. We develop a lightweight simulator\cite{ocst} and release it to the
+community. See the reproducibility paragraph below for more information.  With
+this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in 33
+seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
+2.00GHz. As a point of comparison, the Batsim simulator with no network
+communication modeling and no resource contiguity takes more than 7 hours to
+replay the same trace on the same hardware. The complete simulation campaign
+presented below is executed on a single Dell PowerEdge R730 in 22 hours,
+including input preparation and analysis code.
 
@@
 We use devise the following experimental campaign. For each of the 7 workloads
-considered, we use the original data to resample 100 traces, each of a length
+considered, we use the original data to resample 60 traces, each of a length
 of two years. For each resampled trace, we run simulations using the
@@
 
-The experiment campaign consists of 7 traces $\times$ 100 resamples $\times$
+The experiment campaign consists of 7 traces $\times$ 60 resamples $\times$
 (12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
@@
 precisely, let us denote the resampled traces of a given workload by $T_k$,
-where k is the index of the resampled trace ($1 \le k \le 100$). We also denote
+where k is the index of the resampled trace ($1 \le k \le 60$). We also denote
 the number of jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by
@@
 \begin{equation}
-  \frac{\sum_{k=1}^{k=100}\bar{w}(T_k,\text{EASY}(P_i,40h))-
-  \bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}{\sum_{k=1}^{k=100}\bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}
+  \frac{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(P_i,40h))-
+  \bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}
 \end{equation}
@@
     UniLu-Gaia trace. The average is obtained by resampling the original trace
-    100 times. The dashed lines represent the 10th and 90th percentiles of the
+    60 times. The dashed lines represent the 10th and 90th percentiles of the
   values across this resampling.}
@@
     compared to EASY-FCFS of the FullFeedback, NoisyFeedback and EpsilonGreedy
-    policies. The average is obtained by resampling the original trace 100
+    policies. The average is obtained by resampling the original trace 60 
     times. The dashed lines represent the 10th and 90th percentiles of the
@@
 \begin{itemize}
-  \item Random $P \in P_i, 1 \le k \le 100$. ($\Delta=$1 week)
+  \item Random $P \in P_i, 1 \le k \le 60$. ($\Delta=$1 week)
   \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 week)
@@
   \caption{Share of the policies chosen by Epsilon-Greedy as a function of
-  time. The average choice is obtained by resampling the original trace 100
+  time. The average choice is obtained by resampling the original trace 60 
 times and aggregating by date.}
@@
   \caption{Share of the policies chosen by the Noisy policy. The average choice
-  is obtained by resampling the original trace 100 times and aggregating by
+  is obtained by resampling the original trace 60 times and aggregating by
   date.}
@@
 All the code associated with this work is released under the ~\cite{isc}
-license as a persistent zenodo~\cite{zenodo} archive at~\cite{zenodomain}.
+license as a persistent zenodo~\cite{zenodo} archive~\cite{repro}.
 
@@
 
-\textbf{one-liner : replication}
-
-Running the experiments can be done on any platform equipped with the nix
-package manager by running:
+The archive containing the experimental workflow can be obtained at
+\url{(omitted due to double-blind review)} Running the experiments can be done
+on any platform equipped with the nix package manager by running at the root of
+the extracted archive:
 
 \begin{lstlisting}
-nix-build (fetchTarball ...)
+nix-build -A banditSelection
 \end{lstlisting}
@@
 All the software dependencies of our experiments are automatically managed by Nix.
+The Nix software is available at \url{https://nixos.org/nix}.
 
@@
 Nevertheless, the reader interested in using the tools outside the workflow can
-obtain an environment by running the
-\lstinline[basicstyle=\ttfamily]|nix-shell| command:
-
-\begin{lstlisting}
-nix-shell (fetchTarball ...)
-\end{lstlisting}
-
-Or alternatively, if the user is at the root of the extracted archive:
+obtain an environment by running a \lstinline[basicstyle=\ttfamily]|nix-shell| command
+at the root of the extracted archive:
 
@@
 sizes) can be obtained by modifying the
-\lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|. For modification
+\lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|.
 

### 2017-04-03 paper/src/paper.tex
paper fix

@@
   simulator) recovers existing in-hindsight results that allow to divide the
-  average waiting time up to a factor of 2. Moreover, we show that bandit
-  feedback can be used by a simple multi-armed bandit algorithm to decrease the
-  average waiting time down to 40\% of its original value without using a
-  simulator.
+  average waiting time by almost 2. Moreover, we show that bandit feedback can
+  be used by an epsilon-greedy algorithm to cut the average waiting time by
+  40\% without using a simulator.
 
@@
 management software still has a long way to go. The existing job and resource
-management softwares allow to run tens of thousands of jobs on hundreds of thousands
+management software allow to run tens of thousands of jobs on hundreds of thousands
 of cores. They are based on robust policies, however, despite their positive
 properties such as relative simplicity and prevention of job starvation,
-there is still room for improvement. We propose here to study how to tune the
+there is still room for improvement. Here, we propose to study how to tune the
 ordering of the submission queues in the frame of the classical
@@
 
-Contrarily to most existing studies, which consider the learning of specific
+Unlike most existing studies, which consider the learning of specific
 parameters of a trace, we provide a complete system that uses feedback from the
@@
 Our main contribution is to investigate two original methods for learning a
-good scheduling policy on a per-system basis. We focus in this work on the
+good scheduling policy on a per-system basis. In this work, we focus on the
 average waiting time of the jobs. Each method aims to choose the best queue
@@
 The second method (bandit feedback) is a simulator-less approach based on a
-multi-armed bandit algorithm that derives feedback by measuring system
+multi-armed bandit algorithm, which derives feedback by measuring system
 performance.
 
-Both of these methods are evaluated with a specially developped simulator that is
-open-sourced and packaged for the community.
+Both of these methods are evaluated with a specifically developed simulator,
+which is open-sourced and packaged for the community.
 
 Our main results are as follows: The approach based on the simulator provides
-very good results, close to the best policy in the search set. In essence, this
-means that we are able to recover in an online fashion existing results that
-reduce average waiting times between -11\% and -60\% compared to EASY-FCFS.
-Indeed, we show that the required sample sizes for policy selection in batch
-schedulers are small enough for workload resampling to be avoided in practice.
-The variant based on noisy data, which is closer to actual conditions of
-uncertainties on both jobs and platform, also behaves very well. The alternate
-approach based on a multi-armed bandit also provides reasonable results at a
-much lower cost, reducing the average waiting times down to between -8\% and
--46\% compared to EASY-FCFS.
+very good results, close to the best policy in the search set. In essence, we
+are able to recover in an online fashion existing results that reduce average
+waiting times by 11\% to 60\% compared to EASY-FCFS.  Indeed, we show
+that the required sample sizes for policy selection in batch schedulers are
+small enough for workload resampling to be avoided in practice.  The variant
+based on noisy data, which is closer to the actual conditions of uncertainties on
+both jobs and platform, also behaves very well. The alternate approach based on
+a multi-armed bandit also provides reasonable results at a much lower cost,
+reducing the average waiting times by 8\% to 46\% compared to
+EASY-FCFS.
 
@@
 strategy, it can absolutely be outperformed.  We show how to automate selection
-of a good alternative policy in two cases (depending whether a simulator is
+of a good alternative policy in two cases (depending on whether a simulator is
 available or not), and provide the difference in performance to be expected
@@
 
-This section presents and discuss the most significant results related to job
+This section presents and discusses the most significant results related to job
 scheduling and learning algorithms in this context.  Let us start by recalling
@@
 whose practical ramifications, varying hypotheses, and inherent uncertainty of
-the problem applied on the HPC field have driven practitioners to use simple
+the problem applied to the HPC field have driven practitioners to use simple
 heuristics (and researchers to study their behavior). The two most popular
@@
 while the basic mechanism is used by some actual resource and job management
-softwares (most notably SLURM~\cite{SLURMdocSCHED}), this is rarely done
-without fine tunings by system administrators.
+software (most notably SLURM~\cite{SLURMdocSCHED}), this is rarely done
+without fine tuning by system administrators.
 
@@
 The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
-method to tuning these queues, although it requires simulated scheduling runs
+method to tunine these queues, although it requires simulated scheduling runs
 at decision time and therefore, it costs much more than the natural execution of EASY.
@@
 estimate uncertain parameters, we consider here a more pragmatic approach,
-which directly learn a good scheduling policy from a given policy space.
+which directly learns a good scheduling policy from a given policy space.
 
 Existing work~\cite{jsspp17} takes this approach in an offline manner, by
-splitting workload data in a training/testing fashion. This work studied
+splitting workload data in a training/testing fashion. This work studies
 EASY-Backfilling and shows that the relative performance of some well-known
 priority orders for starting jobs differs between workloads, but is relatively
-stable troughout time for a given workload.
+stable over time for a given workload.
 
@@
 observed. The goal of a MAB algorithm is to maximize the total reward obtained
-in a successive number of rounds.  There exist a bunch of works that address
+in a successive number of rounds. There exist multiple works addressing
 this problem under a variety of constraints, The two most popular settings are
@@
 Confidence Bound" family of algorithms. According to our knowledge, the work
-of~\cite{Banos} is the earliest known work to us for the adversarial case
+of~\cite{Banos} is the earliest for the adversarial case;
 see~\cite{nonstoch} for the "Exponential-weight" family of algorithms. We refer
-to the review~\cite{bubnow} for a comprehensive overview of the field. While
+to~\cite{bubnow} for a comprehensive review of the field. While
 these algorithms bound the cumulative difference in loss to the best arm (the
@@
 
-The selection of the job to run is performed according to a scheduling policy
-that establishes the order in which the jobs are executed.  EASY-Backfilling is
+The selection of the job to run is performed according to a scheduling policy,
+which determines the order in which the jobs are executed. EASY-Backfilling is
 the most widely used policy due to its simple and robust implementation and
@@
 The EASY-FCFS heuristic uses a job queue to select and backfill jobs.  At any
-time that requires a scheduling decision (i.e. job submission or termination),
+time requiring a scheduling decision (i.e. job submission or termination),
 the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
-and starts jobs until it finds a job that can not be started right away. It
-then makes a reservation for this job at the earliest predictable time and
+and starts jobs until it finds a job that can not be started right away. Then,
+it makes a reservation for this job at the earliest predictable time and
 starts \textit{backfilling} the job queue in FCFS order, starting any job that
@@
 
-We present here the strategies we have retained for selecting a policy. We will
+Here, we present the strategies we have retained for selecting a policy. We will
 refer to the period during which a selected policy is applied as the
@@
 We further assume that there is a certain regularity among periods,
-\textit{i.e.} that the distributions underlying the jobs submitted do not
+\textit{i.e.} that the distribution of the job submission process does not
 radically differ between consecutive periods. This assumption is validated in
@@
 
-We thus explore here a more efficient strategy that dispenses with estimating
+Thus, we explore here a more efficient strategy that dispenses with estimating
 the cost of all policies at a given time and only updates the cost of the
@@
 $\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
-\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary here to ensure that policies
+\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary to ensure that policies
 remain comparable over time, as not normalizing would favor the policy that was
@@
 comprehensive experimental campaign.  Subsection~\ref{sub:protocol} outlines
-the experimental protocol used and Subsection~\ref{sub:results} contains the
+our experimental protocol and Subsection~\ref{sub:results} contains the
 experimental results and their discussion.
@@
 
-The Evaluation of computer systems performance is a complex
+The evaluation of computer systems performance is a complex
 task. Here, the two main difficulties are choosing a simulation
@@
 seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
-2.00GHz. As a point of comparison, the Batsim simulator with no network
-communication modeling and no resource contiguity takes more than 7 hours to
+2.00GHz. As a comparison, the Batsim simulator with network
+communication modeling disabled and no resource contiguity takes more than 7 hours to
 replay the same trace on the same hardware. The complete simulation campaign
@@
 do provide guidelines as how to incorporate tracking variant of our algorithms
-in~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
+in Section~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
 this will not be studied in the experiments from this paper, and is left for
@@
 We use devise the following experimental campaign. For each of the 7 workloads
-considered, we use the original data to resample 60 traces, each of a length
+considered, we use the original data to resample 60 traces, each with a length
 of two years. For each resampled trace, we run simulations using the
@@
 threshold $\Theta$ at 40 hours, a sane default. The choice of this default
-stems from the analysis developped in ~\cite{jsspp17}. This value is high
+stems from the analysis developed in ~\cite{jsspp17}. This value is high
 enough for queue reorderings to be leveraged while being low enough to provide
@@
     hyperparameter is a consequence of the natural rythm of human activity:
-    daily or weekly simulation periods seems to facilitate comparability of
+    daily or weekly simulation periods seem to facilitate comparability of
     performance between two periods. Indeed, arbitrary period values would
@@
 
-The experiment campaign consists of 7 traces $\times$ 60 resamples $\times$
+The experimental campaign consists of 7 traces $\times$ 60 resamples $\times$
 (12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
@@
 
-We begin by attracting the attention of the reader to Table~\ref{tab:full},
-which reports the average total waiting time performance improvement over
-EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. More
-precisely, let us denote the resampled traces of a given workload by $T_k$,
-where k is the index of the resampled trace ($1 \le k \le 60$). We also denote
-the number of jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by
-$\bar{w}(T_k,S)$ the cumulative waiting time obtained by a simulation run for a
-strategy $S$:
+We begin by drawing the attention of the reader to Table~\ref{tab:full}, which
+reports the average total waiting time performance improvement over
+EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. Let
+us denote the resampled traces of a given workload by $T_k$, where k is the
+index of the resampled trace ($1 \le k \le 60$). We also denote the number of
+jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by $\bar{w}(T_k,S)$
+the cumulative waiting time obtained by a simulation run for a strategy $S$:
 
@@
 The scheduling of parallel jobs on a given HPC platform is a very hard
-optimization problem with many uncertain parameters. Even thought these
+optimization problem with many uncertain parameters. Even though these
 uncertainties could be reduced, determining efficient strategies remains
@@
 
-More precisely, we showed that it is worth to learn on the scheduling process
+More precisely, we showed that it is worth learning on the scheduling process
 itself. Indeed, reordering the submission queues under EASY-BackFilling leads
@@
 This approach was used in two methods: a method based on a simulator and a
-method based on a multi-armed bandit algorithm. The first one provides very good results,
-reducing the average waiting times of the baseline FCFS ordering policy by a
-factor of -11\% to -60\%. The second one is
-slighly less efficient with an improvement factor of -8\% to
--46\%. However, it is easier to use and less costly to run.
-
-There are two directions for extensions of this work.  First, the question of
+method based on a multi-armed bandit algorithm. The first one provides very
+good results, reducing the average waiting times of the baseline FCFS ordering
+policy by 11\% to 60\%. The second one is slighly less efficient with an
+improvement factor of 8\% to 46\%. However, it is easier to use and cheaper
+to run.
+
+There are two directions to extend this work.  First, the question of
 the existence of a more efficient selection strategy for bandit feedback for

### 2017-04-03 paper/src/paper.tex
tex upd

@@
 the answer to this challenge is to build extreme-scale parallel and distributed
-platforms. The never-ending race for more computing power and storage capacity does
-not only lead to sophisticated specific Exascale platforms, but the objective
-of the community is to design efficient sustained Petascale platforms. This
-large-scale evolution and the increasing complexity in both architecture and
-applications create many scientific and technical problems. Accordingly, system
-management software still has a long way to go. The existing job and resource
-management software allow to run tens of thousands of jobs on hundreds of thousands
-of cores. They are based on robust policies, however, despite their positive
-properties such as relative simplicity and prevention of job starvation,
-there is still room for improvement. Here, we propose to study how to tune the
-ordering of the submission queues in the frame of the classical
-EASY-BackFilling family of heuristics using two new online learning approaches
+platforms. The never-ending race for more computing power and storage capacity
+does not only lead to sophisticated specific Exascale platforms, but the
+objective of the community is to design efficient sustained Petascale
+platforms. This large-scale evolution and the increasing complexity in both
+architecture and applications create many scientific and technical problems.
+Accordingly, system management software still has a long way to go. The
+existing job and resource management software allow to run tens of thousands of
+jobs on hundreds of thousands of cores. These software are based on robust
+policies with positive properties such as relative simplicity and prevention of
+job starvation. However, there is still room for improvement. Here, we propose
+to study how to tune the ordering of the submission queues in the frame of the
+classical EASY-BackFilling family of heuristics using two new online approaches
 (an approach based on a simulator and an approach using multi-armed bandit).
@@
 sophisticated approaches developed \textit{ad hoc} supervised machine learning
-algorithms for reducing uncertainty in problem parameters (e.g. job execution
-times, memory requirements, etc.) via prediction. However, only few studies
+algorithms for reducing uncertainty in problem parameters (e.g.\ job execution
+times, memory requirements, etc.). However, only few studies
 evaluate the impact of such techniques on the performances of the resource
@@
 heuristics for HPC platforms are EASY~\cite{easy} and
-Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.
+Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfilling.
 
@@
 The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
-method to tunine these queues, although it requires simulated scheduling runs
+method to tuning these queues, although it requires simulated scheduling runs
 at decision time and therefore, it costs much more than the natural execution of EASY.
@@
 large scale computing platforms in order to improve system performance. Some
-works use collaborative filtering to colocate tasks in clouds by estimating
+works use collaborative filtering to co-locate tasks in clouds by estimating
 application interference~\cite{7516031}. Some others are closer to the
@@
 known benefits such as high system utilization~\cite{easy}. This strategy has
-no worst case guarantee beyond the absence of starvation (i.e. every job will
+no worst case guarantee beyond the absence of starvation (i.e.\ every job will
 be scheduled at some moment).
@@
 The EASY-FCFS heuristic uses a job queue to select and backfill jobs.  At any
-time requiring a scheduling decision (i.e. job submission or termination),
+time requiring a scheduling decision (i.e.\ job submission or termination),
 the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
@@
 We further assume that there is a certain regularity among periods,
-\textit{i.e.} that the distribution of the job submission process does not
+i.e.\ that the distribution of the job submission process does not
 radically differ between consecutive periods. This assumption is validated in
@@
 policies considered. This indicates that when when the cost metric is averaged
-over different periods, there is a tradeoff to find in between longer periods
+over different periods, there is a trade-off to find in between longer periods
 that would somehow limit the variability, and shorter ones that yield more
@@
 $\epsilon$-greedy exploration method, standard in reinforcement learning
-\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to tradeoff
+\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to trade-off
 exploitation and exploration, and relies, as before, on past estimates of the
@@
 or EXP family of algorithms. See~\cite{bubnow} for a precise definition and
-thourough treatment of this subject. These algorithms however all rely on
+thorough treatment of this subject. These algorithms however all rely on
 pre-normalized rewards. Here, the scale of $w_{\epsilon}(t;P_i)$ is not known
@@
 The choice of a simulator is a critical part of experimental validation that
-raises a tradeoff between precision and runtime. High precision can be obtained
+raises a trade-off between precision and runtime. High precision can be obtained
 by carefully modeling the platform and its network topology, or extracting
-information from the jobs from their sources or post-morted logs. This is the
+information from the jobs from their sources or post-mortem logs. This is the
 approach used by high-fidelity batch scheduling simulators such as
 Batsim~\cite{batsim}. In the present work, the focus is on studying the
-EASY-Backfilling mechanism in itself, without adressing the allocation problem.
+EASY-Backfilling mechanism in itself, without addressing the allocation problem.
 Moreover, the experimental protocol used here requires many simulation runs.
-Therefore, we set the precision/runtime tradeoff at the point which minimizes
+Therefore, we set the precision/runtime trade-off at the point which minimizes
 simulation runtimes. We discard all topological information relative to the
@@
 in this work is subject to high variability. This variability can be so extreme
-that even a timeframe of a fews years can be insufficient to compare scheduling
+that even a time-frame of a fews years can be insufficient to compare scheduling
 algorithms. In our experience inverting some months in the trace at random can
@@
 implementation of these ideas.  More precisely, we want to sample the job
-submission process, which is achieved in the following manner: For evey week in
+submission process, which is achieved in the following manner: For every week in
 a resampled trace, we iterate over every existing user in the original trace.
@@
 
-  \item The simulation-based ("full feedback") strategy with hyperparameter
+  \item The simulation-based ("full feedback") strategy with hyper-parameter
     $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. This choice of
-    hyperparameter is a consequence of the natural rythm of human activity:
+    hyper-parameter is a consequence of the natural rhythm of human activity:
     daily or weekly simulation periods seem to facilitate comparability of
@@
 
-  \item The noisy simulation ("noisy feedback") variant with hyperparameter
+  \item The noisy simulation ("noisy feedback") variant with hyper-parameter
     $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
 
-  \item The bandit feedback strategy with hyperparameters $\lambda=1, \Delta
+  \item The bandit feedback strategy with hyper-parameters $\lambda=1, \Delta
     \in \left\{ 1 \text{day},1 \text{week} \right\}$. The choice of $\lambda=1$
@@
   \item An additional "Random" strategy to be used as a reference point, with
-    hyperparameter $ \Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
+    hyper-parameter $ \Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
     This strategy corresponds to choosing $P_i$ uniformly at random in $\left\{
@@
 that there are traces for which LQF is not optimal. Moreover, these
-are~\textit{a-posteriori} results, which means that we only know post-factum
-that this policy would have been better. In the present work, we are
-studying adaptive policies that obtain a good performance on any given system
-by selecting one of these policies.
+are~\textit{a-posteriori} results, which means that we only know
+textit{post-factum} that this policy would have been better. In the present
+work, we are studying adaptive policies that obtain a good performance on any
+given system by selecting one of these policies.
 
@@
 Figure~\ref{fig:all} reports the evolution of the average cumulative waiting
-time for the UniLu-Gaia tis quantityrace, which means that it shows how the
+time for the UniLu-Gaia, which means that it shows how the
 average improvement in cumulative waiting time evolves as a function of time.
@@
 This figure does not report the variability of the policies against one another,
-and does not allow to compare them beyong comparing their average performance.
+and does not allow to compare them beyond comparing their average performance.
 However, the fact that the band of percentiles are roughly of the size of half
@@
 good results, reducing the average waiting times of the baseline FCFS ordering
-policy by 11\% to 60\%. The second one is slighly less efficient with an
+policy by 11\% to 60\%. The second one is slightly less efficient with an
 improvement factor of 8\% to 46\%. However, it is easier to use and cheaper
@@
 
-This appendix describes the computational artifacts associated with this work. We
-take a lightweight approach to replicability by packaging a declarative
-workflow that generates all the figures from this paper and making it available
-publicly. Moreover, questions regarding replication can be freely directed at
-the authors via e-mail.
+This appendix describes the computational artifacts associated with this work.
+We take a lightweight approach to replicability by packaging and publishing a
+declarative workflow that generates all the figures from this paper. Moreover,
+questions regarding replication can be freely directed to the authors via
+e-mail.
 
@@
 among which distributing complete operating system images, containers, or
-packaging software. As our experiments are cpu bound, we decide to opt for the
-software packaging approach. This allows for replicating the experiments on clusters where virtualization is not
-available or kernels are too old for containers old and decreases runtime.
+packaging software. As our experiments are CPU bound, we decide to opt for the
+software packaging approach. This allows for replicating the experiments on
+clusters where virtualization is not available or kernels are too old for
+containers and decreases simulation runtime.
 
@@
 \begin{itemize}
-  \item {\bf algorithm: Scheduling Simulation Workflow}
-  \item {\bf program: Ocaml, R, Bash, Zymake(all sources and dependencies included)}
-  \item {\bf compilation: Ocaml 4.02.3, GNU bash version 4.4.5(1)-release, R version 3.3.2}
-  \item {\bf transformations: Patched nix standard environment.}
-  \item {\bf binary: Most binaries are cached by Nix channels.}
-  \item {\bf data set: Traces from the parallel workload archive, included.}
-  \item {\bf run-time environment: Nixpkgs 17.03}
-  \item {\bf hardware: Dell PowerEdge R730}
-  \item {\bf output: Files}
-  \item {\bf experiment workflow: \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|, see below}
-  \item {\bf publicly available?: Open-Sourced under the ISC~\cite{isc} license.}
+  \item {\bf Algorithm: Scheduling Simulation Workflow}
+  \item {\bf Program: Ocaml, R, Bash, Zymake(all sources and dependencies included)}
+  \item {\bf Compilation: Ocaml 4.02.3, GNU bash version 4.4.5(1)-release, R version 3.3.2}
+  \item {\bf Transformations: Patched nix standard environment.}
+  \item {\bf Binary: Most binaries are cached by Nix channels.}
+  \item {\bf Data set: Traces from the parallel workload archive, included.}
+  \item {\bf Run-time environment: Nixpkgs 17.03}
+  \item {\bf Hardware: Dell PowerEdge R730}
+  \item {\bf Output: Files}
+  \item {\bf Experiment workflow: \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|, see below}
+  \item {\bf Publicly available?: Open-Sourced under the ISC~\cite{isc} license.}
 \end{itemize}
@@
 The archive containing the experimental workflow can be obtained at
-\url{(omitted due to double-blind review)} Running the experiments can be done
+(URL omitted due to double-blind review). Running the experiments can be done
 on any platform equipped with the nix package manager by running at the root of
@@
 
- The experiments in this paper can run via in less than 24 hours on a
+ The experiments in this paper can run in less than 24 hours on a
  moderately parallel host. More precisely, we had access to a Dell PowerEdge
- R730 equipped with a total of 56 threads @2.4GHz each and 757G of RAM. For
+ R730 equipped with a total of 56 threads @2.4GHz each and 757 Gigabites of RAM. For
  this reason, the experimental campaign workflow was designed to be executed on
@@
 All the software dependencies of our experiments are automatically managed by Nix.
-The Nix software is available at \url{https://nixos.org/nix}.
+The Nix package manager is available at \url{https://nixos.org/nix}.
 
@@
 
-All data from this paper was obtained from the Swf Parallel Workloads
+All data from this paper were obtained from the Swf Parallel Workloads
 Archive~\cite{Feitelson20142967}. The data is cached in gzipped format in the
 persistent archive associated with this workflow, should the Parallel
-Workdloads Archive become unavailable.
+Workloads Archive become unavailable.
 
@@
 
-The experiment dependencies are loaded into the environment by the
+The experiment's dependencies are loaded into the environment by the
 \lstinline[basicstyle=\ttfamily]|nix-build| command.
@@
  All experiments are tied together using zymake~\cite{zymake}, a minimalistic
- workflow system designed for computational experiments\footnote{the zymake system
+ workflow system designed for computational experiments\footnote{The zymake system
  is also packaged by our nix expressions.}. This system is
@@
      this is principally using shell scripts for data manipulation (see file
-     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter| for the
+     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong_filter| for the
      filtering steps used) and ocaml code that implements the resampling method
@@
 
-Readers interested with modifying the experiments can obtain the zenodo archive
-at \url{https://truc.tgz}
-
-The experiments are replicated with the same procedure as before, this time
-invoking the build command with the local folder:
-
-\begin{lstlisting}
-nix-build -A banditSelection
-\end{lstlisting}
+Readers interested in modifying the experiments can obtain the zenodo archive
+at (URL omitted due to peer review).
 
@@
 
+The experiments can then be run as before via:
+
+\begin{lstlisting}
+nix-build -A banditSelection
+\end{lstlisting}
+
+Additionally, the experiments can be launched manually via:
+
+\begin{lstlisting}
+nix-shell -A banditSelection
+zymake -l localhost zymakefile
+\end{lstlisting}
+
 \subsection{Notes}

### 2017-04-03 paper/src/paper.tex
BackFilling

@@
 to study how to tune the ordering of the submission queues in the frame of the
-classical EASY-BackFilling family of heuristics using two new online approaches
+classical EASY-Backfilling family of heuristics using two new online approaches
 (an approach based on a simulator and an approach using multi-armed bandit).
@@
 More precisely, we showed that it is worth learning on the scheduling process
-itself. Indeed, reordering the submission queues under EASY-BackFilling leads
+itself. Indeed, reordering the submission queues under EASY-Backfilling leads
 to considerable gains in performance.

### 2017-04-03 paper/src/paper.tex
paper minor fixes.

@@
 
-\acmConference[SC’17]{Supercomputing}{May 29 -- June 2}{Orlando, FL, USA}
+\acmConference[SC’17]{Supercomputing}{November 12 -- 17}{Denver , CO, USA}
 
@@
 last few years, there was an explosion of the number of works at the interface
-of HPC and BigData fields, dealing with learning algorithms. Recent
+of the HPC and Big Data fields, dealing with learning algorithms. Recent
 sophisticated approaches developed \textit{ad hoc} supervised machine learning
@@
   \caption{Variability in the weekly average waiting time in the KTH-SP2 trace
-  (see Subsection~\ref{sub:traces}) for 7 different policies. The policy set is
+  (see Subsection~\ref{sub:traces}) for 7 different policies with $\Theta = 0$. The policy set is
   reduced as not to obstruct the figure.}
@@
 
-  \item The bandit feedback strategy with hyper-parameters $\lambda=1, \Delta
-    \in \left\{ 1 \text{day},1 \text{week} \right\}$. The choice of $\lambda=1$
-    means that we use no decay in the estimation of the overall cost. Choosing
-    to eliminate decay in this experimental campaign is natural since our
-    resampled data is in a stationary regime \textit{by construction}, as
-    explained in~\ref{ssub:resampling}.
+  \item The bandit feedback strategy with hyper-parameters $\epsilon=0.5,
+    \lambda=1, \Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. The
+    choice of $\lambda=1$ means that we use no decay in the estimation of the
+    overall cost. Choosing to eliminate decay in this experimental campaign is
+    natural since our resampled data is in a stationary regime \textit{by
+    construction}, as explained in~\ref{ssub:resampling}.
+    The choice of $\epsilon=0.5$ is a reasonable default.
 
@@
   \ra{1.3}
-  \caption{Average total waiting time diminution of fixed policies with respect to EASY(FCFS).}
+
+  \caption{Average total waiting time diminution of fixed policies with respect
+  to EASY(FCFS). The precise definition of the value reported is provided in
+Eq~\ref{eq:cum}}
 
@@
 \begin{equation}
+  \label{ref:cum}
   \frac{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(P_i,40h))-
@@
 %TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
-As these directions are data-bound, they will require careful application of
+As both these directions are data-bound, they will require careful application of
 trace resampling techniques in order to be successful.

### 2017-04-03 paper/src/paper.tex
figure fix.

@@
   \centering
-  \includegraphics[scale=0.6]{figures/full-ANL-Intr.pdf}
+  \includegraphics[scale=0.6]{figures/full-UniLu-Ga.pdf}
   \caption{Evolution of the average cumulative waiting time improvement
@@
   \centering
-  \includegraphics[scale=0.6]{figures/mosaicbandit-ANL-Intr.pdf}
-
-  \caption{Share of the policies chosen by Epsilon-Greedy as a function of
-  time. The average choice is obtained by resampling the original trace 60 
-times and aggregating by date.}
+  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}
+
+  \caption{Share of the policies chosen by Epsilon-Greedy($\Delta=$1 day,
+  $\epsilon = 0.5, \lambda = 1 $) as a function of time for the UniLu-Gaia
+trace. The average choice is obtained by resampling the original trace 60 times
+and aggregating by date.}
 
@@
 While this policy obtains an excellent performance, the convergence to a fixed
-policy is slow.  This is not harmful \textit{per se} and just means that the
-variability in the waiting times is indeed high. Additionally, we observe that
-the increasingly dominant policy, LCFS, is the second best overall policy. This
-is perhaps due to the estimation bias stemming from the boundary conditions of
-the weekly simulations.
+policy is slow.  This is not harmful \textit{per se} and just means that it is
+difficult to distinguish between the best policies. Additionally, we observe that
+the two increasingly dominant policy, LAF and LQF, are the best overall policies.
+This means that the estimation converges to the correct point.
 
@@
   \centering
-  \includegraphics[scale=0.6]{figures/mosaic-ANL-Intr.pdf}
-
-  \caption{Share of the policies chosen by the Noisy policy. The average choice
+  \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}
+
+  \caption{Share of the policies chosen by the Noisy policy as a function of time
+    for the UniLu-Gaia trace. The average choice
   is obtained by resampling the original trace 60 times and aggregating by

### 2017-04-03 paper/src/paper.tex
add hyperparameter in caption
add ccs

@@
 
-\acmConference[SC’17]{Supercomputing}{November 12 -- 17}{Denver , CO, USA}
+\acmConference[SC’17]{Supercomputing Conference}{November 12 -- 17}{Denver , Colorado, USA}
 
@@
 %\hyphenation{op-tical net-works semi-conduc-tor}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10010940.10010971.10010980.10010986</concept_id>
+<concept_desc>Software and its engineering~Massively parallel systems</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010169</concept_id>
+<concept_desc>Computing methodologies~Parallel computing methodologies</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010257.10010282.10010284</concept_id>
+<concept_desc>Computing methodologies~Online learning settings</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~Massively parallel systems}
+\ccsdesc[300]{Computing methodologies~Parallel computing methodologies}
+\ccsdesc[300]{Computing methodologies~Online learning settings}
+
+\keywords{Multi-Armed Bandits, EASY-Backfilling, Parallel Job Scheduling, High
+Performance Computing, Simulation.}
 
@@
   The EASY-FCFS heuristic is the basic building block of job scheduling
-  policies in most parallel High Performance Computing (HPC) platforms. Despite
+  policies in most parallel High Performance Computing platforms. Despite
   its good properties (simplicity, no starvation), it could still be improved
@@
   platforms. More precisely, we consider the problem of selecting a reordering
-  policy of the job queue under several feedback modes. We show via a
+  policy for the job queue under several feedback modes. We show via a
   comprehensive experimental campaign that noisy feedback (using a weak
@@
 
-Providing the computing infrastuctures needed to solve actual complex problems
-arising in the various fields of modern society (including climate change,
-health, green energy or security) is a strategic challenge. The main pillar of
-the answer to this challenge is to build extreme-scale parallel and distributed
-platforms. The never-ending race for more computing power and storage capacity
-does not only lead to sophisticated specific Exascale platforms, but the
+Providing the computational infrastuctures needed to solve actual complex
+problems arising in the various fields of modern society (including climate
+change, health, green energy or security) is a strategic challenge. The main
+pillar of the answer to this challenge is to build extreme-scale parallel and
+distributed platforms. The never-ending race for more computing power and
+storage capacity leads to sophisticated specific Exascale platforms, and the
 objective of the community is to design efficient sustained Petascale
@@
 to study how to tune the ordering of the submission queues in the frame of the
-classical EASY-Backfilling family of heuristics using two new online approaches
-(an approach based on a simulator and an approach using multi-armed bandit).
+classical EASY-Backfilling family of heuristics using two new online approaches.
+One approach is based on a simulator and the second uses a multi-armed bandit
+algorithm.
 
@@
 study more sophisticated policies that could leverage these data.  During the
-last few years, there was an explosion of the number of works at the interface
+last few years, there was an explosion in the number of works at the interface
 of the HPC and Big Data fields, dealing with learning algorithms. Recent
@@
 Unlike most existing studies, which consider the learning of specific
-parameters of a trace, we provide a complete system that uses feedback from the
-system for adapting the whole behavior online.  Our focus within this work is
-to learn how the management system behaves within the usual framework of
-EASY-BF. More precisely, we provide strategies for choosing the most
+parameters of a trace, we provide algorithms that use feedback from the
+system in order the whole behavior online. Our focus within this work is
+to learn how the platform behaves within the usual framework of
+EASY-Backfilling described in Section~\ref{sec:rw} . More precisely, we provide strategies for choosing the most
 appropriate queue reordering policy to be used for job selection. This approach
@@
 
-The first method (full feedback) continuously uses simulation on data gathered
+The first method, \textbf{full feedback}, continuously uses simulation on data gathered
 from the past and current system workload. In order to reflect the
@@
 
-The second method (bandit feedback) is a simulator-less approach based on a
+The second method, \textbf{bandit feedback}, is a simulator-less approach based on a
 multi-armed bandit algorithm, which derives feedback by measuring system
@@
 based on noisy data, which is closer to the actual conditions of uncertainties on
-both jobs and platform, also behaves very well. The alternate approach based on
+both jobs and platform, behaves similarly. The alternate approach based on
 a multi-armed bandit also provides reasonable results at a much lower cost,
@@
 significant computational overhead, which mainly explains why most of the
-machines of the top500 ranking~\cite{top500} rather use a variant of EASY-BF
+machines of the top500 ranking~\cite{top500} rather use a variant of EASY-Backfilling
 instead.  There is a large body of work seeking to improve/tune EASY. Indeed,
@@
 partially observable rewards. At every round, an action (represented by an arm
-of the Bandit) must be chosen in a fixed set and the corresponding reward is
+of the Bandit) must be chosen from a fixed set and the corresponding reward is
 observed. The goal of a MAB algorithm is to maximize the total reward obtained
-in a successive number of rounds. There exist multiple works addressing
-this problem under a variety of constraints, The two most popular settings are
-the original stochastic  case and the adversarial case. See~\cite{thompson} for
-the original work on the stochastic case and~\cite{Auer2002} for the "Upper
-Confidence Bound" family of algorithms. According to our knowledge, the work
-of~\cite{Banos} is the earliest for the adversarial case;
-see~\cite{nonstoch} for the "Exponential-weight" family of algorithms. We refer
-to~\cite{bubnow} for a comprehensive review of the field. While
-these algorithms bound the cumulative difference in loss to the best arm (the
-regret), they have functional constraints such as the fact that the rewards
-should be contained in a range. A simple heuristic called
-\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to achieve good
-results in practice, and does not have this requirement.
+in a successive number of rounds. There exist multiple works addressing this
+problem under a variety of constraints, The two most popular settings are the
+original stochastic  case and the adversarial case. See~\cite{thompson} for the
+original work on the stochastic case and~\cite{Auer2002} for the "Upper
+Confidence Bound" family of algorithms. To the best of our knowledge, the work
+of~\cite{Banos} is the earliest for the adversarial case; see~\cite{nonstoch}
+for the "Exponential-weight" family of algorithms. We refer to~\cite{bubnow}
+for a comprehensive review of the field. While these algorithms bound the
+cumulative difference in loss to the best arm (the regret), they have
+functional constraints such as the fact that the rewards should be contained in
+a range. The simple heuristic called \textit{Epsilon-Greedy} introduced
+in~\cite{Auer2002} does not have this requirement.  This heuristic will be used
+in the course of this paper and is described in Subsection~\ref{sub:bandit}
 
@@
 and some resource requirements (number and type of processing units). Then, the
-RJMS drives the search for the resources required to execute this job. Finally,
+Resource and Job Management System (RJMS) drives the search for the resources required to execute this job. Finally,
 the tasks of the job are assigned to the chosen nodes.
@@
 \textit{cumulative} version, which means that one seeks to minimize the average
-waiting time (\textbf{AvgWait}). It is worth noting that the \textbf{MaxWait},
-a.k.a the maximal value of the waiting time of all the jobs is also worthy of
-interest. Unfortunately, these criteria are antagonistic in nature. Section~\ref{sub:th} will outline our
-approach to address the bi-objective aspect of this problem.
+waiting time. It is worth noting that other metrics such as the maximum waiting
+time of all the jobs are also worthy of interest. Unfortunately, this criteria
+is antagonistic in nature with the average waiting time. Section~\ref{sub:th}
+will outline our approach to address the bi-objective aspect of this problem.
 
@@
   \item SEXP: Smallest Expansion Factor First
-  \item LRF: Largest Ratio $\frac{p_j}{q_j}$ First
+  \item LRF: Largest Ratio $\frac{\widetilde{p_j}}{q_j}$ First
   \item SRF: Smallest Ratio First
-  \item LAF: Largest Area $ p_j \times q_j$ First
+  \item LAF: Largest Area $ \widetilde{p_j} \times q_j$ First
   \item SAF: Smallest Area First
@@
 \label{sub:th}
-As existing works point out, reordering the job queue means losing the
-no-starvation guarantee and some individual jobs therefore can wait an undue
-amount of time. It is possible to introduce a thresholding mechanism in order
-to prevent this behavior: When a job's \textit{waiting time so far} exceeds a
-fixed threshold $\Theta$, it is jumped at the head of the queue. We denote by
-EASY($P,\Theta$) the scheduling policy that starts and backfill jobs according to
-the (thresholded) reordering policy $P$. For the sake of completeness,
-Algorithm~\ref{alg:EASY} describes the EASY($P,\Theta$) heuristic.
+
+Reordering the job queue means losing the no-starvation guarantee and some
+individual jobs therefore can wait an undue amount of time. It is possible to
+introduce a thresholding mechanism in order to prevent this behavior: When a
+job's \textit{waiting time so far} exceeds a fixed threshold $\Theta$, it is
+jumped at the head of the queue. We denote by EASY($P,\Theta$) the scheduling
+policy that starts and backfill jobs according to the (thresholded) reordering
+policy $P$. For the sake of completeness, Algorithm~\ref{alg:EASY} describes
+the EASY($P,\Theta$) heuristic.
 
@@
     of the queue (breaking ties in FCFS order).
-    \\ \textit{Starting jobs until the machine is full}
+    \\ \textit{//Starting jobs until the machine is full}
     \FOR {job $j$ in Q do}
@@
     of the currently running jobs.
-    \\ \textit{Backfilling jobs}
+    \\ \textit{//Backfilling jobs}
     \FOR {job $j'$ in $Q\setminus\{j\}$}
@@
 
-We further assume that there is a certain regularity among periods,
-i.e.\ that the distribution of the job submission process does not
-radically differ between consecutive periods. This assumption is validated in
-the study presented in~\cite{jsspp17}. It further entails that the behavior of
-a policy on the previous periods reflects its behavior on the current one, so
-that the selection of a policy can be based on its behavior on previous
-periods. However, there may be some variability between different periods for
-certain cost metrics~\cite{feitelson2001metrics}. This is illustrated in
+We further assume that there is a certain regularity among periods, i.e.\ that
+the distribution of the job submission process does not radically differ
+between consecutive periods. This assumption is validated in the study
+presented in~\cite{jsspp17}. It further entails that the behavior of a policy
+during the previous periods reflects its behavior on the current one, so that
+the selection of a policy can be based on its past behavior. However, there may
+be some variability between different periods for certain cost
+metrics~\cite{feitelson2001metrics}. This is illustrated in
 Figure~\ref{fig:mosn} which displays the average waiting time of various
@@
 average waiting time varies a lot from one week to the other, for all the 7
-policies considered. This indicates that when when the cost metric is averaged
+policies considered. This indicates that when the cost metric is averaged
 over different periods, there is a trade-off to find in between longer periods
@@
 
-The selection of a policy is of course reminiscent of reinforcement learning.
-It is important to note, however, that a pure reinforcement learning approach
-is difficult to develop in our context. Indeed, while we have outlined a
-reduced action space, the state space to consider is infinite. This is not
-prohibitive \textit{per se}, as modern methods~\cite{aprl}
-can bypass dimensionality issues via function approximation. However, these
-methods rely on large amounts of data, and we are studying online (often called
-\textit{on-policy} methods in reinforcement learning) methods. We rely here on
+The selection of a policy is reminiscent of reinforcement learning.  It is
+important to note, however, that a pure reinforcement learning approach is
+difficult to develop in our context. Indeed, while we have outlined a reduced
+action space, the state space to consider is infinite. This is not prohibitive
+\textit{per se}, as modern methods~\cite{aprl} can bypass dimensionality issues
+via function approximation. However, these methods rely on large amounts of
+data, and we are studying online methods\footnote{The Reinforcement learning
+  uses the term "\textit{on-policy}" for these methods.}. We rely here on
   simpler, yet we believe more effective, strategies to solve this problem.
@@
 The choice of a simulator is a critical part of experimental validation that
-raises a trade-off between precision and runtime. High precision can be obtained
-by carefully modeling the platform and its network topology, or extracting
-information from the jobs from their sources or post-mortem logs. This is the
-approach used by high-fidelity batch scheduling simulators such as
+raises a trade-off between precision and runtime. High precision can be
+obtained by carefully modeling the platform and its network topology, or
+extracting information from the jobs from their sources or post-mortem logs.
+This is the approach used by high-fidelity batch scheduling simulators such as
 Batsim~\cite{batsim}. In the present work, the focus is on studying the
-EASY-Backfilling mechanism in itself, without addressing the allocation problem.
-Moreover, the experimental protocol used here requires many simulation runs.
-Therefore, we set the precision/runtime trade-off at the point which minimizes
-simulation runtimes. We discard all topological information relative to the
-platform and use the job processing times of the original workloads. In this
-setup, the processors are considered to be undistinguishable from each other,
-and jobs can be discontinuously mapped to any available processor on the
-system. We develop a lightweight simulator\cite{ocst} and release it to the
-community. See the reproducibility paragraph below for more information.  With
-this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in 33
-seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
-2.00GHz. As a comparison, the Batsim simulator with network
-communication modeling disabled and no resource contiguity takes more than 7 hours to
-replay the same trace on the same hardware. The complete simulation campaign
-presented below is executed on a single Dell PowerEdge R730 in 22 hours,
-including input preparation and analysis code.
+EASY-Backfilling mechanism in itself, without addressing the allocation
+problem. Moreover, the experimental protocol used here requires many
+simulation runs. Therefore, we set the precision/runtime trade-off at the
+point which minimizes simulation runtimes. We discard all topological
+information relative to the platform and use the job processing times of the
+original workloads. In this setup, the processors are considered to be
+undistinguishable from each other, and jobs can be discontinuously mapped to
+any available processor on the system. During this work, we developed a
+lightweight simulator\cite{ocst} and now release it to the community along with
+this article. See the reproducibility paragraph below for more information.
+With this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in
+33 seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
+2.00GHz. As a comparison, the Batsim simulator with network communication
+modeling disabled and no resource contiguity takes more than 7 hours to replay
+the same trace on the same hardware. The complete simulation campaign presented
+below is executed on a single Dell PowerEdge R730 in 22 hours, including input
+preparation and analysis code.
 
@@
 
-We use devise the following experimental campaign. For each of the 7 workloads
+We devise the following experimental campaign. For each of the 7 workloads
 considered, we use the original data to resample 60 traces, each with a length
@@
 EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
-threshold $\Theta$ at 40 hours, a sane default. The choice of this default
-stems from the analysis developed in ~\cite{jsspp17}. This value is high
+threshold $\Theta$ at 40 hours. The choice of this default
+stems from the analysis developed in ~\cite{jsspp17}. Essentially, this value is high
 enough for queue reorderings to be leveraged while being low enough to provide
@@
   to EASY(FCFS). The precise definition of the value reported is provided in
-Eq~\ref{eq:cum}}
+Eq~\eqref{eq:cum}}
 
@@
 
-\begin{equation}
-  \label{ref:cum}
+\begin{equation} \label{eq:cum}
   \frac{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(P_i,40h))-
@@
 are~\textit{a-posteriori} results, which means that we only know
-textit{post-factum} that this policy would have been better. In the present
-work, we are studying adaptive policies that obtain a good performance on any
-given system by selecting one of these policies.
+textit{post-factum} that this policy would have been better.
 
@@
 
+In the present work, we are studying adaptive policies that obtain a good
+performance on any given system by selecting one of these policies.
 Figure~\ref{fig:all} reports the evolution of the average cumulative waiting
-time for the UniLu-Gaia, which means that it shows how the
-average improvement in cumulative waiting time evolves as a function of time.
-In other words, this is the average curve obtained if one were to plot the
-cumulative sum of job waiting times for every resampled trace, incrementing the
-sum when a job finishes. Accordingly, the final values attained by every curve
-correspond exactly to the values reported in Table~\ref{tab:full}. The dashed
-values represent the 10th and 90th percentile of the values.
+time for the UniLu-Gaia, which means that it shows how the average improvement
+in cumulative waiting time evolves as a function of time.  In other words, this
+is the average curve obtained if one were to plot the cumulative sum of job
+waiting times for every resampled trace, incrementing the sum when a job
+finishes. Accordingly, the final values attained by every curve correspond
+exactly to the values reported in Table~\ref{tab:full}. The dashed values
+represent the 10th and 90th percentile of the values.
 
@@
 value of the period length $\Delta$ has an impact on the performance of the
-random policy. Indeed, the shorter period leads to better performance. We
+random policy. Indeed, shorter period lead to better performance. We
 suggest that this is due to the properties of the system under study. Indeed,
@@
 
-  \caption{Share of the policies chosen by the Noisy policy as a function of time
+  \caption{Share of the policies chosen by the Noisy policy ($\Delta=$1 day,
+  $\lambda = 1 $) as a function of time
     for the UniLu-Gaia trace. The average choice
@@
 
-We presented in this work a new way of addressing this problem.  The idea was
+In this work, we presented a new way of addressing this problem. The idea was
 to look at the level of the scheduling process instead of trying to change its
@@
 policy by 11\% to 60\%. The second one is slightly less efficient with an
-improvement factor of 8\% to 46\%. However, it is easier to use and cheaper
-to run.
+improvement factor of 8\% to 46\%. However, the bandit-based method is easier
+to use and cheaper to run.
 
@@
 %@Denis: a essayer.. 1 mois de travail pour tester?
-Second, one can help but wonder what performance could be achieved by extending
+Second, one can't help but wonder what performance could be achieved by extending
 the search space to arbitrary policies. This means doing a complete study
@@
 Authors are sorted in alphabetical order. The authors would like to warmly
-thank Pierre Neyron for help with computational experiments and Arnaud Legrand
-for discussions about reproducibility. We gracefully thank the contributors of
-the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2), Travis Earheart
-and Nancy Wilkins-Diehr (SDSC Blue), Lars Malinowsky (KTH SP2), Dan Dwyer and
-Steve Hotovy (CTC SP2), Joseph Emeras (CEA Curie and UniLu Gaia), Susan
-Coghlan, Narayan Desay, Wei Tang (ANL Intrepid), and of course Dror Feitelson.
-The Metacentrum workload log was graciously provided by the Czech National Grid
-Infrastructure MetaCentrum. This work has been partially supported by the LabEx
+thank Pierre Neyron for help with computational experiments, as well as Michael
+Mercier, Salah Zrigui, Bruno Raffin and Theophile Terraz for their comments and
+Arnaud Legrand for his work on reproducibility. We gracefully thank the
+contributors of the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2),
+Travis Earheart and Nancy Wilkins-Diehr (SDSC Blue), Lars Malinowsky (KTH SP2),
+Dan Dwyer and Steve Hotovy (CTC SP2), Joseph Emeras (CEA Curie and UniLu Gaia),
+Susan Coghlan, Narayan Desay, Wei Tang (ANL Intrepid), and of course Dror
+Feitelson.  The Metacentrum workload log was graciously provided by the Czech
+National Grid Infrastructure MetaCentrum. This work has been partially
+supported by the LabEx
 PERSYVAL-Lab(\grantsponsor{ANR-11-LABX-0025-01}{ANR-11-LABX-0025-01}) funded by
 the French program Investissement d'avenir.
+
 \end{acks}

### 2017-04-21 paper/src/paper.tex
fix
last time fixes

@@
 change, health, green energy or security) is a strategic challenge. The main
-pillar of the answer to this challenge is to build extreme-scale parallel and
+pillar to the answer is to build extreme-scale parallel and
 distributed platforms. The never-ending race for more computing power and
@@
 parameters of a trace, we provide algorithms that use feedback from the
-system in order the whole behavior online. Our focus within this work is
+system in order the whole behavior online. Our focus is
 to learn how the platform behaves within the usual framework of
@@
 good scheduling policy on a per-system basis. In this work, we focus on the
-average waiting time of the jobs. Each method aims to choose the best queue
+average waiting time of the jobs. Each method aims at choosing the best queue
 reordering policy in a fixed search set of semantically diverse options.
@@
 that the required sample sizes for policy selection in batch schedulers are
-small enough for workload resampling to be avoided in practice. The variant
+small enough to avoid workload resampling in practice. The variant
 based on noisy data, which is closer to the actual conditions of uncertainties on
@@
 First-Serve approach to minimizing the waiting times is a common default
-strategy, it can absolutely be outperformed.  We show how to automate selection
+strategy, it can definitively be outperformed.  We show how to automate selection
 of a good alternative policy in two cases (depending on whether a simulator is
@@
 while the basic mechanism is used by some actual resource and job management
-software (most notably SLURM~\cite{SLURMdocSCHED}), this is rarely done
+software (most notably SLURM~\cite{SLURMdocSCHED}), this is seldom done
 without fine tuning by system administrators.
@@
 However, as successful as they may be, these works do not address the
-dependency of scheduling metrics on the workload~\cite{variability}. Indeed
-these studies most often report \textit{post-hoc} performance since they
+dependency of scheduling metrics on the workload~\cite{variability}.
+These studies most often report \textit{post-hoc} performance since they
 compare algorithms after the workload is known.
 
-The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
-method to tuning these queues, although it requires simulated scheduling runs
-at decision time and therefore, it costs much more than the natural execution of EASY.
+The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic method
+to tuning these queues. Although it requires simulated scheduling runs at
+decision time and therefore, making it much more than the natural execution of
+EASY.
 
@@
 There was a recent focus on leveraging the huge amount of data available in
-large scale computing platforms in order to improve system performance. Some
+large scale computing platforms to improve system performance. Some
 works use collaborative filtering to co-locate tasks in clouds by estimating
@@
 application level. For instance, ~\cite{fmodeling} uses binary classification
-to distinguish benign memory faults from application errors in order to execute
+to distinguish benign memory faults from application errors so as to execute
 recovery algorithms.
@@
 estimate uncertain parameters, we consider here a more pragmatic approach,
-which directly learns a good scheduling policy from a given policy space.
-
-Existing work~\cite{jsspp17} takes this approach in an offline manner, by
+that directly learns a good scheduling policy from a given policy space.
+
+Existing work~\cite{jsspp17} takes this approach offlines, by
 splitting workload data in a training/testing fashion. This work studies
@@
 in~\cite{Auer2002} does not have this requirement.  This heuristic will be used
-in the course of this paper and is described in Subsection~\ref{sub:bandit}
+in this paper and is described in Subsection~\ref{sub:bandit}
 
@@
 characteristics are only known when they are released.  Below is the brief
-description (with the corresponding notations) of the characteristics of job
+description of the characteristics of job
 $j$:
@@
 
-\subsection{Brief description of EASY Backfilling}
+\subsection{Brief Description of EASY Backfilling}
 \label{sub:easy}
@@
 The selection of the job to run is performed according to a scheduling policy,
-which determines the order in which the jobs are executed. EASY-Backfilling is
+which determines the order the jobs are executed. EASY-Backfilling is
 the most widely used policy due to its simple and robust implementation and
@@
 
-The EASY-FCFS heuristic uses a job queue to select and backfill jobs.  At any
-time requiring a scheduling decision (i.e.\ job submission or termination),
+EASY-FCFS uses a job queue to select and backfill jobs.  At any
+time a scheduling decision is required (i.e.\ job submission or termination),
 the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
@@
 it makes a reservation for this job at the earliest predictable time and
-starts \textit{backfilling} the job queue in FCFS order, starting any job that
+starts \textit{backfilling} the job queue in FCFS order using any job that
 does not delay the unique reservation.
@@
 job queue reordering and job thresholding. Together, these two building blocks
-constitute a robust framework for tuning EASY.
+make a robust framework for tuning EASY.
 
@@
 \begin{enumerate}
-  \item FCFS: First-Come First-Serve, which is the widely used default policy~\cite{easy}.
+  \item FCFS: First-Come First-Serve, which is the most commonly used policy~\cite{easy}.
   \item LCFS: Last-Come First-Serve.
@@
 individual jobs therefore can wait an undue amount of time. It is possible to
-introduce a thresholding mechanism in order to prevent this behavior: When a
-job's \textit{waiting time so far} exceeds a fixed threshold $\Theta$, it is
-jumped at the head of the queue. We denote by EASY($P,\Theta$) the scheduling
+introduce a thresholding mechanism to prevent this behavior: When a
+queued job's waiting time exceeds a fixed threshold $\Theta$, it is
+at the head of the queue. We denote by EASY($P,\Theta$) the scheduling
 policy that starts and backfill jobs according to the (thresholded) reordering
@@
     \\ \textit{//Starting jobs until the machine is full}
-    \FOR {job $j$ in Q do}
-    \IF {$j$ can be started given the current system use.}
+    \FOR {job $j$ in Q}
+    \IF {$j$ can be started given the current system use}
     \STATE Pop $j$ from Q
@@
     \FOR {job $j'$ in $Q\setminus\{j\}$}
-    \IF {$j'$ can be started without delaying the reservation on $j$.}
+    \IF {$j'$ can be started without delaying $j$'s reservation.}
     \STATE Pop $j'$ from $Q$
@@
 via function approximation. However, these methods rely on large amounts of
-data, and we are studying online methods\footnote{The Reinforcement learning
+data, and we are studying online methods\footnote{The reinforcement learning literature
   uses the term "\textit{on-policy}" for these methods.}. We rely here on
@@
 policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$
-during the period $\Delta_t$ is defined as:
+during period $\Delta_t$ is defined as:
 %
@@
 %
-$\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent
+wheret$\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent
 history (\textit{i.e.} recent periods). One then selects the policy $P$ that
@@
 simulation. Using the same simulator for generating the traces and estimating
-the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact}
+the cost as defined in Eqs \eqref{eq:cost-exact} and \eqref{eq:tot-cost-exact}
 would however be too optimistic and would represent an upper bound on what can
@@
 work on non simulated traces, we randomly introduce noise in the estimate of
-the policy cost defined by Eq.~\ref{eq:cost-exact} by rescaling it, either
+the policy cost defined by Eq.~\eqref{eq:cost-exact} by rescaling it, either
 down or up:
@@
 Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during
-the period $\Delta_t$. We define the cost of policy $P_i$ during the period
+the period $\Delta_t$. We define the cost of policy $P_i$ during period
 $\Delta_t$ as:
@@
 \mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary to ensure that policies
-remain comparable over time, as not normalizing would favor the policy that was
+remain comparable over time. Not normalizing would favor the policy that was
 first selected, even if this choice was random. This normalization however
@@
 large so as to avoid relying on too few points for estimating the cost, and it
-should not be too small to avoid extreme variations between periods. As
+should not be too small either to avoid extreme variations between periods. As
 explained in Section~\ref{sec:experiments}, we rely in this study on periods of
@@
 
-The experimental campaign used in this work uses a mixture of small and large
+The experimental campaign performed in this work uses a mixture of small and large
 systems from different periods. Table~\ref{tab:logs} presents the 7 logs, which
@@
 \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
-clean singular data. More precisely, we apply the following modifications:
+clean erroneous data:
 
@@
   \item If the number of allocated or requested cores of a job  exceeds the
-    size of the machine, we remove the job.
+    size of the machine, we remove the job;
 
@@
     available positive value as the request. If both are negative, we remove
-    the job.
+    the job;
 
@@
 below is executed on a single Dell PowerEdge R730 in 22 hours, including input
-preparation and analysis code.
+preparation and statistical analysis code.
 
@@
 in this work is subject to high variability. This variability can be so extreme
-that even a time-frame of a fews years can be insufficient to compare scheduling
-algorithms. In our experience inverting some months in the trace at random can
-lead to reversal in the results. This motivated decades of research in workload
+that even a time-frame of a fews years can be insufficient to compare
+scheduling algorithms. In our experience, randomly shuffling months in a trace
+at random is enough to generate contradictory conclusions about which algorithm
+is better. This motivated decades of research in workload
 modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} proposes to
@@
 implementation of these ideas.  More precisely, we want to sample the job
-submission process, which is achieved in the following manner: For every week in
-a resampled trace, we iterate over every existing user in the original trace.
-For each of these users, we choose a week of the original trace uniformly at
-random. The jobs from this user in this original week are chosen to be present
-in the week being resampled.  This simple mechanism enables us to preserve the
-dependency within a week.  Although the system state may not be independent
-from one week to an other, we rely on existing work~\cite{jsspp17} and assume
-for these experiments that the system is in a stationary state. While we
-do provide guidelines as how to incorporate tracking variant of our algorithms
-in Section~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
-this will not be studied in the experiments from this paper, and is left for
-future research.
+submission process, which is achieved in the following manner: (a) For every
+week in a resampled trace, we iterate over every existing user in the original
+trace; (b) For each of these users, we choose a week of the original trace
+uniformly at random; (c) The jobs from this user in this original week are
+chosen to be present in the week being resampled.  This simple mechanism
+enables us to preserve the dependency within a week.  Although the system state
+may not be independent from one week to an other, we rely on existing
+work~\cite{jsspp17} and assume for these experiments that the system is in a
+stationary state. While we do provide guidelines as how to incorporate tracking
+variant of our algorithms in Section~\ref{sec:online} by introducing a decaying
+discount factor $\lambda$, this will not be studied in the experiments, and is left for future research.
 
@@
 
-We devise the following experimental campaign. For each of the 7 workloads
+For each of the 7 workloads
 considered, we use the original data to resample 60 traces, each with a length
@@
 
-  \item All fixed $P_i$ for $1<i<12$ from the search set defined in
-    ~\ref{subsec:policies}.  This is done for two reasons. First, it allows to
+  \item All fixed $P_i$ for $1 \leq i \leq 12$ from the search set defined in
+    Subsection ~\ref{subsec:policies}.  This is done for two reasons. First, it allows to
     see which policy works well for which platform. Second, it allows to
     compare how well strategies proposed in this work with the best
-    \textit{a-posteriori} known policy.
-
-  \item The simulation-based ("full feedback") strategy with hyper-parameter
-    $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. This choice of
-    hyper-parameter is a consequence of the natural rhythm of human activity:
-    daily or weekly simulation periods seem to facilitate comparability of
-    performance between two periods. Indeed, arbitrary period values would
-    needlessly increase the variability of the cost metric.
-
-  \item The noisy simulation ("noisy feedback") variant with hyper-parameter
-    $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
-
-  \item The bandit feedback strategy with hyper-parameters $\epsilon=0.5,
-    \lambda=1, \Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. The
-    choice of $\lambda=1$ means that we use no decay in the estimation of the
-    overall cost. Choosing to eliminate decay in this experimental campaign is
-    natural since our resampled data is in a stationary regime \textit{by
-    construction}, as explained in~\ref{ssub:resampling}.
-    The choice of $\epsilon=0.5$ is a reasonable default.
-
-  \item An additional "Random" strategy to be used as a reference point, with
-    hyper-parameter $ \Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
-    This strategy corresponds to choosing $P_i$ uniformly at random in $\left\{
-      P_1 , \cdots, P_{12} \right\}$ during every period $\Delta_t$.
+    \textit{a-posteriori} known policy;
+
+  \item For all the above strategies based on exact simulation (referred to as
+    "full feedback" in the remainder), on noisy simulation (referred to as
+    "noisy feedback") and on $\epsilon$-greedy exploration (referred to as
+    "bandit feedback"), we experiment with two policy periods: $\Delta \in
+    \left\{ 1 \text{day},1 \text{week} \right\}$. This choice is a consequence
+    of the natural rhythm of human activity: Daily or weekly simulation periods
+    seem to ease comparing of performance between two periods.
+    Indeed, arbitrary period values would needlessly increase the variability
+    of the cost metric. For the bandit feedback strategy, $\epsilon$ is set to $0.5$ and
+    $\lambda$ to $1$. This means that we do not use any decay in the estimation
+    of the overall cost. Choosing to eliminate decay in this experimental
+    campaign is natural since our resampled data is in a stationary regime
+    \textit{by construction}, as explained in~\ref{ssub:resampling}. The choice
+    of $\epsilon=0.5$ is a reasonable default;
+
+  \item Lastly, an additional "Random" strategy is used as a reference point,
+    with, as before, $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
+    This strategy corresponds to choosing $P_i$ uniformly at random in
+    $\left\{P_1 , \cdots, P_{12} \right\}$ during every period $\Delta_t$.
 
@@
 The experimental campaign consists of 7 traces $\times$ 60 resamples $\times$
-(12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
+(12+2+2+2) algorithms, which correspond to 12600 scheduling simulations spanning
 two years.  This means a total simulated timespan of 25200 years. Moreover,
@@
 simulations as part of their procedure. Indeed, these strategies re-simulate
-the system 12 times every at every period $\Delta$, which bumps the total
+the system 12 times at every period $\Delta$, which bumps the total
 effective simulation timespan to 92400 years. This motivates the use of a
@@
 invited to peruse the 'artifacts description' appendix of this paper. We
-provide to persistent archives containing the simulator, the analysis code and
+provide to persistent~\cite{zenodo} archives containing the simulator, the analysis code and
 an automated workflow that replicates all experiments and figures present in
@@
     \hline
-    ANL-Intrepid & -25\% & -9\%           & -3\%  & \textbf{-28\%} & -2\%  & -21\% & -24\%          & -21\%          & -10\% & -19\%        & +5\%  \\
     KTH-SP2      & -13\% & \textbf{-16\%} & +5\%  & \textbf{-16\%} & +3\%  & -8\%  & -15\%          & -8\%           & -13\% & -12\%        & +15\% \\
@@
     SDSC-BLUE    & -26\% & -29\%          & 0\%   & -29\%          & +17\% & -17\% & -26\%          & -15\%          & -15\% & \textbf{-32\%} & +23\% \\
+    ANL-Intrepid & -25\% & -9\%           & -3\%  & \textbf{-28\%} & -2\%  & -21\% & -24\%          & -21\%          & -10\% & -19\%        & +5\%  \\
     CEA-Curie    & -46\% & -23\%          & -45\% & \textbf{-57\%} & -19\% & -51\% & -23\%          & \textbf{-57\%} & -24\% & -40\%        & -17\% \\
@@
 
-We begin by drawing the attention of the reader to Table~\ref{tab:full}, which
+We begin by drawing the attention of the reader to Table~\ref{tab:full} that
 reports the average total waiting time performance improvement over
@@
 us denote the resampled traces of a given workload by $T_k$, where k is the
-index of the resampled trace ($1 \le k \le 60$). We also denote the number of
-jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by $\bar{w}(T_k,S)$
-the cumulative waiting time obtained by a simulation run for a strategy $S$:
+index of the resampled trace ($1 \le k \le 60$). Let the number of jobs in the
+resampled trace $T_k$ be $l(T_k)$. Let $\bar{w}(T_k,P)$ be the cumulative
+waiting time obtained by a simulation run for a strategy $P$:
 
@@
 attainable performance in the search set defined in~\ref{subsec:policies}.
-Note that there seems to be some regularity; the SAF policy is remarkably
-inefficient for instance, and the LQF policy works very well. However, note
-that there are traces for which LQF is not optimal. Moreover, these
+Note that there seems to be some regularity; for instance the SAF policy is remarkably
+inefficient and the LQF policy works very well. However there are traces for which LQF is not optimal. Moreover, these
 are~\textit{a-posteriori} results, which means that we only know
-textit{post-factum} that this policy would have been better.
+\textit{post-factum} that this policy would have been better.
 
@@
   \caption{Evolution of the average cumulative waiting time improvement
-    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
+    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 12$ on the
     UniLu-Gaia trace. The average is obtained by resampling the original trace
@@
 Figure~\ref{fig:all} reports the evolution of the average cumulative waiting
-time for the UniLu-Gaia, which means that it shows how the average improvement
+time for the UniLu-Gaia trace. This figure shows how the average improvement
 in cumulative waiting time evolves as a function of time.  In other words, this
@@
 This figure does not report the variability of the policies against one another,
-and does not allow to compare them beyond comparing their average performance.
-However, the fact that the band of percentiles are roughly of the size of half
+and does not allow to compare them beyond their average performance.
+Tthe fact that the size of the percentile band is roughly half
 the best attainable improvement was our original motivation for this work, as
@@
     times. The dashed lines represent the 10th and 90th percentiles of the
-  values across this resampling. Each figure is a different trace, and this
-figure is followed-up in figure~\ref{fig:follow} for the UniLu-Gaia log.}
+  values across this resampling. Each figure is a different trace. This figure
+is followed-up in figure~\ref{fig:follow} for the remaining UniLu-Gaia trace.}
 
@@
 Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
-result concerning the strategies proposed in this paper. These figures show the
+results concerning the strategies proposed in this paper. These figures show the
 average cumulative improvement of over EASY(FCFS,40h) of the three strategies
@@
     $\Delta$     & week   & day   & week           & day            & week  & day   & week   & day \\
-    CTC-SP2      & -20\%  & -23\% & -44\% & -44\% & -44\% & -44\% & -26\%  & -31\% \\
-    KTH-SP2      & -6\%   & -8\%  & -12\% & -11\% & -12\% & -12\% & -7\%   & -10\% \\
-    SDSC-SP2     & -4\%   & -7\%  & -13\% & -13\% & -13\% & -13\% & -5\%   & -8\% \\
-    SDSC-BLUE    & -12\%  & -14\% & -31\% & -31\% & -30\% & -29\% & -12\%  & -17\% \\
-    ANL-Intrepid & -13\%  & -20\% & -22\% & -26\% & -19\% & -24\% & -13\%  & -20\% \\
-    CEA-Curie    & -35\%  & -42\% & -53\% & -47\% & -53\% & -48\% & -36\%  & -46\% \\
-    Unilu-Gaia   & -29\%  & -31\% & -59\% & -60\% & -58\% & -58\% & -33\%  & -34\% \\
+    KTH-SP2      & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
+    CTC-SP2      & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
+    SDSC-SP2     & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
+    SDSC-BLUE    & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
+    ANL-Intrepid & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
+    CEA-Curie    & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
+    Unilu-Gaia   & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
     \hline
@@
 improvement over EASY(FCFS,40h) of all the proposed strategies along with the
-random variant.
-
-We make the following observations.
+random variant.  We make the following observations.
 
@@
 value of the period length $\Delta$ has an impact on the performance of the
-random policy. Indeed, shorter period lead to better performance. We
-suggest that this is due to the properties of the system under study. Indeed,
-when a bad policy is used, the job queue builds up. By switching more
+random policy. Indeed, shorter periods lead to better performance. We
+ that this is due to the properties of the system under study.
+When a bad policy is used, the job queue builds up. By switching more
 frequently from bad to good policies, the system never has time to build up the
@@
 accumulation of waiting jobs, there is less contribution to system inertia from
-good decisions, as the number of jobs in queue is below bounded by zero. This
+good decisions, as the number of jobs in queue can not be negative. This
 interpretation also means that randomized bandit strategies are robust by
@@
 
-\textbf{The Simulated Feedback strategy consistently performs as well as the
-best fixed policy.} The values attained by the simulation-based strategy match
+\textbf{The Simulated Feedback strategy consistently performs almost as well as the
+best fixed policy.} The values attained by the exact simulation based strategy almost match
 the highlighted values from Table~\ref{tab:full}. This corresponds to the curve
@@
 \textbf{Noisy simulations can be used.} The performance attained by the noisy
-variant almost matches that of the simulation-based strategy. This indicates
+variant almost matches that of the exact-simulation based strategy. This indicates
 that it is not necessary to have access to a precise simulator. Rather, we may
@@
 a quarter to the whole gap between the Random and Full Feedback strategies.
-
-\begin{figure}[]
-  \centering
-  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}
-
-  \caption{Share of the policies chosen by Epsilon-Greedy($\Delta=$1 day,
-  $\epsilon = 0.5, \lambda = 1 $) as a function of time for the UniLu-Gaia
-trace. The average choice is obtained by resampling the original trace 60 times
-and aggregating by date.}
-
-  \label{fig:mosb}
-\end{figure}
-
-Figure~\ref{fig:mosb} gives insight regarding the behavior of the Noisy
-(simulated) Feedback policy with $\Delta=$ 1 week. This figure reports the
-share of the policies chosen by the strategy on average as a function of time.
-This is an area plot, which means that the height of the colored region
-represents the value. Moreover, the proportion of times each policies are
-chosen by the strategy is aggregated across resampled traces.
-
-While this policy obtains an excellent performance, the convergence to a fixed
-policy is slow. This is not harmful \textit{per se} and just means that it is
-difficult to distinguish between the best policies. Additionally, we observe that
-the two increasingly dominant policy, LAF and LQF, are the best overall policies.
-This means that the estimation converges to the correct point.
 
@@
 
-Figure~\ref{fig:mosn} is the analogue of Figure~\ref{fig:mosb}, this time for
-the Epsilon-Greedy policy with $\Delta$ = 1 week.
-
-Observe that this time the choice is much less stable. The convergence is
+\textbf{Stability of the policy selection.}
+Figure~\ref{fig:mosn} gives insight regarding the behavior of the noisy
+feedback policy with $\Delta=$ 1 week. This figure reports the
+share of the policies chosen by the strategy on average as a function of time.
+This is an area plot: The height of the colored region
+represents the value. Moreover, the proportion of times each policy is
+chosen by the strategy is aggregated across resampled traces.
+
+While this policy obtains an excellent performance, the convergence to a fixed
+policy is slow. This is not harmful \textit{per se} and just means that it is
+difficult to distinguish between the best policies. Additionally, we observe that
+the two increasingly dominant policies, LAF and LQF, are the best overall policies.
+This means that the estimation converges to the correct point.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}
+
+  \caption{Share of the policies chosen by Epsilon-Greedy($\Delta=$1 day,
+  $\epsilon = 0.5, \lambda = 1 $) as a function of time for the UniLu-Gaia
+trace. The average choice is obtained by resampling the original trace 60 times
+and aggregating by date.}
+
+  \label{fig:mosb}
+\end{figure}
+
+Figure~\ref{fig:mosb} is the analogue of Figure~\ref{fig:mosn}, this time for
+the Epsilon-Greedy policy with $\Delta$ = 1 week.  Observe that here the choice of policy is much less stable. The convergence is
 extremely slow, which is expected given the variability of the reward we give
@@
 Second, one can't help but wonder what performance could be achieved by extending
-the search space to arbitrary policies. This means doing a complete study
-including lower bounding the cost and studying learning performance using some
+the search space to arbitrary policies. This means studying learning performance using some
 function approximator.
@@
 As both these directions are data-bound, they will require careful application of
-trace resampling techniques in order to be successful.
+trace resampling techniques to be successful.
 
